### 2021-11-15 Paper_online_Learning.Rmd
update papaeR

---
title: "Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends"
author:
- familyname: Talagala 
  othernames: Priyanga Dilini
  address: Department of Compuational Mathematics, University of Moratuwa, Sri Lanka
  email: priyangad@uom.lk
  correspondingauthor: true
- familyname: Talagala
  othernames: Thiyanga
  address: Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
  email: tstalagala@gmail.com
abstract: "A brief summary of our work"
keywords: "rmarkdown, templates"
wpnumber: 1
jelcodes: C10,C14,C22
blind: false
cover: true
toc: false
bibliography: references.bib
biblio-style: authoryear-comp
output:
  templatehub::workingpaper:
    fig_caption: yes
    fig_height: 10
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
```

```{r main, child='main.Rmd'}
```

### 2021-11-15 Paper_online_Learning.tex
update papaeR

\documentclass[11pt,a4paper,]{article}
\usepackage{lmodern}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
            pdfkeywords={rmarkdown, templates},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
\usepackage[style=authoryear-comp,]{biblatex}
\addbibresource{references.bib}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\title{Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends}

%% MONASH STUFF

%% CAPTIONS
\RequirePackage{caption}
\DeclareCaptionStyle{italic}[justification=centering]
 {labelfont={bf},textfont={it},labelsep=colon}
\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}

%% FONT
\RequirePackage{bera}
\RequirePackage{mathpazo}

%% HEADERS AND FOOTERS
\RequirePackage{fancyhdr}
\pagestyle{fancy}
\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
\makeatletter
\lhead{\textsf{\expandafter{\@title}}}
\makeatother
\rhead{}
\cfoot{}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}
\fancypagestyle{plain}{%
\fancyhf{} % clear all header and footer fields
\fancyfoot[C]{\sffamily\thepage} % except the center
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%% MATHS
\RequirePackage{bm,amsmath}
\allowdisplaybreaks

%% GRAPHICS
\RequirePackage{graphicx}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.8}

%\RequirePackage[section]{placeins}

%% SECTION TITLES
\RequirePackage[compact,sf,bf]{titlesec}
\titleformat{\section}[block]
  {\fontsize{15}{17}\bfseries\sffamily}
  {\thesection}
  {0.4em}{}
\titleformat{\subsection}[block]
  {\fontsize{12}{14}\bfseries\sffamily}
  {\thesubsection}
  {0.4em}{}
\titlespacing{\section}{0pt}{*5}{*1}
\titlespacing{\subsection}{0pt}{*2}{*0.2}

%% TITLE PAGE
\def\Date{\number\day}
\def\Month{\ifcase\month\or
 January\or February\or March\or April\or May\or June\or
 July\or August\or September\or October\or November\or December\fi}
\def\Year{\number\year}

\makeatletter
\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
\def\nojel{\def\showjel{}}
\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
\def\cover{{\sffamily\setcounter{page}{0}
        \thispagestyle{empty}
        \vspace*{2cm}
        \begin{center}
        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
                \LARGE\@author\end{onehalfspace}
        }}
        \end{center}
        \vfill
                \begin{center}\Large
                \Month~\Year\\[1cm]
                Working Paper \@wp
        \end{center}\vspace*{2cm}}}
\def\pageone{{\sffamily\setstretch{1}%
        \thispagestyle{empty}%
        \vbox to \textheight{%
        \raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
        \vspace{2cm}\par
        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
        \hspace{1cm}\showjel\vss}}}
\def\blindtitle{{\sffamily
     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
        }}
\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}

\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
\let\maketitle\titlepage
\makeatother

%% SPACING
\RequirePackage{setspace}
\spacing{1.5}

%% LINE AND PAGE BREAKING
\sloppy
\clubpenalty = 10000
\widowpenalty = 10000
\brokenpenalty = 10000
\RequirePackage{microtype}

%% PARAGRAPH BREAKS
\setlength{\parskip}{1.4ex}
\setlength{\parindent}{0em}

%% HYPERLINKS
\RequirePackage{xcolor} % Needed for links
\definecolor{darkblue}{rgb}{0,0,.6}
\RequirePackage{url}

\makeatletter
\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
\makeatother
\hypersetup{
     citecolor=0 0 0,
     breaklinks=true,
     bookmarksopen=true,
     bookmarksnumbered=true,
     linkcolor=darkblue,
     urlcolor=blue,
     citecolor=darkblue,
     colorlinks=true}

%% KEYWORDS
\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}

%% ABSTRACT
\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
  {\end{minipage}}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage[showonlyrefs]{mathtools}
\usepackage[no-weekday]{eukdate}

%% BIBLIOGRAPHY

\makeatletter
\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
\makeatother
\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}

\DeclareFieldFormat{url}{\texttt{\url{#1}}}
\DeclareFieldFormat[article]{pages}{#1}
\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
\DeclareFieldFormat[inproceedings]{title}{#1}
\DeclareFieldFormat{shorthandwidth}{#1}
% No dot before number of articles
\usepackage{xpatch}
\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
% Remove In: for an article.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{in}\intitlepunct}}}

\makeatletter
\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
\makeatother
\renewcommand*{\finalnamedelim}{%
  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
  \addspace\&\space}

\wp{1}
\jel{C10,C14,C22}

\RequirePackage[absolute,overlay]{textpos}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}
\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}

\author{Priyanga Dilini~Talagala, Thiyanga~Talagala}
\addresses{\textbf{Priyanga Dilini Talagala}\newline
Department of Compuational Mathematics, University of Moratuwa, Sri Lanka
\newline{Email: \href{mailto:priyangad@uom.lk}{\nolinkurl{priyangad@uom.lk}}}\newline Corresponding author\\[1cm]
\textbf{Thiyanga Talagala}\newline
Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
\newline{Email: \href{mailto:tstalagala@gmail.com}{\nolinkurl{tstalagala@gmail.com}}}\\[1cm]
}

\date{\sf\Date~\Month~\Year}
\makeatletter
 \lfoot{\sf Talagala, Talagala: \@date}
\makeatother

%% Any special functions or other packages can be loaded here.

\begin{document}
\maketitle
\begin{abstract}
A brief summary of our work
\end{abstract}
\begin{keywords}
rmarkdown, templates
\end{keywords}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The ongoing pandemic of COVID-19 has now become the most important priority of governments and media of many countries all around the world. Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International (PHEIC) Concern on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}. As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}. In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment to minimize the spread of the virus. In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.

The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with a Chinese national who came to Sri Lanka as a tourist. The first locally transmitted COVID-19 case in Sri Lanka was then reported on 11 March 2020 followed by an increasing trend of the COVID-19 confirmed cases. In response to this increasing trend of COVID 19 cases, lockdowns and curfews were imposed several times across the country with the aim of reducing the spread of the virus. As of 29 September 2021, 515524 cases of COVID-19 have been reported in Sri Lanka, resulting with 12786 deaths \autocite{HealthSL2020,dong2020interactive}.

\begin{figure}[h]

{\centering \includegraphics[width=1\textwidth]{figure/covidImpactWorld-1}

}

\caption{Global tracking of COVID-19 caused school closures and re-openings}\label{fig:covidImpactWorld}
\end{figure}

\begin{figure}[h]

{\centering \includegraphics[width=1\textwidth]{figure/covidImpactContinent-1}

}

\caption{Region wise tracking of COVID-19 caused school closures and re-openings}\label{fig:covidImpactContinent}
\end{figure}

The ongoing COVID-19 pandemic has brought the world to a standstill. Many sectors have been affected ever since the outbreak of COVID 19, worldwide. Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}. Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020. With the start of the pandemic a near total closure of schools were observed all across the globe. Africa and Oceania are in better positions in comparison to other regions with their increasing trend of fully open schools. In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}.

The speed of these closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders. Sudden unexpected movement from classrooms to homes at large scale made students, teachers and parents vulnerable. The parents with limited education and resources; and students who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement. Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services. Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning. Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}.

Eventhough distance education has a long history that goes back to almost two centuries \autocite{spector2014handbook} with significant modifications,
alterations, and addition in the process of delivery and communication \autocite{moore2011learning}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously. Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to unexpected massive shutdown worldwide. Lack of experience, lack of awareness, lack of readiness for distance education put various stakeholders in the education system in panic, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic Figures \ref{fig:distanceLearningWorldAnalysis}. The search spikes coincide with increasing Covid-19 counts and related google trend searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.

\begin{figure}[h]

{\centering \includegraphics[width=1\textwidth]{figure/distanceLearningWorldAnalysis-1}

}

\caption{Google trends footprint analysis}\label{fig:distanceLearningWorldAnalysis}
\end{figure}

In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions. In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education? (2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic? (3) Which solutions have a wide attention and public interest both at the national and global level?

Confusion and stress for teachers: When schools close, especially unexpectedly and for unknown duration, teachers are often unsure of their obligations and how to maintain connections with students to support learning. Transitions to distance learning platforms tend to be messy and frustrating, even in the best circumstances. Limited time, and lack of knowledge about the best online tools and techniques available are other major challenges teachers encounter during the pandemic. In many contexts, school closures lead to furloughs or separations for teachers.\\
Parents unprepared for distance and home schooling: When schools close, parents are often asked to facilitate the learning of children at home and can struggle to perform this task. This is especially true for parents with limited education and resources.
Challenges creating, maintaining, and improving distance learning: Demand for distance learning skyrockets when schools close and often overwhelms existing portals to remote education. Moving learning from classrooms to homes at scale and in a hurry presents enormous challenges, both human and technical.

The design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content. It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves \autocite{moore2011learning}.

As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal \& Wilson, 2010; Volery \& Lord, 2000) \autocite{moore2011learning}. There are many different distance learning solutions. Identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features. The high popularity or increasing online attention towards a product can indicate its prominence in the market, its usability and its impact. Further, developers also tend to improve the their products and serivices in response to increasing online search \autocite{willis2020using}

A big question for faculty members designing new online courses concerns their role as a teacher: what do I do when I teach online? Designing the course materials, interacting with students and giving them feedback, and assessing student work are obvious tasks for the online teacher. What is not
obvious, however, is how those tasks are carried out.

As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, \& Watts, 2010), and politics (Reilly, Richey, \& Taylor, 2012). We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study \autocite{vaughan2014web}. However their (\textcite{vaughan2014web}) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.

We primarily analyze quantitative digital footprint data on the Internet from January 2020 to 11.03.2020, including their representativeness.The google trend footprint will allow the teachers to narrow down thier search space and select prominent distance education solutions for their teaching purposes.

The remainder of this paper is organised as follows. Section 2 presents the related work to lay the foundation for the stray algorithm. Section 3 describes the limitations of the HDoutliers algorithm that hinder its performance. Section 4 presents the improved algorithm, stray, that addresses the limitations of the HDoutliers algorithm. Section 5 presents a comprehensive evaluation, illustrating the key features of the stray algorithm. Section 6 includes an application of stray algorithm related to pedestrian behaviour in the city of Melbourne, Australia. Section 7 concludes the article and presents future research directions.

\hypertarget{background}{%
\subsection{Background}\label{background}}

\hypertarget{benefits-of-competition}{%
\subsubsection{Benefits of Competition}\label{benefits-of-competition}}

As Uber was queried more on Google, he was able to find a decrease in complaints of broken credit card machines as the taxis improved their service in response to the increased. Complaints may provide a proxy of the quality of traditional taxis in the city and thus provide a means to investigate the service quality differences in yellow taxis since the arrival of Uber \autocite{willis2020using}.

In total, our approach to employ novel data from a news media outlet to approximate the public interest and Uber's popularity enabled us to illustrate how Uber disrupted the New York City well established taxi market.

Both of these behavioural mechanisms can potentially explain the increase in the number of complaints about taxi service.
\#\# Analysis

\url{https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/}

Google has analyzed the billions of search requests it processes every day, and identified the terms that have had the highest spike this year compared to 2019. Unsurprisingly, ``coronavirus'' topped the overall list.

The search trends show the impact of the COVID pandemic on so many aspects of life. Relatively obscure teleconferencing company Zoom became a household name (and highly profitable business) as employees adapted to working from home, and friends and families turned to video calls instead of meeting in person .

Google Classroom was one of several online resources that provided education to millions of children affected by school closures. Searches for ``how to foster a dog'' reached an all-time high as social distancing triggered a yearning for the companionship of a pet.

\url{https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/}

Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception as- sociated with health-threatening events such as an Ebola outbreak {[}5, 37{]}. Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease \autocite{alicino2015assessing}.

The Internet represents a great opportunity for public health agencies to disseminate healthcare-related infor- mation quickly, effectively and cheaply, provided that re- liable news are shared with as much of the population as possible, both in developed and developing countries, where the diffusion of new technologies among the population is rapidly increasing. Google Trends -- a pub- licly available data source -- could be used as a proxy of the proper diffusion of strategies based on health educa- tion messages, allowing to fill the translation gap be- tween best evidence (what some experts know) and practice (what most people do or believe) {[}38{]}. More- over, it could be exploited to determine the level of up- take of information about a public health event. Even during the 2014 Ebola outbreak, some stakeholders lev- eraged new technologies to plan proper communication strategies and address public concern {[}39, 40 \autocite{alicino2015assessing}.

Hara and Kling (2000) argue that more research about online courses from the students' perspective is needed. In a case study of graduate students enrolled in an online course, they found that many aspects of the course frustrated students, and that the instructor remained unaware of the continuing level of frustration. They suggest that many researchers bring an optimistic, romantic view of technology that may dampen their ability to look at hard questions and apply rigorous research methods. Much of the research, they maintain, has been advocacy and theorizing about future possibilities \autocite{wallace2003online}.

\printbibliography

\end{document}

### 2021-11-15 main.Rmd
update papaeR

--- 2021-11-15 Paper_online_Learning.tex
@@
-\documentclass[11pt,a4paper,]{article}
-\usepackage{lmodern}
+```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
+library(knitr)
+opts_chunk$set(
+  warning = FALSE,
+  message = FALSE,
+  echo = FALSE,
+  fig.path = 'figure/',
+  cache.path = 'cache/',
+  fig.align = 'center',
+  fig.show = 'hold',
+  cache = FALSE,
+  external = TRUE,
+  dev = "png",
+  #fig.height = 8,
+  #fig.width = 12,
+  out.width = "1\\textwidth",
+  size= 'small',
+  fig.pos = "h" ##  h, here, t, top, b, bottom, p, page of float
+  #fig.asp = 0.5
+)
+writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
+read_chunk('src/main.R')
+```
 
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
-\hypersetup{
-            pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
-            pdfkeywords={rmarkdown, templates},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{geometry}
-\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
-\usepackage[style=authoryear-comp,]{biblatex}
-\addbibresource{references.bib}
-\usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
+```{r load, cache=FALSE}
+```
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\title{Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends}
-
-%% MONASH STUFF
-
-%% CAPTIONS
-\RequirePackage{caption}
-\DeclareCaptionStyle{italic}[justification=centering]
- {labelfont={bf},textfont={it},labelsep=colon}
-\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
-\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
-
-%% FONT
-\RequirePackage{bera}
-\RequirePackage{mathpazo}
-
-%% HEADERS AND FOOTERS
-\RequirePackage{fancyhdr}
-\pagestyle{fancy}
-\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
-\makeatletter
-\lhead{\textsf{\expandafter{\@title}}}
-\makeatother
-\rhead{}
-\cfoot{}
-\setlength{\headheight}{15pt}
-\renewcommand{\headrulewidth}{0.4pt}
-\renewcommand{\footrulewidth}{0.4pt}
-\fancypagestyle{plain}{%
-\fancyhf{} % clear all header and footer fields
-\fancyfoot[C]{\sffamily\thepage} % except the center
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}}
-
-%% MATHS
-\RequirePackage{bm,amsmath}
-\allowdisplaybreaks
-
-%% GRAPHICS
-\RequirePackage{graphicx}
-\setcounter{topnumber}{2}
-\setcounter{bottomnumber}{2}
-\setcounter{totalnumber}{4}
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\bottomfraction}{0.85}
-\renewcommand{\textfraction}{0.15}
-\renewcommand{\floatpagefraction}{0.8}
-
-%\RequirePackage[section]{placeins}
-
-%% SECTION TITLES
-\RequirePackage[compact,sf,bf]{titlesec}
-\titleformat{\section}[block]
-  {\fontsize{15}{17}\bfseries\sffamily}
-  {\thesection}
-  {0.4em}{}
-\titleformat{\subsection}[block]
-  {\fontsize{12}{14}\bfseries\sffamily}
-  {\thesubsection}
-  {0.4em}{}
-\titlespacing{\section}{0pt}{*5}{*1}
-\titlespacing{\subsection}{0pt}{*2}{*0.2}
-
-%% TITLE PAGE
-\def\Date{\number\day}
-\def\Month{\ifcase\month\or
- January\or February\or March\or April\or May\or June\or
- July\or August\or September\or October\or November\or December\fi}
-\def\Year{\number\year}
-
-\makeatletter
-\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
-\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
-\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
-\def\nojel{\def\showjel{}}
-\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
-\def\cover{{\sffamily\setcounter{page}{0}
-        \thispagestyle{empty}
-        \vspace*{2cm}
-        \begin{center}
-        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
-                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-                \LARGE\@author\end{onehalfspace}
-        }}
-        \end{center}
-        \vfill
-                \begin{center}\Large
-                \Month~\Year\\[1cm]
-                Working Paper \@wp
-        \end{center}\vspace*{2cm}}}
-\def\pageone{{\sffamily\setstretch{1}%
-        \thispagestyle{empty}%
-        \vbox to \textheight{%
-        \raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
-        \vspace{2cm}\par
-        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
-        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
-        \hspace{1cm}\showjel\vss}}}
-\def\blindtitle{{\sffamily
-     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-        }}
-\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}
-
-\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
-\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
-\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
-\let\maketitle\titlepage
-\makeatother
-
-%% SPACING
-\RequirePackage{setspace}
-\spacing{1.5}
-
-%% LINE AND PAGE BREAKING
-\sloppy
-\clubpenalty = 10000
-\widowpenalty = 10000
-\brokenpenalty = 10000
-\RequirePackage{microtype}
-
-%% PARAGRAPH BREAKS
-\setlength{\parskip}{1.4ex}
-\setlength{\parindent}{0em}
-
-%% HYPERLINKS
-\RequirePackage{xcolor} % Needed for links
-\definecolor{darkblue}{rgb}{0,0,.6}
-\RequirePackage{url}
-
-\makeatletter
-\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
-\makeatother
-\hypersetup{
-     citecolor=0 0 0,
-     breaklinks=true,
-     bookmarksopen=true,
-     bookmarksnumbered=true,
-     linkcolor=darkblue,
-     urlcolor=blue,
-     citecolor=darkblue,
-     colorlinks=true}
-
-%% KEYWORDS
-\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}
-
-%% ABSTRACT
-\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
-\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
-  {\end{minipage}}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage[showonlyrefs]{mathtools}
-\usepackage[no-weekday]{eukdate}
-
-%% BIBLIOGRAPHY
-
-\makeatletter
-\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
-\makeatother
-\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}
-
-\DeclareFieldFormat{url}{\texttt{\url{#1}}}
-\DeclareFieldFormat[article]{pages}{#1}
-\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
-\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
-\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
-\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
-\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
-\DeclareFieldFormat[inproceedings]{title}{#1}
-\DeclareFieldFormat{shorthandwidth}{#1}
-% No dot before number of articles
-\usepackage{xpatch}
-\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
-% Remove In: for an article.
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
-
-\makeatletter
-\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
-\makeatother
-\renewcommand*{\finalnamedelim}{%
-  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
-  \addspace\&\space}
-
-\wp{1}
-\jel{C10,C14,C22}
-
-\RequirePackage[absolute,overlay]{textpos}
-\setlength{\TPHorizModule}{1cm}
-\setlength{\TPVertModule}{1cm}
-\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}
-
-\author{Priyanga Dilini~Talagala, Thiyanga~Talagala}
-\addresses{\textbf{Priyanga Dilini Talagala}\newline
-Department of Compuational Mathematics, University of Moratuwa, Sri Lanka
-\newline{Email: \href{mailto:priyangad@uom.lk}{\nolinkurl{priyangad@uom.lk}}}\newline Corresponding author\\[1cm]
-\textbf{Thiyanga Talagala}\newline
-Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
-\newline{Email: \href{mailto:tstalagala@gmail.com}{\nolinkurl{tstalagala@gmail.com}}}\\[1cm]
-}
-
-\date{\sf\Date~\Month~\Year}
-\makeatletter
- \lfoot{\sf Talagala, Talagala: \@date}
-\makeatother
-
-%% Any special functions or other packages can be loaded here.
-
-\begin{document}
-\maketitle
-\begin{abstract}
-A brief summary of our work
-\end{abstract}
-\begin{keywords}
-rmarkdown, templates
-\end{keywords}
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+# Introduction
 
 The ongoing pandemic of COVID-19 has now become the most important priority of governments and  media of many countries all around the world.
-Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International (PHEIC) Concern on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}.
-As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}.
+Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International (PHEIC) Concern on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline].
+As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].
 In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment  to minimize the spread of the virus.
-In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.
+In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as  travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus [@chang2020modelling].
 
@@
 In response to this increasing trend of COVID 19 cases, lockdowns and curfews were imposed several times across the country with the aim of reducing the spread of the virus.
-As of 29 September 2021, 515524 cases of COVID-19 have been reported in Sri Lanka, resulting with 12786 deaths \autocite{HealthSL2020,dong2020interactive}.
+As of 29 September 2021, 515524 cases of COVID-19 have been reported in Sri Lanka, resulting with 12786 deaths [@HealthSL2020; @dong2020interactive].
 
-\begin{figure}[h]
+```{r covidImpactWorld, fig.cap = "Global tracking of COVID-19 caused school closures and re-openings", fig.height=4}
+```
 
-{\centering \includegraphics[width=1\textwidth]{figure/covidImpactWorld-1}
-
-}
-
-\caption{Global tracking of COVID-19 caused school closures and re-openings}\label{fig:covidImpactWorld}
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\textwidth]{figure/covidImpactContinent-1}
-
-}
-
-\caption{Region wise tracking of COVID-19 caused school closures and re-openings}\label{fig:covidImpactContinent}
-\end{figure}
+```{r covidImpactContinent, fig.cap = "Region wise tracking of COVID-19 caused school closures and re-openings", fig.height=8}
+```
 
@@
 Many sectors have been affected ever since the outbreak of COVID 19, worldwide.
-Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}.
+Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education].
 Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020.
@@
 Africa and Oceania are in better positions in comparison to other regions with their increasing trend of fully open schools.
-In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}.
+In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].
 
@@
 Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}.
+Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic [@drane2020impact; @daniel2020education; @unescoadverse2020].
 
-Eventhough distance education has a long history that goes back to almost two centuries \autocite{spector2014handbook} with significant modifications,
-alterations, and addition in the process of delivery and communication \autocite{moore2011learning}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
+Eventhough distance education has a long history that goes back to almost two centuries [@spector2014handbook] with significant  modifications,
+alterations, and addition in the process of delivery and communication  [@moore2011learning],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
 Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to  unexpected  massive shutdown worldwide.
@@
 
-\begin{figure}[h]
-
-{\centering \includegraphics[width=1\textwidth]{figure/distanceLearningWorldAnalysis-1}
-
-}
-
-\caption{Google trends footprint analysis}\label{fig:distanceLearningWorldAnalysis}
-\end{figure}
+```{r distanceLearningWorldAnalysis, fig.cap= "Google trends footprint analysis", fig.height=6}
+```
 
@@
 Limited time, and lack of knowledge about the best online tools and techniques available are other major challenges teachers encounter during the pandemic.
-In many contexts, school closures lead to furloughs or separations for teachers.\\
+In many contexts, school closures lead to furloughs or separations for teachers.
 Parents unprepared for distance and home schooling: When schools close, parents are often asked to facilitate the learning of children at home and can struggle to perform this task.
@@
 The design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
-It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves \autocite{moore2011learning}.
+It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves [@moore2011learning].
 
-As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal \& Wilson, 2010; Volery \& Lord, 2000) \autocite{moore2011learning}.
+As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal & Wilson, 2010; Volery & Lord, 2000) [@moore2011learning].
 There are many different distance learning solutions.
@@
 The high popularity or increasing online attention towards a product can indicate its prominence in the market, its usability and its impact.
-Further, developers also tend to improve the their products and serivices in response to increasing online search \autocite{willis2020using}
+Further,  developers also tend to improve the their products and serivices in response to increasing online search [@willis2020using]
 
@@
 
-As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, \& Watts, 2010), and politics (Reilly, Richey, \& Taylor, 2012).
-We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study \autocite{vaughan2014web}.
-However their (\textcite{vaughan2014web}) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
+As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, & Watts, 2010), and politics (Reilly, Richey, & Taylor, 2012).
+We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study [@vaughan2014web].
+However their (@vaughan2014web) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
 
@@
 
-\hypertarget{background}{%
-\subsection{Background}\label{background}}
+## Background
 
-\hypertarget{benefits-of-competition}{%
-\subsubsection{Benefits of Competition}\label{benefits-of-competition}}
+### Benefits of Competition
+As Uber was queried more on Google, he was able to find a decrease in complaints of broken credit card machines as the taxis improved their service in response to the increased.
+Complaints may provide a proxy of the quality of traditional taxis in the city and thus provide a means to investigate the service quality differences in yellow taxis since the arrival of Uber [@willis2020using].
 
-As Uber was queried more on Google, he was able to find a decrease in complaints of broken credit card machines as the taxis improved their service in response to the increased.
-Complaints may provide a proxy of the quality of traditional taxis in the city and thus provide a means to investigate the service quality differences in yellow taxis since the arrival of Uber \autocite{willis2020using}.
-
-In total, our approach to employ novel data from a news media outlet to approximate the public interest and Uber's popularity enabled us to illustrate how Uber disrupted the New York City well established taxi market.
+In total, our approach to employ novel data from a news media outlet to approximate the public interest and Uber’s popularity enabled us to illustrate how Uber disrupted the New York City well established taxi market.
 
 Both of these behavioural mechanisms can potentially explain the increase in the number of complaints about taxi service.
-\#\# Analysis
+## Analysis
 
-\url{https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/}
+https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/
 
 Google has analyzed the billions of search requests it processes every day, and identified the terms that have had the highest spike this year compared to 2019.
-Unsurprisingly, ``coronavirus'' topped the overall list.
+Unsurprisingly, “coronavirus” topped the overall list.
 
@@
 Google Classroom was one of several online resources that provided education to millions of children affected by school closures.
-Searches for ``how to foster a dog'' reached an all-time high as social distancing triggered a yearning for the companionship of a pet.
+Searches for “how to foster a dog” reached an all-time high as social distancing triggered a yearning for the companionship of a pet.
 
-\url{https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/}
+https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/
 
-Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception as- sociated with health-threatening events such as an Ebola outbreak {[}5, 37{]}.
-Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease \autocite{alicino2015assessing}.
+Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception as- sociated with health-threatening events such as an Ebola outbreak [5, 37].
+Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease [@alicino2015assessing].
 
 The Internet represents a great opportunity for public health agencies to disseminate healthcare-related infor- mation quickly, effectively and cheaply, provided that re- liable news are shared with as much of the population as possible, both in developed and developing countries, where the diffusion of new technologies among the population is rapidly increasing.
-Google Trends -- a pub- licly available data source -- could be used as a proxy of the proper diffusion of strategies based on health educa- tion messages, allowing to fill the translation gap be- tween best evidence (what some experts know) and practice (what most people do or believe) {[}38{]}.
+Google Trends – a pub- licly available data source – could be used as a proxy of the proper diffusion of strategies based on health educa- tion messages, allowing to fill the translation gap be- tween best evidence (what some experts know) and practice (what most people do or believe) [38].
 More- over, it could be exploited to determine the level of up- take of information about a public health event.
-Even during the 2014 Ebola outbreak, some stakeholders lev- eraged new technologies to plan proper communication strategies and address public concern {[}39, 40 \autocite{alicino2015assessing}.
+Even during the 2014 Ebola outbreak, some stakeholders lev- eraged new technologies to plan proper communication strategies and address public concern [39, 40 [@alicino2015assessing].
 
-Hara and Kling (2000) argue that more research about online courses from the students' perspective is needed.
+Hara and Kling (2000) argue that more research about online courses from the students’ perspective is needed.
 In a case study of graduate students enrolled in an online course, they found that many aspects of the course frustrated students, and that the instructor remained unaware of the continuing level of frustration.
 They suggest that many researchers bring an optimistic, romantic view of technology that may dampen their ability to look at hard questions and apply rigorous research methods.
-Much of the research, they maintain, has been advocacy and theorizing about future possibilities \autocite{wallace2003online}.
-
-\printbibliography
-
-\end{document}
+Much of the research, they maintain, has been advocacy and theorizing about future possibilities [@wallace2003online].

### 2021-11-15 src/main.R

@@ -1,0 +1,2132 @@
+### Supplemental Materials (Data and scripts)
+###  Distance Learning Solutions with COVID-19 Pandemic
+
+## ---- load
+library(tidyverse)
+library (gtrendsR)
+library(RColorBrewer)
+library(ggpubr)
+library(viridis)
+library(fpp3)
+library(countrycode)
+library(coronavirus)
+library(patchwork)
+#set the time window
+# Consider 2020 
+time <- ("2019-12-01 2021-08-15")
+
+## ---- covidImpactWorld
+## COVID-19 Impact on Education (global data)
+## Data source: https://en.unesco.org/covid19/educationresponse
+## Data download: 16-08-2021
+## Write (download from the above link)
+world_impact <- read.csv(
+  here::here("data", "covid_impact_education_full.csv"))
+wimpact <- world_impact %>%
+  mutate(Date = as.Date(Date, "%d/%m/%Y"),
+         Status = as.factor(Status)) 
+
+wimpact <- wimpact %>% select(-ISO) %>%
+  group_by(Date, Status) %>%
+  summarize(Count = n())%>%
+  as_tsibble(index = Date, key = Status) 
+
+p <- wimpact %>% autoplot(Count) +
+  geom_line(size=1) +
+  xlab ("Time") +
+  # viridis::scale_colour_viridis(discrete = TRUE)+
+  scale_color_brewer(palette="Dark2")+
+  ylab("Number of Countries")+
+  theme_bw() + 
+  theme( text = element_text(size = 14), legend.position = "bottom")+
+  guides(colour = guide_legend(nrow = 2, title="Status"))
+
+print(p)
+
+
+## ---- covidImpactContinent
+## COVID-19 Impact on Education - continent wise 
+## Data source: https://en.unesco.org/covid19/educationresponse
+## Data download: 16-08-2021
+world_impact <- read.csv(
+  here::here("data", "covid_impact_education_full.csv"))
+wimpact <- world_impact %>%
+  mutate(Date = as.Date(Date, "%d/%m/%Y"),
+         Status = as.factor(Status)) 
+
+wimpact$Continent <- countrycode(sourcevar = wimpact[, "ISO"],
+                                 origin = "iso3c",
+                                 destination = "continent")
+
[…]

### 2021-11-15 Paper_online_Learning.tex
update paper

@@
 Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020.
-With the start of the pandemic a near total closure of schools were observed all across the globe.
-Africa and Oceania are in better positions in comparison to other regions with their increasing trend of fully open schools.
+With the start of the pandemic a near total closure of schools were observed all around the world.
 In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}.
-
-The speed of these closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders.
-Sudden unexpected movement from classrooms to homes at large scale made students, teachers and parents vulnerable.
-The parents with limited education and resources; and students who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services.
-Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}.
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactWorld}).
 
@@
 alterations, and addition in the process of delivery and communication \autocite{moore2011learning}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
-Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to unexpected massive shutdown worldwide.
-Lack of experience, lack of awareness, lack of readiness for distance education put various stakeholders in the education system in panic, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic Figures \ref{fig:distanceLearningWorldAnalysis}.
-The search spikes coincide with increasing Covid-19 counts and related google trend searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.
+Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
+Surprisingly, the search spikes of distance education related searches coincide with increasing Covid-19 counts and related internet searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.
 
@@
 
+The speed of these closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders.
+The parents with limited education and resources; and students who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
+Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services.
+Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}.
+How to preparing course materials, How to interact with students, how to assess and evaluate student work, how to provide feedback are some major concerns they had when preparing for their first online teaching experience.
+Limited time, and lack of knowledge about the best online tools and techniques available were some of the major challenges teachers encountered during the online course preparation.
+This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home.
+
+In response to this grey situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
+there are many different distance learning solutions.
+The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content \autocite{moore2011learning}.
+However, getting familiar with all these available distance solutions is not feasible or practical due to limited time availability and need of urgent responses.
+
+The popularity of a product greatly influences consumer purchasing decisions can indicate its prominence in the market, its usability and its impact \autocite{ahn2006utilizing}.
+Further, according to \textcite{willis2020using}, developers also tend to improve the their products and services in response to increasing online search.
+Therefore, the Internet represents a great opportunity to learn about public attention towards a product or service.
+It's also a way to narrow down thier search space and thereby identify suitable tools and techniques to create an appropriate learning environment.
+
+Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior \autocite{nuti2014use}.
 In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions.
@@
 (3) Which solutions have a wide attention and public interest both at the national and global level?
+We primarily analyze quantitative digital footprint data on the Internet from December 2019 to August 2021.The google trend footprint will allow the teachers to narrow down thier search space and select prominent distance education solutions for their teaching purposes.
 
-Confusion and stress for teachers: When schools close, especially unexpectedly and for unknown duration, teachers are often unsure of their obligations and how to maintain connections with students to support learning.
-Transitions to distance learning platforms tend to be messy and frustrating, even in the best circumstances.
-Limited time, and lack of knowledge about the best online tools and techniques available are other major challenges teachers encounter during the pandemic.
-In many contexts, school closures lead to furloughs or separations for teachers.\\
-Parents unprepared for distance and home schooling: When schools close, parents are often asked to facilitate the learning of children at home and can struggle to perform this task.
-This is especially true for parents with limited education and resources.
-Challenges creating, maintaining, and improving distance learning: Demand for distance learning skyrockets when schools close and often overwhelms existing portals to remote education.
-Moving learning from classrooms to homes at scale and in a hurry presents enormous challenges, both human and technical.
-
-The design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
-It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves \autocite{moore2011learning}.
-
-As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal \& Wilson, 2010; Volery \& Lord, 2000) \autocite{moore2011learning}.
-There are many different distance learning solutions.
-Identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
-The high popularity or increasing online attention towards a product can indicate its prominence in the market, its usability and its impact.
-Further, developers also tend to improve the their products and serivices in response to increasing online search \autocite{willis2020using}
-
-A big question for faculty members designing new online courses concerns their role as a teacher: what do I do when I teach online?
-Designing the course materials, interacting with students and giving them feedback, and assessing student work are obvious tasks for the online teacher.
-What is not
-obvious, however, is how those tasks are carried out.
-
-As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, \& Watts, 2010), and politics (Reilly, Richey, \& Taylor, 2012).
-We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study \autocite{vaughan2014web}.
-However their (\textcite{vaughan2014web}) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
-
-We primarily analyze quantitative digital footprint data on the Internet from January 2020 to 11.03.2020, including their representativeness.The google trend footprint will allow the teachers to narrow down thier search space and select prominent distance education solutions for their teaching purposes.
-
-The remainder of this paper is organised as follows.
-Section 2 presents the related work to lay the foundation for the stray algorithm.
-Section 3 describes the limitations of the HDoutliers algorithm that hinder its performance.
-Section 4 presents the improved algorithm, stray, that addresses the limitations of the HDoutliers algorithm.
-Section 5 presents a comprehensive evaluation, illustrating the key features of the stray algorithm.
-Section 6 includes an application of stray algorithm related to pedestrian behaviour in the city of Melbourne, Australia.
-Section 7 concludes the article and presents future research directions.
+The remainder of this paper is organized as follows.
+Section 2 presents the related work to lay the foundation for the Google footprint analysis.
+Section 3 presents the methodology followed in the study.
+Section 4 includes Google footprint analysis.
+Section 5 concludes the article and presents future research directions.
 
@@
 
+As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, \& Watts, 2010), and politics (Reilly, Richey, \& Taylor, 2012).
+We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study \autocite{vaughan2014web}.
+However their (\textcite{vaughan2014web}) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
+
+\hypertarget{analysis}{%
+\subsection{Analysis}\label{analysis}}
+
+Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to unexpected massive shutdown worldwide.
+
 \printbibliography

### 2021-11-15 main.Rmd
update paper

@@
 
-```{r covidImpactContinent, fig.cap = "Region wise tracking of COVID-19 caused school closures and re-openings", fig.height=8}
+```{r covidImpactContinent, fig.cap = "Region wise tracking of COVID-19 caused school closures and re-openings", fig.height=6}
 ```
@@
 Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020.
-With the start of the pandemic a near total closure of schools were observed all across the globe.
-Africa and Oceania are in better positions in comparison to other regions with their increasing trend of fully open schools.
+With the start of the pandemic a near total closure of schools were observed all around the world.
 In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactWorld}).
+
+Eventhough distance education has a long history that goes back to almost two centuries [@spector2014handbook] with significant  modifications,
+alterations, and addition in the process of delivery and communication  [@moore2011learning],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
+Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
+Surprisingly, the search spikes of distance education related searches coincide with increasing Covid-19 counts and related internet searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.
+
+```{r distanceLearningWorldAnalysis, fig.cap= "Google trends footprint analysis", fig.height= 8}
+```
 
 The speed of these closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders.
-Sudden unexpected movement from classrooms to homes at large scale made  students, teachers and parents vulnerable.
 The parents  with limited education and  resources;  and  students  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
@@
 Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic [@drane2020impact; @daniel2020education; @unescoadverse2020].
+How to preparing course materials, How to interact with students, how to  assess and evaluate student work, how to provide feedback are some major concerns they had when preparing for their first  online  teaching experience.
+Limited time, and lack of knowledge about the best online tools and techniques available were some of the major challenges teachers encountered during the online course preparation.
+This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home.
 
-Eventhough distance education has a long history that goes back to almost two centuries [@spector2014handbook] with significant  modifications,
-alterations, and addition in the process of delivery and communication  [@moore2011learning],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
-Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to  unexpected  massive shutdown worldwide.
-Lack of experience, lack of awareness, lack of readiness for distance education put various stakeholders in the education system in panic, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic Figures \ref{fig:distanceLearningWorldAnalysis}.
-The search spikes coincide with increasing Covid-19 counts and related google trend searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.
+In response to this grey situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
+there are many different distance learning solutions.
+The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content [@moore2011learning].
+However, getting familiar with all these available distance solutions  is  not feasible or practical due to limited time availability and need of urgent responses.
 
-```{r distanceLearningWorldAnalysis, fig.cap= "Google trends footprint analysis", fig.height=6}
-```
+The popularity of a product greatly influences consumer purchasing decisions can indicate its prominence in the market, its usability and its impact  [@ahn2006utilizing].
+Further, according to @willis2020using, developers also tend to improve the their products and services in response to increasing online search.
+Therefore, the Internet represents a great opportunity to learn about public attention towards a product or service.
+It's also a way to narrow down thier search space and  thereby identify suitable tools and techniques to create an appropriate learning environment.
 
+Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior [@nuti2014use].
 In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions.
@@
 (3) Which solutions have a wide attention and public interest both at the national and global level?
+We primarily analyze quantitative digital footprint data on the Internet from December 2019 to August 2021.The google trend footprint will allow the teachers to narrow down thier search space and select  prominent distance education solutions for their teaching purposes.
 
-Confusion and stress for teachers: When schools close, especially unexpectedly and for unknown duration, teachers are often unsure of their obligations and how to maintain connections with students to support learning.
-Transitions to distance learning platforms tend to be messy and frustrating, even in the best circumstances.
-Limited time, and lack of knowledge about the best online tools and techniques available are other major challenges teachers encounter during the pandemic.
-In many contexts, school closures lead to furloughs or separations for teachers.
-Parents unprepared for distance and home schooling: When schools close, parents are often asked to facilitate the learning of children at home and can struggle to perform this task.
-This is especially true for parents with limited education and resources.
-Challenges creating, maintaining, and improving distance learning: Demand for distance learning skyrockets when schools close and often overwhelms existing portals to remote education.
-Moving learning from classrooms to homes at scale and in a hurry presents enormous challenges, both human and technical.
-
-The design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
-It is important to know how the learning environment is used, and the influences of the tools and techniques that distinguish the differences in learning outcomes as the technology evolves [@moore2011learning].
-
-As learning technology and its associated fields continue to evolve, practitioners and researchers have yet to agree on common definitions and terminologies (Lowenthal & Wilson, 2010; Volery & Lord, 2000) [@moore2011learning].
-There are many different distance learning solutions.
-Identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
-The high popularity or increasing online attention towards a product can indicate its prominence in the market, its usability and its impact.
-Further,  developers also tend to improve the their products and serivices in response to increasing online search [@willis2020using]
-
-A big question for faculty members designing new online courses concerns their role as a teacher: what do I do when I teach online?
-Designing the course materials, interacting with students and giving them feedback, and assessing student work are obvious tasks for the online teacher.
-What is not
-obvious, however, is how those tasks are carried out.
-
-As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, & Watts, 2010), and politics (Reilly, Richey, & Taylor, 2012).
-We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study [@vaughan2014web].
-However their (@vaughan2014web) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
-
-We primarily analyze quantitative digital footprint data on the Internet from January 2020 to 11.03.2020, including their representativeness.The google trend footprint will allow the teachers to narrow down thier search space and select  prominent distance education solutions for their teaching purposes.
-
-The remainder of this paper is organised as follows.
-Section 2 presents the related work to lay the foundation for the stray algorithm.
-Section 3 describes the limitations of the HDoutliers algorithm that hinder its performance.
-Section 4 presents the improved algorithm, stray, that addresses the limitations of the HDoutliers algorithm.
-Section 5 presents a comprehensive evaluation, illustrating the key features of the stray algorithm.
-Section 6 includes an application of stray algorithm related to pedestrian behaviour in the city of Melbourne, Australia.
-Section 7 concludes the article and presents future research directions.
+The remainder of this paper is organized as follows.
+Section 2 presents the related work to lay the foundation for the Google footprint analysis.
+Section 3 presents the methodology followed in the study.
+Section 4 includes Google footprint analysis.
+Section 5 concludes the article and presents future research directions.
 
@@
 Much of the research, they maintain, has been advocacy and theorizing about future possibilities [@wallace2003online].
+
+As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, & Watts, 2010), and politics (Reilly, Richey, & Taylor, 2012).
+We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study [@vaughan2014web].
+However their (@vaughan2014web) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
+
+## Analysis
+
+Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to  unexpected  massive shutdown worldwide.

### 2021-12-16 Paper_online_Learning.Rmd
update online learning

@@
 - familyname: Talagala
-  othernames: Thiyanga
+  othernames: Thiyanga S.
   address: Department of Statistics, University of Sri Jayewardenepura, Sri Lanka

### 2021-12-16 Paper_online_Learning.tex
update online learning

@@
 
-\author{Priyanga Dilini~Talagala, Thiyanga~Talagala}
+\author{Priyanga Dilini~Talagala, Thiyanga S.~Talagala}
 \addresses{\textbf{Priyanga Dilini Talagala}\newline
@@
 \newline{Email: \href{mailto:priyangad@uom.lk}{\nolinkurl{priyangad@uom.lk}}}\newline Corresponding author\\[1cm]
-\textbf{Thiyanga Talagala}\newline
+\textbf{Thiyanga S. Talagala}\newline
 Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
@@
 
-The ongoing pandemic of COVID-19 has now become the most important priority of governments and media of many countries all around the world.
-Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International (PHEIC) Concern on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}.
+The ongoing pandemic of COVID-19 is still one of the most important priorities of governments and media of many countries all around the world.
+Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}.
 As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}.
@@
 
-The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with a Chinese national who came to Sri Lanka as a tourist.
+The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with a person who came to Sri Lanka as a tourist.
 The first locally transmitted COVID-19 case in Sri Lanka was then reported on 11 March 2020 followed by an increasing trend of the COVID-19 confirmed cases.
@@
 In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactWorld}).
-
-Eventhough distance education has a long history that goes back to almost two centuries \autocite{spector2014handbook} with significant modifications,
-alterations, and addition in the process of delivery and communication \autocite{moore2011learning}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
-Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
-Surprisingly, the search spikes of distance education related searches coincide with increasing Covid-19 counts and related internet searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactContinent}).
 
@@
 
-\caption{Google trends footprint analysis}\label{fig:distanceLearningWorldAnalysis}
+\caption{Visualization of data from Goolge search trends.
+(a) COVID-19 cases worldwide (b) Google search trends of COVID-19 related terms (c) Google search trends of distance education related terms.}\label{fig:distanceLearningWorldAnalysis}
 \end{figure}
 
-The speed of these closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders.
-The parents with limited education and resources; and students who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
+Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
+alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
+Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
+Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.
+
+The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders: teachers, students and parents.
+Teachers were not fully aware of their obligations and how to maintain connections with students to support thier learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic.
+Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
 Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services.
-Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic \autocite{drane2020impact,daniel2020education,unescoadverse2020}.
-How to preparing course materials, How to interact with students, how to assess and evaluate student work, how to provide feedback are some major concerns they had when preparing for their first online teaching experience.
-Limited time, and lack of knowledge about the best online tools and techniques available were some of the major challenges teachers encountered during the online course preparation.
+\autocite{drane2020impact,daniel2020education,unescoadverse2020}.
 This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home.
+Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown.
+This situation left internet as the only medium to support all these educational processes.
 
-In response to this grey situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
-there are many different distance learning solutions.
+In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure.
+There are many different distance learning solutions.
 The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content \autocite{moore2011learning}.
@@
 
-The popularity of a product greatly influences consumer purchasing decisions can indicate its prominence in the market, its usability and its impact \autocite{ahn2006utilizing}.
-Further, according to \textcite{willis2020using}, developers also tend to improve the their products and services in response to increasing online search.
-Therefore, the Internet represents a great opportunity to learn about public attention towards a product or service.
-It's also a way to narrow down thier search space and thereby identify suitable tools and techniques to create an appropriate learning environment.
+According to \textcite{ahn2006utilizing}, the popularity of a product greatly influences consumer purchasing decisions.
+It can also indicate the prominence of the product in the market, its usability and its impact.
+Further, according to \textcite{willis2020using}, developers also tend to improve their products and services in response to increasing online search.
+Therefore, the internet represents a great opportunity to learn about public attention towards a product or service.
+It is also a way to narrow down the search space and thereby identify most suitable product and services that meets customer needs.
 
 Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior \autocite{nuti2014use}.
-In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions.
+In this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
 In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education?
-(2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic?
-(3) Which solutions have a wide attention and public interest both at the national and global level?
-We primarily analyze quantitative digital footprint data on the Internet from December 2019 to August 2021.The google trend footprint will allow the teachers to narrow down thier search space and select prominent distance education solutions for their teaching purposes.
+(2) What solutions are in place to meet the need of distance education in terms during COVID 19 pandemic?
+(3) Which distance learning solutions have a wide attention and public interest both at the national and global level?
+We primarily analyze quantitative digital footprint data on the internet from December 2019 to August 2021.
+The resulted google trend footprint provides a fast first step to identify the most popular distance education tools available for different education purposes.
+It also allows the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching.
+
+Quality education is essential to sustainable development and it is a key United Nations (UN) sustainable development goal.
+Having the right tool is imperative to successful completion of a task at hand.
+The findings of this study provides an initial guidance to select a right tool.
+Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.
 
@@
 
-Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception as- sociated with health-threatening events such as an Ebola outbreak {[}5, 37{]}.
+Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception associated with health-threatening events such as an Ebola outbreak {[}5, 37{]}.
 Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease \autocite{alicino2015assessing}.

### 2021-12-16 main.Rmd
update online learning

@@
 
-The ongoing pandemic of COVID-19 has now become the most important priority of governments and  media of many countries all around the world.
-Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International (PHEIC) Concern on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline].
+The ongoing pandemic of COVID-19 is still one of the most important priorities of governments and  media of many countries all around the world.
+Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline].
 As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].
@@
 
-The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with  a Chinese national who came to Sri Lanka as a tourist.
+The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with  a person who came to Sri Lanka as a tourist.
 The first locally transmitted COVID-19 case in Sri Lanka was then reported on  11 March 2020 followed by an increasing trend of the COVID-19 confirmed cases.
@@
 In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactWorld}).
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactContinent}).
 
-Eventhough distance education has a long history that goes back to almost two centuries [@spector2014handbook] with significant  modifications,
-alterations, and addition in the process of delivery and communication  [@moore2011learning],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
-Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
-Surprisingly, the search spikes of distance education related searches coincide with increasing Covid-19 counts and related internet searches, eventhough distance education has a long history in contrast to Covid-19 pandemic.
-
-```{r distanceLearningWorldAnalysis, fig.cap= "Google trends footprint analysis", fig.height= 8}
+```{r distanceLearningWorldAnalysis, fig.cap= "Visualization of data from Goolge search trends.
+(a) COVID-19 cases worldwide (b) Google search trends of COVID-19 related terms (c) Google search trends of distance education related terms.", fig.height= 8}
 ```
 
-The speed of these closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders.
-The parents  with limited education and  resources;  and  students  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
+Eventhough distance education has a long history that goes back to almost two centuries with significant  modifications,
+alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
+Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
+Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.
+
+The speed of the physical classroom closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders: teachers, students and parents.
+Teachers were not fully aware of their obligations and how to maintain connections with students to support thier learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic.
+Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
 Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services.
-Teachers on the other hand were not fully aware of their obligations and how to maintain connections with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic [@drane2020impact; @daniel2020education; @unescoadverse2020].
-How to preparing course materials, How to interact with students, how to  assess and evaluate student work, how to provide feedback are some major concerns they had when preparing for their first  online  teaching experience.
-Limited time, and lack of knowledge about the best online tools and techniques available were some of the major challenges teachers encountered during the online course preparation.
+[@drane2020impact; @daniel2020education; @unescoadverse2020].
 This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home.
+Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown.
+This situation left internet as the only medium to support all these educational processes.
 
-In response to this grey situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
-there are many different distance learning solutions.
+In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure.
+There are many different distance learning solutions.
 The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content [@moore2011learning].
@@
 
-The popularity of a product greatly influences consumer purchasing decisions can indicate its prominence in the market, its usability and its impact  [@ahn2006utilizing].
-Further, according to @willis2020using, developers also tend to improve the their products and services in response to increasing online search.
-Therefore, the Internet represents a great opportunity to learn about public attention towards a product or service.
-It's also a way to narrow down thier search space and  thereby identify suitable tools and techniques to create an appropriate learning environment.
+According to @ahn2006utilizing, the popularity of a product greatly influences consumer purchasing decisions.
+It can also indicate the prominence of the product in the market, its usability and its impact.
+Further, according to @willis2020using, developers also tend to improve their products and services in response to increasing online search.
+Therefore, the internet represents a great opportunity to learn about public attention towards a product or service.
+It is also a way to narrow down the search space and  thereby identify most suitable product and services that meets customer needs.
 
 Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior [@nuti2014use].
-In this paper we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in distance education solutions.
+In this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
 In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID 19 pandemic on education?
-(2) What solutions are in place to meet the need of distance education in terms of coverage, up-take, and learning during COVID 19 pandemic?
-(3) Which solutions have a wide attention and public interest both at the national and global level?
-We primarily analyze quantitative digital footprint data on the Internet from December 2019 to August 2021.The google trend footprint will allow the teachers to narrow down thier search space and select  prominent distance education solutions for their teaching purposes.
+(2) What solutions are in place to meet the need of distance education in terms during COVID 19 pandemic?
+(3) Which distance learning solutions have a wide attention and public interest both at the national and global level?
+We primarily analyze quantitative digital footprint data on the internet from December 2019 to August 2021.
+The resulted google trend footprint provides a fast first step to identify the most popular distance education tools available for different education purposes.
+It  also allows the teachers to narrow down the search space and  deepen their exploration on  prominent distance education solutions to support their online teaching.
+
+Quality education is essential to sustainable development and it is a key  United Nations (UN) sustainable development goal.
+Having the right tool is imperative to successful completion of a task at hand.
+The findings of this study provides an initial guidance to select a right tool.
+Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.
 
@@
 
-Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception as- sociated with health-threatening events such as an Ebola outbreak [5, 37].
+Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception associated with health-threatening events such as an Ebola outbreak [5, 37].
 Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease [@alicino2015assessing].

### 2021-12-16 src/main.R

@@ -203,7 +203,9 @@ p1 <- covidcasesW1 %>%  autoplot() +
   scale_color_brewer(palette="Dark2") +
   scale_x_date(limits = as.Date(c('2017-12-03','2021-08-08'))) +
   theme_bw()+
-  theme(legend.position = "bottom", text = element_text(size = 14))
+  theme(legend.position = "bottom", text = element_text(size = 14)) +
+  ggtitle("(a)")+
+  xlab("")
 
 
 # covid vaccine cases
@@ -252,7 +254,9 @@ p3<- dataW %>%  ggplot(aes(x=date, y=hits, colour = keyword))+
   scale_x_date(limits = as.Date(c('2017-12-03','2021-08-08'))) +
   theme_bw()+
   theme(legend.position = "bottom",
-        text = element_text(size = 14))
+        text = element_text(size = 14)) +
+  ggtitle("(b)")+
+  xlab("")
 
 
 ## Covid vacicne search trend
@@ -308,7 +312,8 @@ p5<- dataW%>% ggplot(aes(x=date, y=hits, colour = keyword))+
   theme_bw()+
   theme(legend.position = "bottom",
         text = element_text(size = 14))+
-  guides(colour = guide_legend(nrow = 2, title="keywords"))
+  guides(colour = guide_legend(nrow = 2, title="keywords"))+
+  ggtitle("(c)")
 
 p <- p1/p3/p5
 

### 2021-12-16 Paper_online_Learning.tex
update introduction

@@
 
-The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with a person who came to Sri Lanka as a tourist.
-The first locally transmitted COVID-19 case in Sri Lanka was then reported on 11 March 2020 followed by an increasing trend of the COVID-19 confirmed cases.
-In response to this increasing trend of COVID 19 cases, lockdowns and curfews were imposed several times across the country with the aim of reducing the spread of the virus.
-As of 29 September 2021, 515524 cases of COVID-19 have been reported in Sri Lanka, resulting with 12786 deaths \autocite{HealthSL2020,dong2020interactive}.
-
 \begin{figure}[h]
@@
 With the start of the pandemic a near total closure of schools were observed all around the world.
-In line with that, the Sri Lankan government also ordered to close schools from 12 March 2020 and its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances \autocite{wikieducation}.
 Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactContinent}).
@@
 (2) What solutions are in place to meet the need of distance education in terms during COVID 19 pandemic?
-(3) Which distance learning solutions have a wide attention and public interest both at the national and global level?
+(3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?
 We primarily analyze quantitative digital footprint data on the internet from December 2019 to August 2021.
@@
 
-\hypertarget{benefits-of-competition}{%
-\subsubsection{Benefits of Competition}\label{benefits-of-competition}}
-
-As Uber was queried more on Google, he was able to find a decrease in complaints of broken credit card machines as the taxis improved their service in response to the increased.
-Complaints may provide a proxy of the quality of traditional taxis in the city and thus provide a means to investigate the service quality differences in yellow taxis since the arrival of Uber \autocite{willis2020using}.
-
-In total, our approach to employ novel data from a news media outlet to approximate the public interest and Uber's popularity enabled us to illustrate how Uber disrupted the New York City well established taxi market.
-
-Both of these behavioural mechanisms can potentially explain the increase in the number of complaints about taxi service.
-\#\# Analysis
-
-\url{https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/}
-
-Google has analyzed the billions of search requests it processes every day, and identified the terms that have had the highest spike this year compared to 2019.
-Unsurprisingly, ``coronavirus'' topped the overall list.
-
-The search trends show the impact of the COVID pandemic on so many aspects of life.
-Relatively obscure teleconferencing company Zoom became a household name (and highly profitable business) as employees adapted to working from home, and friends and families turned to video calls instead of meeting in person .
-
-Google Classroom was one of several online resources that provided education to millions of children affected by school closures.
-Searches for ``how to foster a dog'' reached an all-time high as social distancing triggered a yearning for the companionship of a pet.
-
-\url{https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/}
-
-Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception associated with health-threatening events such as an Ebola outbreak {[}5, 37{]}.
-Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease \autocite{alicino2015assessing}.
-
-The Internet represents a great opportunity for public health agencies to disseminate healthcare-related infor- mation quickly, effectively and cheaply, provided that re- liable news are shared with as much of the population as possible, both in developed and developing countries, where the diffusion of new technologies among the population is rapidly increasing.
-Google Trends -- a pub- licly available data source -- could be used as a proxy of the proper diffusion of strategies based on health educa- tion messages, allowing to fill the translation gap be- tween best evidence (what some experts know) and practice (what most people do or believe) {[}38{]}.
-More- over, it could be exploited to determine the level of up- take of information about a public health event.
-Even during the 2014 Ebola outbreak, some stakeholders lev- eraged new technologies to plan proper communication strategies and address public concern {[}39, 40 \autocite{alicino2015assessing}.
-
-Hara and Kling (2000) argue that more research about online courses from the students' perspective is needed.
-In a case study of graduate students enrolled in an online course, they found that many aspects of the course frustrated students, and that the instructor remained unaware of the continuing level of frustration.
-They suggest that many researchers bring an optimistic, romantic view of technology that may dampen their ability to look at hard questions and apply rigorous research methods.
-Much of the research, they maintain, has been advocacy and theorizing about future possibilities \autocite{wallace2003online}.
-
-As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, \& Watts, 2010), and politics (Reilly, Richey, \& Taylor, 2012).
-We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study \autocite{vaughan2014web}.
-However their (\textcite{vaughan2014web}) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
-
 \hypertarget{analysis}{%
@@
 
-Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to unexpected massive shutdown worldwide.
+\hypertarget{discussion-and-conclusion}{%
+\subsection{Discussion and Conclusion}\label{discussion-and-conclusion}}
 

### 2021-12-16 main.Rmd
update introduction

@@
 
+<!--
 The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with  a person who came to Sri Lanka as a tourist.
@@
 As of 29 September 2021, 515524 cases of COVID-19 have been reported in Sri Lanka, resulting with 12786 deaths [@HealthSL2020; @dong2020interactive].
+-->
 
@@
 With the start of the pandemic a near total closure of schools were observed all around the world.
-In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].
 Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactContinent}).
+
+<!--In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].-->
 
@@
 
-### Benefits of Competition
-As Uber was queried more on Google, he was able to find a decrease in complaints of broken credit card machines as the taxis improved their service in response to the increased.
-Complaints may provide a proxy of the quality of traditional taxis in the city and thus provide a means to investigate the service quality differences in yellow taxis since the arrival of Uber [@willis2020using].
-
-In total, our approach to employ novel data from a news media outlet to approximate the public interest and Uber’s popularity enabled us to illustrate how Uber disrupted the New York City well established taxi market.
-
-Both of these behavioural mechanisms can potentially explain the increase in the number of complaints about taxi service.
 ## Analysis
 
-https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/
-
-Google has analyzed the billions of search requests it processes every day, and identified the terms that have had the highest spike this year compared to 2019.
-Unsurprisingly, “coronavirus” topped the overall list.
-
-The search trends show the impact of the COVID pandemic on so many aspects of life.
-Relatively obscure teleconferencing company Zoom became a household name (and highly profitable business) as employees adapted to working from home, and friends and families turned to video calls instead of meeting in person .
-
-Google Classroom was one of several online resources that provided education to millions of children affected by school closures.
-Searches for “how to foster a dog” reached an all-time high as social distancing triggered a yearning for the companionship of a pet.
-
-https://www.weforum.org/agenda/2020/12/google-most-popular-search-2020/
-
-Notably, GT, together with social media, could be use- fully exploited by global and local health agencies, as well as by public health practitioners, as a tool to identify disease-related information needs, social and behavioural barriers to infection control, and misinformation, as well as to better understand sentiment and risk perception associated with health-threatening events such as an Ebola outbreak [5, 37].
-Indeed, analyses of the temporal trend of GT-based query volumes together with their geographic distribution and main topics of searches can furnish a quantifiable and valuable measure of public attention and information needs about a particular communicable disease [@alicino2015assessing].
-
-The Internet represents a great opportunity for public health agencies to disseminate healthcare-related infor- mation quickly, effectively and cheaply, provided that re- liable news are shared with as much of the population as possible, both in developed and developing countries, where the diffusion of new technologies among the population is rapidly increasing.
-Google Trends – a pub- licly available data source – could be used as a proxy of the proper diffusion of strategies based on health educa- tion messages, allowing to fill the translation gap be- tween best evidence (what some experts know) and practice (what most people do or believe) [38].
-More- over, it could be exploited to determine the level of up- take of information about a public health event.
-Even during the 2014 Ebola outbreak, some stakeholders lev- eraged new technologies to plan proper communication strategies and address public concern [39, 40 [@alicino2015assessing].
-
-Hara and Kling (2000) argue that more research about online courses from the students’ perspective is needed.
-In a case study of graduate students enrolled in an online course, they found that many aspects of the course frustrated students, and that the instructor remained unaware of the continuing level of frustration.
-They suggest that many researchers bring an optimistic, romantic view of technology that may dampen their ability to look at hard questions and apply rigorous research methods.
-Much of the research, they maintain, has been advocacy and theorizing about future possibilities [@wallace2003online].
-
-As shown in this literature review, the most common applications of Google search data in social science to date have been in economics, finance, and business, though there have also been studies in other fields, including media and technology (Rech, 2007), the entertainment industry (Goel, Hofman, Lahaie, Pennock, & Watts, 2010), and politics (Reilly, Richey, & Taylor, 2012).
-We did not, however, find any studies using Google Trends data in the higher education sector, which is our area of study [@vaughan2014web].
-However their (@vaughan2014web) fous is on Our study determines whether Google Trends data can be used to estimate or predict academic fame by testing the correlation between search volume data and university ranking data.
-
-## Analysis
-
-Every single step including planning, developing new tools and techniques, conducting awareness programs and workshops about online education, shifting towards online education happens through online due to  unexpected  massive shutdown worldwide.
+## Discussion and Conclusion

### 2021-12-16 Paper_online_Learning.tex
update background

@@
 It also allows the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching.
+Further, identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
 
@@
 
+Google Trends, is a website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, the largest search engine in the world, across various regions, subject domains and languages.
+Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution \autocite{alicino2015assessing}.
+
+The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding} and health care \autocite{nuti2014use,alicino2015assessing,arora2019google}.
+According to the study conducted by \textcite{willis2020using} on taxi preferences found that online quires can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+Several recent studies have also been conducted to investigate the relation between Google search trend patterns and the COVID-19 related concerns \autocite{husnayain2020applications,effenberger2020association}.
+The study conducted by \textcite{husnayain2020applications} focuses on search terms related to the coronavirus, handwashing, and face masks to monitor public restlessness toward COVID-19.
+However, there are only a limited number of research attempts related to education section using Google trend search query data \autocite{vaughan2014web}.
+The study conducted by \textcite{vaughan2014web} have used google trend search queries to predict academic fame by testing the correlation between search volume data and university ranking data.
+\textcite{kansal2021google} have used a text mining approach to identify emerging patterns of words and phrases during COVID-19 pandemic, keeping education prospects as the focus.
+However, their study was limited to a small number of learning platforms.
+However, according to \textcite{moore2011learning} the design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
+Therefore, this study provides a detailed investigation on distance education solutions focusing on distance learning needs.
+
+Eventhough google trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool.
+According to \textcite{nuti2014use} the data retrieval process should be transparent.
+This will increase the trustworthiness of both the results and the generalizability of the findings.
+Furthermore, researchers should clearly document the rationale and data retrieval process to ensure the reproducibility of results \autocite{nuti2014use}.
+
+\hypertarget{methodology}{%
+\subsection{Methodology}\label{methodology}}
+
+We visualise our data and we then employ time series modelling to capture
+
+\hypertarget{impact-of-covid-19-on-education}{%
+\subsubsection{Impact of covid 19 on education}\label{impact-of-covid-19-on-education}}
+
+In response to school closures, UNESCO recommended the use of distance learning programmes and open educational applications and platforms that schools and teachers can use to reach learners remotely and limit the disruption of education.
+
 \hypertarget{analysis}{%

### 2021-12-16 main.Rmd
update background

@@
 It  also allows the teachers to narrow down the search space and  deepen their exploration on  prominent distance education solutions to support their online teaching.
+Further, identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
 
@@
 
+Google Trends, is a website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, the largest search engine in the world, across various regions, subject domains and languages.
+Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution  [@alicino2015assessing].
+
+The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding] and health care [@nuti2014use; @alicino2015assessing; @arora2019google].
+According to the study conducted by @willis2020using on taxi preferences found that online quires can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+Several recent studies have also been conducted  to investigate the relation between Google search trend patterns and the COVID-19 related concerns [@husnayain2020applications; @effenberger2020association].
+The study conducted by @husnayain2020applications focuses on search terms related to the coronavirus, handwashing, and face masks  to monitor public restlessness toward COVID-19.
+However, there are only a limited number of research attempts related to education section  using Google trend search query data [@vaughan2014web].
+The study conducted by @vaughan2014web have used  google trend search queries to predict academic fame by testing the correlation between search volume data and university ranking data.
+@kansal2021google have used a text mining approach to identify emerging patterns of words and phrases during COVID-19 pandemic, keeping education prospects as the focus.
+However, their study was limited to a small number of  learning platforms.
+However, according to @moore2011learning the design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
+Therefore, this study    provides a detailed investigation on distance  education solutions focusing on  distance learning needs.
+
+Eventhough google trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool.
+According to @nuti2014use the data retrieval process should be transparent.
+This will increase the trustworthiness of both the results and the generalizability of the findings.
+Furthermore, researchers should  clearly document the rationale and data retrieval process to ensure the reproducibility of results [@nuti2014use].
+
+## Methodology
+
+We visualise our data and we then employ time series modelling to capture
+
+### Impact of covid 19 on education
+
+In response to school closures, UNESCO recommended the use of distance learning programmes and open educational applications and platforms that schools and teachers can use to reach learners remotely and limit the disruption of education.
+
 ## Analysis

### 2021-12-25 Paper_online_Learning.tex
update methodology

@@
 Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
-alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
-Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
+alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}.
+Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.
 Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.
@@
 Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services.
-\autocite{drane2020impact,daniel2020education,unescoadverse2020}.
+Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services, long-term educational disengagement, digital exclusion, poor technology management, increased psychosocial challenges, decreases in mathematical achievement, a lack of engagement, increased early marriages , psychosocial challenges \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}.
 This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home.
+Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures.
 Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown.
@@
 There are many different distance learning solutions.
+Yet very little is known about the uptake of different solutions or about their effectiveness.
 The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content \autocite{moore2011learning}.
 However, getting familiar with all these available distance solutions is not feasible or practical due to limited time availability and need of urgent responses.
+According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.
 
-According to \textcite{ahn2006utilizing}, the popularity of a product greatly influences consumer purchasing decisions.
-It can also indicate the prominence of the product in the market, its usability and its impact.
+Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting distance education solutions.
+Further, according to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as popularity can indicate the prominence of the product in the market, its usability and its impact.
 Further, according to \textcite{willis2020using}, developers also tend to improve their products and services in response to increasing online search.
@@
 Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior \autocite{nuti2014use}.
-In this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
+As there is still the main struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
 In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education?
@@
 
-Google Trends, is a website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, the largest search engine in the world, across various regions, subject domains and languages.
-Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution \autocite{alicino2015assessing}.
+Google Trends, is a freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages \autocite{carneiro2009google}.
+Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution \autocite{alicino2015assessing,cook2011assessing}.
+According to \textcite{jarynowski2020perception} the public concerns on different matters can also take an epidemic nature starting with the phase of growing interest which is known as ``early adoption'', to the phase of general interest which is known as ``majority'' to eventually lose popularity which is known as ``lagers stage''.
+Figure \ref{fig:distanceLearningWorldAnalysis} also confirms this life cycle explanation.
 
-The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding} and health care \autocite{nuti2014use,alicino2015assessing,arora2019google}.
-According to the study conducted by \textcite{willis2020using} on taxi preferences found that online quires can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding} and health care \autocite{nuti2014use,alicino2015assessing,arora2019google,carneiro2009google,cook2011assessing} as it provides data about social phenomena in a more timely manner than traditional data collection processes \autocite{vaughan2014web}.
+According to the study conducted by \textcite{willis2020using} on taxi preferences found that online queries can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+In November, 2008, Google launched an internet-based surveillance tool, Google Flu Trends (GFT), that uses aggregated Google search data to provide a near real-time support in predicting influenza outbreaks \autocite{cook2011assessing}.
 Several recent studies have also been conducted to investigate the relation between Google search trend patterns and the COVID-19 related concerns \autocite{husnayain2020applications,effenberger2020association}.
@@
 However, according to \textcite{moore2011learning} the design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
-Therefore, this study provides a detailed investigation on distance education solutions focusing on distance learning needs.
+Therefore, this study provides a detailed investigation on distance education solutions focusing on different distance learning needs.
+In contrast to the work in \textcite{kansal2021google}, the work in our paper has a very specific focus: the distance learning solutions for various education purposes.
 
 Eventhough google trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool.
+It may contain inaccuracies \autocite{carneiro2009google}.
 According to \textcite{nuti2014use} the data retrieval process should be transparent.
@@
 Furthermore, researchers should clearly document the rationale and data retrieval process to ensure the reproducibility of results \autocite{nuti2014use}.
+.
 
@@
 
+In this work we use Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
+We use weekly Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
+We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of Covid-19 pandemic.
+Through this we want to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
+We also study the correlation between weekly web searches and the number of weekly new cases of covid-19.
+
+First ``Distance education'' and related terms were searched.
+
 We visualise our data and we then employ time series modelling to capture
 
-\hypertarget{impact-of-covid-19-on-education}{%
-\subsubsection{Impact of covid 19 on education}\label{impact-of-covid-19-on-education}}
+\hypertarget{keyword-selection}{%
+\subsection{Keyword selection}\label{keyword-selection}}
 
-In response to school closures, UNESCO recommended the use of distance learning programmes and open educational applications and platforms that schools and teachers can use to reach learners remotely and limit the disruption of education.
+There are different types of interactions that facilitate learners to construct knowledge in distance education environments.
+These are learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction \autocite{wallace2003online}.
+
+Google Trends determines categories based on search patterns.
+According to \textcite{vaughan2014web} more relevant accurate data can be obtained by limiting the search in Google Trends to a specific category as it helps to reduce noise in the data.
+In this work our focus was given to education category.
+Google Trends does not provide the absolute search volume of a given search term.
+Instead it provides a relative search volume of a particular term adjusted according to the total searches of the geography and time range it represents.
+The resulted series scale from 0 to 100, where each data point in the series represents the search interest relative to the highest point on the series for the selected region and time.
+\autocite{alicino2015assessing,vaughan2014web}.
+A value of 100 represents the peak popularity of the term.
+
+The term distance education is known by a variety of names. The terms such as distance learning, e-Learning, online learning environments, home study, independent study, external
+study, correspondence education, off campus study, open learning and open education are used commonly and interchangeably \autocite{moore2011learning}.
+It often describes the effort of providing access to learning for those who are geographically distant.
+The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
+Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places and/or varying forms of instructional materials \autocite{moore2011learning}.
+It is important to decide what term to search for in Google Trends as different terms have different search volumes.
+In the past literature several strategies have been used to select suitable search terms for a given study.
+Some studies have selected search terms based on intuition and some through brainstorming processes \autocite{vaughan2014web}.
+In this study our choice of search term took into account the list of distance learning solutions published by UNESCO \autocite{unesco2020DLsolutions}.
+Eventhough, the solutions they had recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base and evidence of impact \autocite{unesco2020DLsolutions}.
+Different solutions are categorized based on distance learning needs.
+For each keyword, the worldwide search queries were performed with a keyword being used as the ``search term''.
+That allowed us to search the exact string of text typed by the user.
+However, according to \textcite{vaughan2014web}, short forms or acronyms had higher search volumes than the corresponding complete names, in general.
+However we did not use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
+
+In google trend It is possible to search for up to five queries each time {[}21{]}.
+
+Google Trends does not report the absolute search volume of a single term but does report the relative search volumes for up to five multiple terms \autocite{vaughan2014web}.
+
+We entered up to five universities at a time and recorded their relative ranking scores, we then entered a different group of universities, some of which overlapped with the previous group, and we recorded the relative ranking scores again.
+Through this repeated process of relative comparison, we were able to obtain relative ranking scores for all uni- versities in the study.
+(Vaughan \& Romero-Frias 2014)
 
@@
 
-\hypertarget{discussion-and-conclusion}{%
-\subsection{Discussion and Conclusion}\label{discussion-and-conclusion}}
+bola the most searched item on Yahoo engine and the fifth most searched term on Google in 2014 {[}2, 3{]}.
+
+Duifferent attesnion for different etrms
+The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
+The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane \& Galyen 2011).
+
+\hypertarget{exams}{%
+\subsection{exams}\label{exams}}
+
+Strategies to postpone, skip or adminsiter examinations at a distance raise serious concerns about fairness, especialy when access to learning becomes variable.
+Disruptions to assessments results in stress for students and their families and can trigger disengagement.
+\autocite{unesco2020covid,unescoadverse2020}
+\#\# Discussion and Conclusion
 

### 2021-12-25 main.Rmd
update methodology

@@
 Eventhough distance education has a long history that goes back to almost two centuries with significant  modifications,
-alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously.
-Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}).
+alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously [@richmond2020critical].
+Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
 Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.
@@
 Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services.
-[@drane2020impact; @daniel2020education; @unescoadverse2020].
+Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services, long-term educational disengagement, digital exclusion, poor technology management, increased psychosocial challenges, decreases in mathematical achievement, a lack of engagement, increased early marriages ,  psychosocial challenges [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher].
 This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home.
+Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures.
 Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown.
@@
 There are many different distance learning solutions.
+Yet very little is known about the uptake of different solutions or about their effectiveness.
 The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content [@moore2011learning].
 However, getting familiar with all these available distance solutions  is  not feasible or practical due to limited time availability and need of urgent responses.
+According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught.
 
-According to @ahn2006utilizing, the popularity of a product greatly influences consumer purchasing decisions.
-It can also indicate the prominence of the product in the market, its usability and its impact.
+Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting distance education solutions.
+Further, according to @ahn2006utilizing, the popularity of a product can also greatly influence consumer purchasing decisions as popularity can  indicate the prominence of the product in the market, its usability and its impact.
 Further, according to @willis2020using, developers also tend to improve their products and services in response to increasing online search.
@@
 Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior [@nuti2014use].
-In this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
+As there is still the main struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
 In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID 19 pandemic on education?
@@
 
-Google Trends, is a website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, the largest search engine in the world, across various regions, subject domains and languages.
-Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution  [@alicino2015assessing].
+Google Trends, is a  freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages [@carneiro2009google].
+Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution  [@alicino2015assessing; @cook2011assessing].
+According to @jarynowski2020perception the public concerns on different matters can also take an epidemic nature  starting with the phase of growing interest which is known as "early adoption", to the phase of general interest which is known as "majority" to eventually lose popularity which is known as "lagers stage".
+Figure \ref{fig:distanceLearningWorldAnalysis} also confirms this life cycle explanation.
 
-The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding] and health care [@nuti2014use; @alicino2015assessing; @arora2019google].
-According to the study conducted by @willis2020using on taxi preferences found that online quires can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding] and health care [@nuti2014use; @alicino2015assessing; @arora2019google; @carneiro2009google; @cook2011assessing] as it provides data about social phenomena in a more timely manner than traditional data collection processes [@vaughan2014web].
+According to the study conducted by @willis2020using on taxi preferences found that online queries can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+In November, 2008, Google launched an internet-based surveillance tool, Google Flu Trends  (GFT), that uses aggregated Google search data to provide a near real-time support in predicting influenza outbreaks [@cook2011assessing].
 Several recent studies have also been conducted  to investigate the relation between Google search trend patterns and the COVID-19 related concerns [@husnayain2020applications; @effenberger2020association].
@@
 However, according to @moore2011learning the design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
-Therefore, this study    provides a detailed investigation on distance  education solutions focusing on  distance learning needs.
+Therefore, this study    provides a detailed investigation on distance  education solutions focusing on different distance learning needs.
+In contrast to the work in @kansal2021google, the work in our paper has a very specific focus: the distance learning solutions for various education purposes.
 
 Eventhough google trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool.
+It may contain inaccuracies [@carneiro2009google].
 According to @nuti2014use the data retrieval process should be transparent.
@@
 Furthermore, researchers should  clearly document the rationale and data retrieval process to ensure the reproducibility of results [@nuti2014use].
+.
 
@@
 
+In this work we use  Google Analytics search index series  as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
+We use weekly  Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
+We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the  period of Covid-19 pandemic.
+Through this we want to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
+We also study the correlation between weekly web searches and the number of  weekly new cases of covid-19.
+
+First "Distance education” and related terms were searched.
+
 We visualise our data and we then employ time series modelling to capture
 
-### Impact of covid 19 on education
+## Keyword selection
 
-In response to school closures, UNESCO recommended the use of distance learning programmes and open educational applications and platforms that schools and teachers can use to reach learners remotely and limit the disruption of education.
+There are different types  of interactions that facilitate learners to construct knowledge in distance education environments.
+These are learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction [@wallace2003online].
+
+Google Trends determines categories based on search patterns.
+According to @vaughan2014web  more relevant accurate data can be obtained by  limiting the search in Google Trends to a specific category as it helps  to reduce noise in the data.
+In this work our focus was given to education category.
+Google Trends does not provide the absolute search volume of a given search term.
+Instead it provides a relative search volume of a  particular term adjusted according to  the total searches of the geography and time range it represents.
+The resulted series  scale from 0 to 100, where each data point in the series represents the search interest relative to the highest point on the series for the selected region and time.
+[@alicino2015assessing; @vaughan2014web].
+A value of 100 represents the peak popularity of the term.
+
+The term distance education is known by a variety of names. The terms such as distance learning, e-Learning,  online learning environments, home study, independent study, external
+study, correspondence education, off campus study, open learning and open education  are used commonly and interchangeably  [@moore2011learning].
+It often describes the effort of providing access to learning for those who are geographically distant.
+The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
+Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places  and/or  varying forms of instructional materials [@moore2011learning].
+It is important to decide what term to search for in Google Trends as different terms have different search volumes.
+In the past literature several strategies have been used to select suitable search terms for a given study.
+Some studies have  selected search terms based on intuition and some through  brainstorming processes [@vaughan2014web].
+In this study our choice of search term  took into account the list of distance learning solutions published by UNESCO [@unesco2020DLsolutions].
+Eventhough, the solutions they had recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base and evidence of impact [@unesco2020DLsolutions].
+Different solutions are categorized based on distance learning needs.
+For each keyword, the worldwide search queries were performed with a keyword being used  as the "search term".
+That allowed us to search the exact string of text typed by the user.
+However, according to @vaughan2014web,  short forms or acronyms had higher search volumes than the corresponding complete names, in general.
+However we did not  use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
+
+<!-- completed up to page 30-->
+
+In google trend It is possible to search for up to five queries each time [21].
+
+Google Trends does not report the absolute search volume of a single term but does report the relative search volumes for up to five multiple terms [@vaughan2014web].
+
+ We entered up to five universities at a time and recorded their relative ranking scores, we then entered a different group of universities, some of which overlapped with the previous group, and we recorded the relative ranking scores again.
+Through this repeated process of relative comparison, we were able to obtain relative ranking scores for all uni- versities in the study.
+(Vaughan & Romero-Frias 2014)
 
@@
 
+bola the most searched item on Yahoo engine and the fifth most searched term on Google in 2014 [2, 3].
+
+Duifferent attesnion for different etrms 
+The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
+The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane & Galyen 2011).
+
+## exams
+
+Strategies to postpone, skip or adminsiter examinations at a distance raise serious concerns about fairness, especialy when access to learning becomes variable.
+Disruptions to assessments results in stress for students and their families and can trigger disengagement.
+[@unesco2020covid; @unescoadverse2020]
 ## Discussion and Conclusion

### 2021-12-25 Paper_online_Learning.tex
update analysis

@@
 Through this we want to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also study the correlation between weekly web searches and the number of weekly new cases of covid-19.
-
-First ``Distance education'' and related terms were searched.
-
-We visualise our data and we then employ time series modelling to capture
-
-\hypertarget{keyword-selection}{%
-\subsection{Keyword selection}\label{keyword-selection}}
+We also study the correlation between weekly web searches and the number of weekly cases of covid-19.
 
@@
 
-In google trend It is possible to search for up to five queries each time {[}21{]}.
+Google Trends allows to search for up to five queries each time \autocite{vaughan2014web}.
+We did our analysis under 11 sub segmentations namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
+The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to Covid 19 pandemic.
 
-Google Trends does not report the absolute search volume of a single term but does report the relative search volumes for up to five multiple terms \autocite{vaughan2014web}.
+In google trends, it is possible to search for only up to five queries at a time.
+Therefore, under each segmentation we entered up to five distance learning tools at a time and recorded their relative ranking scores.
+An iterative pairwise comparison was first used to identify the series that has highest search volume during the study period.
+Through this iterative process of relative comparison, we were able to identify the tool with the highest relative ranking score.
+Then the relative ranking scores for all the other tools in the given segmentation was obtained relative to the previously detected series with the highest ranking scores \autocite{vaughan2014web}.
 
-We entered up to five universities at a time and recorded their relative ranking scores, we then entered a different group of universities, some of which overlapped with the previous group, and we recorded the relative ranking scores again.
-Through this repeated process of relative comparison, we were able to obtain relative ranking scores for all uni- versities in the study.
-(Vaughan \& Romero-Frias 2014)
+Both visual and analytical approaches were used to analyse the changes in web search queries at the global level over the entire period of the Covid-19 pandemic to identify the emerging evidence about the impact of COVID-19 pandemic on distance education.
+We also evaluated the relationship between weekly web searches and the number of weekly global cases of covid-19.
+Cross correlation analysis and the dynamic time warping algorithm \autocite{giorgino2009computing} were performed to investigate the relationship and similarities between the signatures of covid-19 related search volume data and distance education related search volume data.
 
-\hypertarget{analysis}{%
-\subsection{Analysis}\label{analysis}}
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
-bola the most searched item on Yahoo engine and the fifth most searched term on Google in 2014 {[}2, 3{]}.
+Search volume index graphs for Covid 19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic.
+Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic.
+According to \textcite{moore2011learning}, ``distance education'' is the most renowned descriptor used when referencing distance learning.
+However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``never version'' and ``improved version'' of distance learning.
+The google search trend pattern corresponds to the term ``online learning'' (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout Covid-19 pandemic.
+Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``Distance learning'' and ``online teaching'' demonstrate a similar trend during the study period.
+Covid-19 related web searches declined in the months after the WHO announcement of the Covid-19 pandemic, despite the increase in Covid-19 cases.
+A similar decline could be detected in distance education related terms.
+further, a second was observed only in education related terms after the WHO announcement of the Covid-19 pandemic.
+This could be due to the near total closures of schools happened with the declaration of Covid-19 as a pandemic.
 
-Duifferent attesnion for different etrms
-The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
-The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane \& Galyen 2011).
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth]{figure/dtw1-1} \includegraphics[width=1\textwidth]{figure/dtw1-2}
+
+}
+
+\caption{Visualization of Dynamic Time Warping Alignments between 'online learning' and 'corona'}\label{fig:dtw1}
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth]{figure/dtw2-1} \includegraphics[width=1\textwidth]{figure/dtw2-2}
+
+}
+
+\caption{Visualization of Dynamic Time Warping Alignments  'online learning' and 'covid'}\label{fig:dtw2}
+\end{figure}
+
+The pairwise comparison performed using Dynamic Time Warping algorithm.
+Under this approach the two time series are stretched or compressed locally in order to make one resemble the other as much as possible.
+The distance between the two series is then computed by summing the distances of individual aligned elements \autocite{giorgino2009computing}.
+Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact from covid 19 on distance education.
+Both pairwise comparison show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)).
+The normalized distance between the two series in the two comparisons are 0.1326197 and 0.1828299, respectively.
+This also provides further confirmation for the similarity of the google trend search profiles.
+
+The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to covid 19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after covid 19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+
+The degree of digression of blue line from the ideal 45-degree straight line indicates a dissimilarity of the two time
+series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+In 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide.
+By the time it was know as ``corona'', it was not a global concern and was mostly limited to China.
+During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to Covid 19 pandemic all around the world.
+This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis.
+
+Later with the spread of the virus worldwide, WHO named it as COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
+With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
+With this sudden unexpected movement, online learning also became an equally important global concern, similar to covid 19 pandemic.
+This similar behaviour in public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line from approximately 120 to 150 on the Query Index axis.
+This also confirms the impact of covid 19 on distance education and online learning.
+
+Then, a cross correlation analysis was performed to estimate the lag between covid 19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
+A close relationship was evident between covid related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''.
+The significant cross-correlation coefficients at lag zero confirms contemporaneous relationship between the covid related search terms and distance learning related search terms.
+The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
+However, no significant synchronization was observed between the search patterns of ``distance education'' and ``online proctoring'' and covid 19 related search trends (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth]{figure/ccfAnalysis-1}
+
+}
+
+\caption{Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic}\label{fig:ccfAnalysis}
+\end{figure}
+
+\newpage
+
+\hypertarget{gt-footprint}{%
+\subsection{GT footprint}\label{gt-footprint}}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth]{figure/plot1-1}
+
+}
+
+\caption{GT footprint}\label{fig:plot1}
+\end{figure}
+
+\newpage
+
+\hypertarget{gt-footprint-1}{%
+\subsection{GT footprint}\label{gt-footprint-1}}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth]{figure/plot2-1}
+
+}
+
+\caption{GT footprint}\label{fig:plot2}
+\end{figure}
 
@@
 \autocite{unesco2020covid,unescoadverse2020}
-\#\# Discussion and Conclusion
+
+\hypertarget{discussion-and-conclusion}{%
+\subsection{Discussion and Conclusion}\label{discussion-and-conclusion}}
+
+The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
+The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane \& Galyen 2011).
 

### 2021-12-25 main.Rmd
update analysis

@@
 Through this we want to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also study the correlation between weekly web searches and the number of  weekly new cases of covid-19.
-
-First "Distance education” and related terms were searched.
-
-We visualise our data and we then employ time series modelling to capture
-
-## Keyword selection
+We also study the correlation between weekly web searches and the number of weekly  cases of covid-19.
 
@@
 
-In google trend It is possible to search for up to five queries each time [21].
+Google Trends allows to search for up to five queries each time [@vaughan2014web].
+We did our analysis under 11 sub segmentations namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
+The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to Covid 19 pandemic.
 
-Google Trends does not report the absolute search volume of a single term but does report the relative search volumes for up to five multiple terms [@vaughan2014web].
+In google trends, it is possible to search for only up to five queries at a  time.
+Therefore, under each segmentation we entered up to five distance learning tools at a time and recorded their relative ranking scores.
+An iterative pairwise comparison was first used to identify the series that has highest search volume during the study period.
+Through this iterative process of relative comparison, we were able to identify the tool with the highest relative ranking score.
+Then the relative ranking scores for all the other tools in the given segmentation was obtained relative to the previously detected series with the highest ranking scores [@vaughan2014web].
 
- We entered up to five universities at a time and recorded their relative ranking scores, we then entered a different group of universities, some of which overlapped with the previous group, and we recorded the relative ranking scores again.
-Through this repeated process of relative comparison, we were able to obtain relative ranking scores for all uni- versities in the study.
-(Vaughan & Romero-Frias 2014)
+Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level over the entire period of the Covid-19 pandemic to identify the emerging evidence about the impact of  COVID-19 pandemic on distance education.
+We also evaluated the  relationship between weekly web searches and the number of  weekly global cases of covid-19.
+Cross correlation analysis and  the dynamic time warping algorithm [@giorgino2009computing] were performed  to investigate the relationship and similarities between the signatures of  covid-19 related search volume data  and distance education related search volume data.
 
-## Analysis
+## Results
 
-bola the most searched item on Yahoo engine and the fifth most searched term on Google in 2014 [2, 3].
+Search volume index graphs for Covid 19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic.
+Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic.
+According to @moore2011learning, "distance education" is the most renowned descriptor used when referencing distance learning.
+However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version", "never version" and "improved version" of distance learning.
+The google search trend pattern corresponds to the term "online learning"  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout  Covid-19 pandemic.
+Despite quantitative differences in terms of relative search volume, the series generated by the two terms "Distance learning" and "online teaching" demonstrate a similar trend during the study period.
+Covid-19 related web searches declined in the months after the WHO announcement of the Covid-19 pandemic, despite the increase in Covid-19 cases.
+A similar decline could be detected in distance education related terms.
+further,  a second was observed only in education related terms  after the WHO announcement of the Covid-19 pandemic.
+This could be due to the near total closures of schools happened  with the declaration of  Covid-19 as a pandemic.
 
-Duifferent attesnion for different etrms 
-The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
-The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane & Galyen 2011).
+```{r dtw, fig.height=5, fig.cap= "Visualization of Dynamic Time Warping Alignments"}
+```
+
+```{r dtw1, fig.height=4, fig.cap= "Visualization of Dynamic Time Warping Alignments between 'online learning' and 'corona'", dependson="dtw"}
+```
+
+```{r dtw2, fig.height=4, fig.cap= "Visualization of Dynamic Time Warping Alignments  'online learning' and 'covid'", dependson="dtw"}
+```
+
+The pairwise comparison performed  using  Dynamic Time Warping algorithm.
+Under this approach the two time series are stretched or compressed  locally in order to make one resemble the other as much as possible.
+The distance between the two series is then computed by summing the distances of individual aligned elements [@giorgino2009computing].
+Two pairwise comparisons, "corona"  with "online learning" and   "covid" with "online learning" were conducted to see the impact  from covid 19 on distance education.
+Both pairwise comparison show a significant  areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and  \ref{fig:dtw2} (a)).
+The normalized distance between the two series in the two comparisons are  0.1326197 and 0.1828299, respectively.
+This also provides further confirmation for  the similarity of the google trend  search profiles.
+
+The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to covid 19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after covid 19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+
+The degree of digression of blue line from the ideal 45-degree straight line indicates a dissimilarity of the two time
+series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+In 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide.
+By the time it was know as "corona", it was not a global concern and was mostly limited to China.
+During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to Covid 19 pandemic all around the world.
+This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis.
+
+Later with the spread of the virus worldwide, WHO  named it as  COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
+With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
+With this sudden unexpected movement, online learning  also became an equally important global concern, similar to covid 19 pandemic.
+This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line  from approximately 120 to 150 on the Query Index axis.
+This also confirms the impact of covid 19 on distance education and online learning.
+
+Then, a cross correlation analysis was performed to estimate the lag between covid 19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
+A close relationship was evident between covid related search patterns and the search patterns of the terms "Online learning", "Distance learning" and "Online teaching".
+The significant cross-correlation coefficients  at lag zero confirms  contemporaneous relationship between the covid related search terms and distance learning related search terms.
+The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
+However, no significant  synchronization was observed  between the search patterns of "distance education" and "online proctoring" and covid 19 related search trends (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
+
+```{r ccfAnalysis, fig.height=8, fig.cap= "Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic"}
+```
+
+```{r Digital_lms_data_download, eval=FALSE}
+```
+
+```{r DigitalLmsAnalysis, cache=FALSE}
+```
+
+```{r mobile_phones_apps_data_download, eval=FALSE}
+```
+
+```{r mobile_phones_apps_analysis, dependson="DigitalLmsAnalysis"}
+```
+
+```{r offline_functionality_data_download, eval=FALSE}
+```
+
+```{r offline_functionality_analysis, dependson="mobile_phones_apps_analysis"}
+```
+
+```{r mooc_data_download, eval=FALSE}
+```
+
+```{r mooc_analysis, dependson="offline_functionality_analysis"}
+```
+
+```{r Self_directed_learning_data_download, eval=FALSE}
+```
+
+```{r Self_directed_learning_analysis, dependson="mooc_analysis"}
+```
+
+```{r Mobile_reading_data_download, eval=FALSE}
+```
+
+```{r Mobile_reading_analysis, dependson="Self_directed_learning_analysis"}
+```
+
+```{r Collaboration_platforms_data_download, eval=FALSE}
+```
+
+```{r Collaboration_platforms_eduonly_analysis, dependson="Mobile_reading_analysis"}
+```
+
+```{r Tools_for_teachers_data_download, eval=FALSE}
+```
+
+```{r Tools_for_teachers_analysis, dependson="Collaboration_platforms_eduonly_analysis"}
+```
+
+```{r External_repositories_DL_data_download, eval=FALSE}
+```
+
+```{r External_repositories_DL_analysis, dependson="Tools_for_teachers_analysis"}
+```
+
+```{r Online_proctoring_data_download, eval=FALSE}
+```
+
+```{r Online_proctoring_analysis, dependson="External_repositories_DL_analysis"}
+```
+
+```{r psychosocial_support_data_download, eval=FALSE}
+```
+
+\newpage
+## GT footprint
+
+```{r plot1, fig.height=12, fig.cap="GT footprint"}
+```
+
+\newpage
+
+## GT footprint
+```{r plot2, fig.height=12, fig.cap="GT footprint"}
+```
 
@@
 [@unesco2020covid; @unescoadverse2020]
+
 ## Discussion and Conclusion
+
+The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
+The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane & Galyen 2011).

### 2021-12-25 src/main.R

@@ -320,81 +320,181 @@ p <- p1/p3/p5
 print(p)
 
 
-## ---- ccf_analysis_covid_online_edu
-# Cross-Correlation Function Estimation
-data_world <- read.csv( here::here(
-  "Paper_online_Learning_Dec", "data",
-  "education_search_world.csv"))
-# hits is a character variable. Convert hits into a numeric variable
-hits1<- ifelse(data_world$hits ==  "<1", 0,
-               data_world$hits)
-data_world$hits <- as.numeric(hits1)
-keyword_ord<- data_world %>% 
-  group_by(keyword) %>%
-  summarise(total_hits = 
-              sum(hits, na.rm = TRUE)) %>%
-  arrange(total_hits) %>% 
-  select(keyword) %>% 
-  as_vector()
-dataW_edu <- data_world %>%
-  select(date, keyword, hits) %>%
-  mutate(keyword = factor(keyword, 
-                          levels = keyword_ord))
-dataW_edu$date = as.Date(dataW$date, "%Y-%m-%d")
-print(head(dataW_edu))
 
-d_edu<- dataW_edu %>% 
-  filter(keyword == "Online teaching") %>%
-  select(-keyword) 
-d_edu$date = as.Date(d_edu$date, "%Y-%m-%d")
-d_edu <- d_edu %>%as_tsibble(index = date) 
-print(head(d_edu))
+## ---- dtw
 
+# Covid cases
+covidcasesW <- read.csv(here::here( "data", "covid_cases.csv"))
+covidcasesW <- covidcasesW %>% select(-X)
+
+## covid search google trends
 data_covid <- read.csv( here::here(
-  "Paper_online_Learning_Dec", "data", 
-  "covid_search_world.csv"))
+  "data", "covid_search_world.csv"))
 hits1<- ifelse(data_covid$hits ==  "<1", 0, data_covid$hits)
 data_covid$hits <- as.numeric(hits1)
-keyword_ord<- data_covid %>% 
-  group_by(keyword) %>%
-  summarise(total_hits = sum(hits, na.rm = TRUE)) %>%
-  arrange(total_hits) %>% select(keyword) %>% as_vector()
+# covid search google trends - arrange in wide format
+covid_search <- data_covid %>%
+  select(date, hits, keyword) %>%
+  pivot_wider(names_from = keyword, values_from = hits)
 
-dataW_covid <- data_covid %>%
-  select(date, keyword, hits) %>%
[…]

### 2021-12-26 Paper_online_Learning.tex
update results and discussion

@@
 
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
+\hypertarget{results-and-discussion}{%
+\subsection{Results and Discussion}\label{results-and-discussion}}
 
@@
 The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
-However, no significant synchronization was observed between the search patterns of ``distance education'' and ``online proctoring'' and covid 19 related search trends (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
+However, no significant synchronization was observed between the Covid 19 related search trends and the search patterns of ``distance education'' or ``online proctoring'' (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
 
@@
 \end{figure}
+
+The list of distance education solutions published by UNESCO provides many different distance education solutions for different education purposes.
+Figure \ref{fig:plot1} and Figure \ref{fig:plot2} provide a Google trend footprint that can be used as a proxy for measuring global attention towards different distance learning solutions.
+This provides a fast first step to identify the most popular tools available for different educational purposes.
+Under each and every category Figure \ref{fig:plot1} (a) - (e) and Figure \ref{fig:plot2} (f) - (j), one or two tools have got a significantly higher level of attention in comparison to the remaining options available undue the same category.
+The time series across multiple panels are not comparable as they represent ``relative'' search volume of given terms and they are comparable only with a given panel.
+The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each category.
+
+Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during covid 19 pandemic (Figure \ref{fig:plot1} (a)).
+This might be due to it's easy and fast setup and usage \autocite{sudarsana2019use}.
+Under systems built for use on basic mobile phones, both `kaiOS' and `Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
+`KaiOs' has the ability provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
+`Ubongo' has also received a considerable attention, eventhough its target groups are school-age children and their parents in Africa.
+It facilitates localized, multi-platform Entertainment-Education to African families at low cost and massive scale and is available in both Kiswahili and English.
+A significant peak in `Ubongo' series can be observed during 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
+Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during covid-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+The can be a huge relief for financially distressed families during Covid-19 pandemic.
+Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has got a considerable attention in comparison to other options available.
+`Udemy' has got the second highest attention.
+However, `Canvas' has received far more attention than `Udemy'.
+Under self-directed learning content `Quizlet' has got a considerably high attention throughout the pandemic in comparison to the other options available under the same category.
+It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and laguages.
+The second highest option only provides support for language learning.
+`Quizlet'`s ability provide support for multiple subject might be the reason for it to outperform it competitor in the market.
+Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option.
+Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had got a considerable attention at the start of the pandemic in March 2020.
+However, `WhatsApp' has outperformed `Zoom' overtime through with increasing public attention during the pandemic.
+In contrary, the global attention towards `Zoom' has decreased over time.
+`Teams' has also got considerable attention in comparison to the other remaining options available int he market.
+Under tools for teachers to create of digital learning content, `Trello' which provides a visual collaboration platform for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
+`Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic.
+A significant peak can be observed in the `Nearpod' series, during the first academic break.
+Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got a significant global attention.
+`Brookings' provides a large catalogue of learning innovations eventhough its focus is not limited to distance learning solutions.
+`UNHCR' also provides an extensive list of distance learning solutions from the United Nations agency for refugees.
 
@@
 
+The attention in the education sector is low compared to all category.
+However a significant peak can be observed during the first academic break after declaring Covid 19 as a pandemic.
+Similar peak can be observed in 2021 .
+But that significant compared to 2020 peak.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=1\textwidth]{figure/psychosocialSupportAnalysis-1}
+
+}
+
+\caption{Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' category}\label{fig:psychosocialSupportAnalysis}
+\end{figure}
+
 \hypertarget{exams}{%
@@
 
-\hypertarget{discussion-and-conclusion}{%
-\subsection{Discussion and Conclusion}\label{discussion-and-conclusion}}
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
 
@@
 
+\url{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0248029}
+
+First and second waves of coronavirus disease-19: A comparative study in hospitalized patients in Reus, Spain
+
 \printbibliography

### 2021-12-26 main.Rmd
update results and discussion

@@
 
-## Results
+## Results and Discussion
 
@@
 The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
-However, no significant  synchronization was observed  between the search patterns of "distance education" and "online proctoring" and covid 19 related search trends (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
+However, no significant  synchronization was observed  between the Covid 19 related search trends and the search patterns of "distance education" or "online proctoring" (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
 
-```{r ccfAnalysis, fig.height=8, fig.cap= "Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic"}
+```{r ccfAnalysis, fig.height=7, fig.cap= "Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic"}
 ```
+
+The list of distance education solutions published by UNESCO provides many different distance education solutions for different education purposes.
+Figure \ref{fig:plot1} and  Figure \ref{fig:plot2} provide a Google trend footprint that can be used as a proxy for measuring global attention towards different distance learning solutions.
+This provides a fast first step to identify the most popular tools available for different educational purposes.
+Under each and every category Figure \ref{fig:plot1} (a) - (e) and  Figure \ref{fig:plot2} (f) - (j), one or two  tools have got a significantly higher level of attention in comparison to the remaining options available undue the same category.
+The time series across multiple panels are not comparable as they represent "relative" search volume of given terms and they are comparable only with a given panel.
+The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each category.
+
+Under digital learning management systems "Google Classroom", which  helps  teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during covid 19 pandemic (Figure \ref{fig:plot1} (a)).
+This might be due to it's easy and fast setup and usage [@sudarsana2019use].
+Under systems built for use on basic mobile phones, both 'kaiOS' and  'Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
+'KaiOs'  has the ability provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
+'Ubongo' has also received a considerable attention, eventhough its target groups are school-age children and their parents in Africa.
+It facilitates localized, multi-platform Entertainment-Education  to African families at low cost and massive scale and is  available in both Kiswahili and English.
+A significant peak in  'Ubongo' series can be observed during 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
+Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during covid-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+The can be a huge relief for financially distressed families during Covid-19 pandemic.
+Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has got a considerable attention in comparison to other options available.
+'Udemy' has got the second highest attention.
+However, 'Canvas' has received far more attention than 'Udemy'.
+Under self-directed learning content 'Quizlet' has got a considerably high attention  throughout the pandemic in comparison to the other options available under the same category.
+It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and laguages.
+The second highest option only provides  support for language learning.
+'Quizlet''s ability provide support for multiple subject might be the reason for it to outperform it  competitor in the market.
+Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option.
+Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had got a considerable attention at the start of the pandemic in March 2020.
+However, 'WhatsApp' has outperformed 'Zoom' overtime through with increasing public attention during the pandemic.
+In contrary, the global attention towards 'Zoom'  has decreased over time.
+'Teams' has also got considerable attention in comparison to the other remaining options available int he market.
+Under tools for teachers to create of digital learning content, 'Trello'  which provides a visual collaboration platform  for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
+'Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic.
+A significant peak can be observed in the 'Nearpod' series, during the first academic break.
+Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got a significant global attention.
+'Brookings' provides a large catalogue of learning innovations eventhough its focus is not limited to distance learning solutions.
+'UNHCR' also  provides an extensive list of distance learning solutions from the United Nations agency for refugees.
 
@@
 
+The attention in the education sector is low compared to all category.
+However a significant peak can be observed during the first academic break  after declaring  Covid 19 as a pandemic.
+Similar peak can be observed in 2021 .
+But that significant compared to 2020 peak.
+
+```{r psychosocialSupportAnalysis, fig.height= 3, fig.cap="Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' category"}
+```
+
 ## exams
@@
 
-## Discussion and Conclusion
+## Conclusions
 
@@
 The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane & Galyen 2011).
+
+https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0248029
+
+First and second waves of coronavirus disease-19: A comparative study in hospitalized patients in Reus, Spain

### 2021-12-26 src/main.R

@@ -451,39 +451,39 @@ p5 <- data %>%
 
 
 pa <- data %>% 
-  mutate(diff_corona = difference(`covid 19`),
+  mutate(diff_corona = difference(covid),
          diff_online = difference(`Online Learning`)) %>%
   CCF(diff_corona, diff_online ) %>%
   autoplot()+
-  labs(title= "(f) 'Covid 19' with 'Online learning")
+  labs(title= "(f) 'Covid' with 'Online learning")
 
 pb <- data %>% 
-  mutate(diff_corona = difference(`covid 19`),
+  mutate(diff_corona = difference(covid),
          diff_dist = difference(`Distance learning`)) %>%
   CCF(diff_corona, diff_dist ) %>%
   autoplot()+
-  labs(title= "(g) 'Covid 19' with 'Distance learning'")
+  labs(title= "(g) 'Covid' with 'Distance learning'")
 
 pc <- data %>% 
-  mutate(diff_corona = difference(`covid 19`),
+  mutate(diff_corona = difference(covid),
          diff_teach = difference(`Online teaching`)) %>%
   CCF(diff_corona, diff_teach ) %>%
   autoplot()+
-  labs(title= "(h) 'Covid 19' with 'Online teaching'")
+  labs(title= "(h) 'Covid' with 'Online teaching'")
 
 pd <- data %>% 
-  mutate(diff_corona = difference(`covid 19`),
+  mutate(diff_corona = difference(covid),
          diff_edu = difference(`Distance education`)) %>%
   CCF(diff_corona, diff_edu ) %>%
   autoplot()+
-  labs(title= "(i) 'Covid 19' with 'Distance education'")
+  labs(title= "(i) 'Covid' with 'Distance education'")
 
 pe <- data %>% 
-  mutate(diff_corona = difference(`covid 19`),
+  mutate(diff_corona = difference(covid),
          diff_proc = difference(`online proctoring`)) %>%
   CCF(diff_corona, diff_proc ) %>%
   autoplot()+
-  labs(title= "(j) 'Covid 19' with 'Online proctoring'")
+  labs(title= "(j) 'Covid' with 'Online proctoring'")
 
 
 (p1 | pa ) / 
@@ -675,7 +675,7 @@ keywords= c("Cell-Ed", "Eneza Education", "Funzi",
 keywords1 <- c("Ubongo", "Cell-Ed", "Eneza Education", "Funzi", "Ubongo")
 trends1 <-  gtrends(keywords1, gprop =channel, time = time )
[…]

### 2021-12-26 Paper_online_Learning.tex
upadte results and sicussion

@@
 The can be a huge relief for financially distressed families during Covid-19 pandemic.
-Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has got a considerable attention in comparison to other options available.
+Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
 `Udemy' has got the second highest attention.
 However, `Canvas' has received far more attention than `Udemy'.
-Under self-directed learning content `Quizlet' has got a considerably high attention throughout the pandemic in comparison to the other options available under the same category.
-It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and laguages.
+Under self-directed learning content `Quizlet' has got a considerably high attention throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)).
+It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and languages.
 The second highest option only provides support for language learning.
 `Quizlet'`s ability provide support for multiple subject might be the reason for it to outperform it competitor in the market.
-Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option.
-Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had got a considerable attention at the start of the pandemic in March 2020.
+Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
+Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
 However, `WhatsApp' has outperformed `Zoom' overtime through with increasing public attention during the pandemic.
@@
 Under tools for teachers to create of digital learning content, `Trello' which provides a visual collaboration platform for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
-`Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic.
+`Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).
 A significant peak can be observed in the `Nearpod' series, during the first academic break.
-Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got a significant global attention.
+Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got a significant global attention (Figure \ref{fig:plot2} (i)).
 `Brookings' provides a large catalogue of learning innovations eventhough its focus is not limited to distance learning solutions.
 `UNHCR' also provides an extensive list of distance learning solutions from the United Nations agency for refugees.
+Student assessment is vital in ensuring the quality of education.
+However, due unexpected sudden school closure, administering examinations at a distance became a major concern.
+Under tools for online proctoring, `Pearson VUE' which provides computer based testing services in
+over 150 countries worldwide with more than 5,000 authorized test centers has got a considerably high public attention during the pandemic (Figure \ref{fig:plot2} (j)).
 
-\newpage
-
-\hypertarget{gt-footprint}{%
-\subsection{GT footprint}\label{gt-footprint}}
+At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
+These peaks aligns with the near-total closures of schools due to covid 19 in Figure \ref{fig:covidImpactWorld}.
+The major concerns that arise with the unexpected sudden movement from physical classroom to virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
+Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
+The inability to reopen school fully due to the increasing Covid-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.
 
@@
 
-\caption{GT footprint}\label{fig:plot1}
+\caption{Google search trends footprint analysis}\label{fig:plot1}
 \end{figure}
-
-\newpage
-
-\hypertarget{gt-footprint-1}{%
-\subsection{GT footprint}\label{gt-footprint-1}}
 
@@
 
-\caption{GT footprint}\label{fig:plot2}
+\caption{Google search trends footprint analysis}\label{fig:plot2}
 \end{figure}
@@
 
-\hypertarget{exams}{%
-\subsection{exams}\label{exams}}
-
-Strategies to postpone, skip or adminsiter examinations at a distance raise serious concerns about fairness, especialy when access to learning becomes variable.
-Disruptions to assessments results in stress for students and their families and can trigger disengagement.
-\autocite{unesco2020covid,unescoadverse2020}
+Good psychosocial support is also vital to improve both students' and teachers' mental health and well-being during a pandemic.
+This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two suddent spikes in the search trend series of the term `Psychosocial support' during the academic breaks in August 2020 and in August 20201 under `Education' category.
+This highlight the urgent need of attention and solutions towards these aspects.
+This type of an analysis based on secondary data is very vital as these types of aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
 

### 2021-12-26 main.Rmd
upadte results and sicussion

@@
 The can be a huge relief for financially distressed families during Covid-19 pandemic.
-Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has got a considerable attention in comparison to other options available.
+Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
 'Udemy' has got the second highest attention.
 However, 'Canvas' has received far more attention than 'Udemy'.
-Under self-directed learning content 'Quizlet' has got a considerably high attention  throughout the pandemic in comparison to the other options available under the same category.
-It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and laguages.
+Under self-directed learning content 'Quizlet' has got a considerably high attention  throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)).
+It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and languages.
 The second highest option only provides  support for language learning.
 'Quizlet''s ability provide support for multiple subject might be the reason for it to outperform it  competitor in the market.
-Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option.
-Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had got a considerable attention at the start of the pandemic in March 2020.
+Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
+Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
 However, 'WhatsApp' has outperformed 'Zoom' overtime through with increasing public attention during the pandemic.
@@
 Under tools for teachers to create of digital learning content, 'Trello'  which provides a visual collaboration platform  for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
-'Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic.
+'Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).
 A significant peak can be observed in the 'Nearpod' series, during the first academic break.
-Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got a significant global attention.
+Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got a significant global attention (Figure \ref{fig:plot2} (i)).
 'Brookings' provides a large catalogue of learning innovations eventhough its focus is not limited to distance learning solutions.
 'UNHCR' also  provides an extensive list of distance learning solutions from the United Nations agency for refugees.
+Student assessment is vital in ensuring the quality of education.
+However, due unexpected sudden school closure, administering examinations at a distance became a major concern.
+Under tools for online proctoring, 'Pearson VUE' which provides computer based testing services in
+over 150 countries worldwide with more than 5,000 authorized test centers has got  a considerably high public attention during the pandemic (Figure \ref{fig:plot2} (j)).
+
+At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a))  and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
+These peaks aligns with the near-total closures of schools due to covid 19 in Figure \ref{fig:covidImpactWorld}.
+The major concerns that arise with the unexpected sudden movement from physical classroom to  virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
+Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
+The inability to reopen school fully due to the increasing Covid-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reason for this considerable public attention around August, 2020.
 
@@
 
-\newpage
-## GT footprint
-
-```{r plot1, fig.height=12, fig.cap="GT footprint"}
+```{r plot1, fig.height=12, fig.cap="Google search trends footprint analysis"}
 ```
 
-\newpage
-
-## GT footprint
-```{r plot2, fig.height=12, fig.cap="GT footprint"}
+```{r plot2, fig.height=12, fig.cap="Google search trends footprint analysis"}
 ```
@@
 
-## exams
-
-Strategies to postpone, skip or adminsiter examinations at a distance raise serious concerns about fairness, especialy when access to learning becomes variable.
-Disruptions to assessments results in stress for students and their families and can trigger disengagement.
-[@unesco2020covid; @unescoadverse2020]
+Good psychosocial support is also vital to improve both students' and  teachers' mental health and well-being during a pandemic.
+This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two suddent spikes in the search trend series of  the term 'Psychosocial support' during the academic breaks in August 2020 and in August 20201 under 'Education' category.
+This highlight the urgent need of attention and solutions towards these aspects.
+This type of an analysis based on secondary data is very vital as these types of aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
 

### 2021-12-26 Paper_online_Learning.Rmd
update paper

@@
   othernames: Thiyanga S.
-  address: Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
-  email: tstalagala@gmail.com
-abstract: "A brief summary of our work"
-keywords: "rmarkdown, templates"
+  address: Department of Statistics, Faculty of Applied Sciences, University of Sri Jayewardenepura, Sri Lanka
+  email: ttalagala@sjp.ac.lk
+abstract: "Distance Education has long history.
+However Coid-19 has made a new era of distance education.
+With the increasing demand, many different distance learning solutions have been introduced for different education purposes.
+In this study we invetigated the impact of Covid 19 on ditance education and the possibility of using Google analytics search index series  as a proxy to quantify the  popularity and the public interest for different distance education solutions.
+Both visual and analytical approaches were used to  analyse web search queries at the global level during Covid-19 pandemic.This study intorduced a Google trend footprint of various ditance leraning solutions for diffrent education purposes, that can be used as a proxy for measuring global attension towards different learning solutions.
+The findings provide a fast first step to identify the most popular tools available for different educational purposes.
+R code and data are available in the online supplementary materials.
+"
+keywords:  "Online-learning, Distance education solutions, Covid-19 Pandemic,  Google trend search queries "
 wpnumber: 1
-jelcodes: C10,C14,C22
+jelcodes: C82,I20,I21
 blind: false

### 2021-12-26 Paper_online_Learning.tex
update paper

@@
             pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
-            pdfkeywords={rmarkdown, templates},
+            pdfkeywords={Online-learning, Distance education solutions, Covid-19 Pandemic, Google trend search queries},
             pdfborder={0 0 0},
@@
 \wp{1}
-\jel{C10,C14,C22}
+\jel{C82,I20,I21}
 
@@
 \textbf{Thiyanga S. Talagala}\newline
-Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
-\newline{Email: \href{mailto:tstalagala@gmail.com}{\nolinkurl{tstalagala@gmail.com}}}\\[1cm]
+Department of Statistics, Faculty of Applied Sciences, University of Sri Jayewardenepura, Sri Lanka
+\newline{Email: \href{mailto:ttalagala@sjp.ac.lk}{\nolinkurl{ttalagala@sjp.ac.lk}}}\\[1cm]
 }
@@
 \begin{abstract}
-A brief summary of our work
+Distance Education has long history.
+However Coid-19 has made a new era of distance education.
+With the increasing demand, many different distance learning solutions have been introduced for different education purposes.
+In this study we invetigated the impact of Covid 19 on ditance education and the possibility of using Google analytics search index series as a proxy to quantify the popularity and the public interest for different distance education solutions.
+Both visual and analytical approaches were used to analyse web search queries at the global level during Covid-19 pandemic.This study intorduced a Google trend footprint of various ditance leraning solutions for diffrent education purposes, that can be used as a proxy for measuring global attension towards different learning solutions.
+The findings provide a fast first step to identify the most popular tools available for different educational purposes.
+R code and data are available in the online supplementary materials.
 \end{abstract}
 \begin{keywords}
-rmarkdown, templates
+Online-learning, Distance education solutions, Covid-19 Pandemic, Google trend search queries
 \end{keywords}
@@
 Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}.
-Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020.
+Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 \autocite{unesco2020covid}.
 With the start of the pandemic a near total closure of schools were observed all around the world.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactContinent}).
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools where classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
 
@@
 The remainder of this paper is organized as follows.
-Section 2 presents the related work to lay the foundation for the Google footprint analysis.
+Section 2 presents the related work to lay the foundation for the Google trend footprint analysis of different distance learning solutions.
 Section 3 presents the methodology followed in the study.
-Section 4 includes Google footprint analysis.
+Section 4 includes results and discussion.
 Section 5 concludes the article and presents future research directions.
@@
 \hypertarget{background}{%
-\subsection{Background}\label{background}}
+\section{Background}\label{background}}
+
+The term distance education is known by a variety of names. The terms such as distance learning, e-Learning, online learning environments, home study, independent study, external
+study, correspondence education, off campus study, open learning and open education are used commonly and interchangeably \autocite{moore2011learning}.
+It often describes the effort of providing access to learning for those who are geographically distant.
+The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
+Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places and/or varying forms of instructional materials \autocite{moore2011learning}.
+There are different types of interactions that facilitate learners to construct knowledge in distance education environments.
+These are learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction \autocite{wallace2003online}.
 
@@
 \hypertarget{methodology}{%
-\subsection{Methodology}\label{methodology}}
+\section{Methodology}\label{methodology}}
 
-In this work we use Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
-We use weekly Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
+To identify the impact of covid 19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO \autocite{unesco2020covid}.
+Data related to Covid-19 cases was retrieved from the R software \autocite{rsoftware} package \texttt{coronavirus} \autocite{coronavirusr}.
+The package pulls data from the Johns
+Hopkins University Center for Systems Science and Engineering (JHU CCSE) data repository \autocite{dong2020interactive}.
+
+In this work we also used Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
+We used weekly Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
 We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of Covid-19 pandemic.
-Through this we want to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also study the correlation between weekly web searches and the number of weekly cases of covid-19.
-
-There are different types of interactions that facilitate learners to construct knowledge in distance education environments.
-These are learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction \autocite{wallace2003online}.
+Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
+We also studied the cross correlation between weekly web searches and the weekly cases of covid-19.
 
@@
 
-The term distance education is known by a variety of names. The terms such as distance learning, e-Learning, online learning environments, home study, independent study, external
-study, correspondence education, off campus study, open learning and open education are used commonly and interchangeably \autocite{moore2011learning}.
-It often describes the effort of providing access to learning for those who are geographically distant.
-The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
-Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places and/or varying forms of instructional materials \autocite{moore2011learning}.
 It is important to decide what term to search for in Google Trends as different terms have different search volumes.
@@
 \hypertarget{results-and-discussion}{%
-\subsection{Results and Discussion}\label{results-and-discussion}}
+\section{Results and Discussion}\label{results-and-discussion}}
 
@@
 
-\caption{Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' category}\label{fig:psychosocialSupportAnalysis}
+\caption{Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' search category}\label{fig:psychosocialSupportAnalysis}
 \end{figure}
@@
 Good psychosocial support is also vital to improve both students' and teachers' mental health and well-being during a pandemic.
-This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two suddent spikes in the search trend series of the term `Psychosocial support' during the academic breaks in August 2020 and in August 20201 under `Education' category.
+This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two sudden spikes in the search volume index graphs of the term `Psychosocial support' which correspond well with the academic breaks in August in both years 2020 and 2021, under `Education' category.
 This highlight the urgent need of attention and solutions towards these aspects.
@@
 
-\hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
+\hypertarget{conclusions-and-further-work}{%
+\section{Conclusions and Further work}\label{conclusions-and-further-work}}
 
-The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
-The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane \& Galyen 2011).
+In this study we investigated whether the Google trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
+This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during covid 19 pandemic.
+A few previous published studies \autocite{vaughan2014web,kansal2021google} have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
 
-\url{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0248029}
+Our findings suggest that web search interest in distance learning solutions at the global was strongly influenced by the Covid-19 pandemic.
+Google trend search queries showed a strong cross correlation between covid-19 related search terms and distance education related search terms.
+The high popularity and increasing public attention towards different distance learning solutions during Covid-19 pandemic was also investigated through a detailed google trend footprint analysis.
+The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
+According to \textcite{wallace2003online}, the learning tools can have a significant impact on how students engage in online learning.
+The finding will also helpful for developers to identify their competitors in the market.
+High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the covid-19 pandemic.
+These findings offer important insights to various stakeholders of education to make better decision with less effort and time.
 
-First and second waves of coronavirus disease-19: A comparative study in hospitalized patients in Reus, Spain
+Google Trends only captures the search behaviour of a a sub population with Internet access, eventhough it is the most common worldwide search engine.
+Our study focused on global attention toward distance learning solutions during Covid-19 pandemic.
+Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.
+This can lead to variations in the search output and in turn the study findings for different regions and time frames.
+Further, Google Trends data were not available as absolute values and available only in the form of relative volume.
+Therfore the scores ranges from 0 to 100 do not allow the researchers to make meaningful comparisons across different education needs.
+Further investigation is needed at national and regional level to identify the specific measures that can be taken to ensure the quality of distance education at national and regional level.
+
+\hypertarget{supplementary-materials}{%
+\section*{Supplementary Materials}\label{supplementary-materials}}
+\addcontentsline{toc}{section}{Supplementary Materials}
+
+\textbf{Data and scripts}: Datasets and R code to reproduce all figures in this article (\url{https://github.com/pridiltal/Covid19_Online_Teaching_paper}).
+
+\hypertarget{acknowledgment}{%
+\section*{Acknowledgment}\label{acknowledgment}}
+\addcontentsline{toc}{section}{Acknowledgment}
+
+This work was carried out with the aid of a grant form UNESCO and the International Development Research Centre, Ottawa, Canada.
+The views expressed herein do not necessarily represent those of UNESCO, IDRC or its Board of Governors.
 

### 2021-12-26 main.Rmd
update paper

@@
 
-<!--
-The first confirmed case of Covid-19 was reported from Sri Lanka on 27 January 2020 with  a person who came to Sri Lanka as a tourist.
-The first locally transmitted COVID-19 case in Sri Lanka was then reported on  11 March 2020 followed by an increasing trend of the COVID-19 confirmed cases.
-In response to this increasing trend of COVID 19 cases, lockdowns and curfews were imposed several times across the country with the aim of reducing the spread of the virus.
-As of 29 September 2021, 515524 cases of COVID-19 have been reported in Sri Lanka, resulting with 12786 deaths [@HealthSL2020; @dong2020interactive].
--->
-
 ```{r covidImpactWorld, fig.cap = "Global tracking of COVID-19 caused school closures and re-openings", fig.height=4}
@@
 Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education].
-Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020.
+Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 [@unesco2020covid].
 With the start of the pandemic a near total closure of schools were observed all around the world.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools (Figures \ref{fig:covidImpactContinent}).
-
-<!--In line with that, the Sri Lankan  government also ordered to close schools from 12 March 2020 and  its planned reopening of schools was delayed several times due to unexpected Covid-19-related circumstances [@wikieducation].-->
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools where classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
 
@@
 The remainder of this paper is organized as follows.
-Section 2 presents the related work to lay the foundation for the Google footprint analysis.
+Section 2 presents the related work to lay the foundation for the Google trend footprint analysis of different distance learning solutions.
 Section 3 presents the methodology followed in the study.
-Section 4 includes Google footprint analysis.
+Section 4 includes results and discussion.
 Section 5 concludes the article and presents future research directions.
 
-## Background
+# Background
+
+The term distance education is known by a variety of names. The terms such as distance learning, e-Learning,  online learning environments, home study, independent study, external
+study, correspondence education, off campus study, open learning and open education  are used commonly and interchangeably  [@moore2011learning].
+It often describes the effort of providing access to learning for those who are geographically distant.
+The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
+Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places  and/or  varying forms of instructional materials [@moore2011learning].
+There are different types  of interactions that facilitate learners to construct knowledge in distance education environments.
+These are learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction [@wallace2003online].
 
@@
 
-## Methodology
+# Methodology
 
-In this work we use  Google Analytics search index series  as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
-We use weekly  Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
+To identify the impact of covid 19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO [@unesco2020covid].
+Data related to Covid-19 cases was retrieved from the R software [@rsoftware] package `coronavirus` [@coronavirusr].
+The  package pulls data from the Johns
+Hopkins University Center for Systems Science and Engineering (JHU CCSE) data repository [@dong2020interactive].
+
+In this work we also used  Google Analytics search index series  as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
+We used weekly  Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
 We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the  period of Covid-19 pandemic.
-Through this we want to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also study the correlation between weekly web searches and the number of weekly  cases of covid-19.
-
-There are different types  of interactions that facilitate learners to construct knowledge in distance education environments.
-These are learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction [@wallace2003online].
+Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
+We also studied the cross correlation between weekly web searches and the  weekly  cases of covid-19.
 
@@
 
-The term distance education is known by a variety of names. The terms such as distance learning, e-Learning,  online learning environments, home study, independent study, external
-study, correspondence education, off campus study, open learning and open education  are used commonly and interchangeably  [@moore2011learning].
-It often describes the effort of providing access to learning for those who are geographically distant.
-The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
-Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places  and/or  varying forms of instructional materials [@moore2011learning].
 It is important to decide what term to search for in Google Trends as different terms have different search volumes.
@@
 
-## Results and Discussion
+# Results and Discussion
 
@@
 
-```{r psychosocialSupportAnalysis, fig.height= 3, fig.cap="Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' category"}
+```{r psychosocialSupportAnalysis, fig.height= 3, fig.cap="Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' search category"}
 ```
@@
 Good psychosocial support is also vital to improve both students' and  teachers' mental health and well-being during a pandemic.
-This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two suddent spikes in the search trend series of  the term 'Psychosocial support' during the academic breaks in August 2020 and in August 20201 under 'Education' category.
+This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two sudden spikes in the search volume index graphs of the term 'Psychosocial support'  which correspond well with  the academic breaks in August in both years 2020 and 2021, under 'Education' category.
 This highlight the urgent need of attention and solutions towards these aspects.
@@
 
-## Conclusions
+# Conclusions and Further work
 
-The lack of consistency in terminology inevitably affects not only the researchers who would like to build upon the findings, but also impacts designers who are creating similar types of environments.
-The findings show great differences in the meaning of foundational terms that are used in the field, but also provide implications internationally for the referencing, sharing, and the collaboration of results detailed in varying research studies (Moore, Dickson-Deane & Galyen 2011).
+In this study  we investigated whether the Google  trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
+This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during covid 19 pandemic.
+A few previous published studies [@vaughan2014web; @kansal2021google] have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
 
-https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0248029
+Our findings suggest that web search interest in distance learning solutions at the global  was strongly influenced by the Covid-19 pandemic.
+Google trend search queries showed a strong cross correlation between covid-19 related search terms and distance education related search terms.
+The high popularity and increasing public attention towards different distance learning solutions during Covid-19 pandemic was also investigated through a detailed google trend footprint analysis.
+The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
+According to @wallace2003online, the learning tools can have a significant impact on how students engage in online learning.
+The finding will also helpful for developers to identify their competitors in the market.
+High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the covid-19 pandemic.
+These findings offer important insights to  various stakeholders of education  to make better decision with less effort and time.
 
-First and second waves of coronavirus disease-19: A comparative study in hospitalized patients in Reus, Spain
+Google Trends only captures the search behaviour of a a sub population with Internet access, eventhough it  is the most common worldwide search engine.
+Our study focused on global attention toward distance learning solutions during Covid-19 pandemic.
+Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.
+This can lead to variations in the search output and in turn the study findings for different regions and time frames.
+Further, Google Trends data were not available as absolute values and available only in the form of relative volume.
+Therfore the scores ranges from 0 to 100 do not allow the researchers to make meaningful comparisons across different education needs.
+Further investigation is needed at national and regional level to identify the specific measures that can be taken to ensure the quality of distance education at national and regional level.
+
+# Supplementary Materials {-}
+
+**Data and scripts**: Datasets and R code to reproduce all figures in this article (https://github.com/pridiltal/Covid19_Online_Teaching_paper).
+
+# Acknowledgment {-}
+
+This work was carried out with the aid of a grant form UNESCO and the International Development Research Centre, Ottawa, Canada.
+The views expressed herein do not necessarily represent those of UNESCO, IDRC or its Board of Governors.

### 2021-12-26 src/main.R

@@ -1236,145 +1236,3 @@ p11 <- dataW %>%
 
 print(p11)
 
-
-## ---- other
-
-keywords= c("Google Classroom - Topic", 
-            "Learning management system - Topic",
-            "Moodle - Topic",
-            "Virtual learning environment",
-            "zoom")
-# filter only  education
-category <- categories[categories$name=="Education",2] %>% as.numeric()
-#set the geographic area: DE = Germany
-country=c('LK')  # use this to find country code https://en.wikipedia.org/wiki/ISO_3166-2
-#set the time window
-time=("2019-01-01 2021-08-15")
-#set channels 
-channel='web'
-trends = gtrends(keywords, gprop =channel,geo=country, time = time, category = category )
-
-
-
-
-
-
-## ---- extract_data_most_popular
-## Most popular quiries in education 
-
-# set category- use 'data(categories)' to retreive valid codes. 74- Education
-category <- categories[categories$name=="Education",2] %>% as.numeric()
-#set the geographic area: DE = Germany
-# use this to find country code https://en.wikipedia.org/wiki/ISO_3166-2
-country <- c('LK') 
-#set the time window
-# First Sri Lankan found on 3 March 2020
-time <- ("2020-03-03 2021-08-15")
-#set channels 
-channel <- 'web'
-trends <- gtrends( gprop =channel,geo=country, time = time , category = category)
-#select only releted topics
-related_topics_edu <- trends$related_topics
-# filter only rising values
-related_topics_edu %>% filter(related_topics == "rising" ) %>% 
-  select(value) %>% unique()
-
-
-
-## ---- extract_data_distance_learning
-## MAnusally selected the followin ketwords related to online education in Sri Lanka
-keywords= c("Google Classroom", 
-            "Learning management system",
-            "Moodle",
-            "Virtual learning environment",
-            "zoom")
-# filter only  education category
-category <- categories[categories$name=="Education",2] %>% as.numeric()
-#set the geographic area: DE = Germany
-country=c('LK')  # use this to find country code https://en.wikipedia.org/wiki/ISO_3166-2
-#set the time window
[…]

### 2021-12-26 Paper_online_Learning.Rmd
update introduction

--- 2021-12-16 Paper_online_Learning.Rmd
@@
   othernames: Thiyanga S.
-  address: Department of Statistics, University of Sri Jayewardenepura, Sri Lanka
-  email: tstalagala@gmail.com
-abstract: "A brief summary of our work"
-keywords: "rmarkdown, templates"
+  address: Department of Statistics, Faculty of Applied Sciences, University of Sri Jayewardenepura, Sri Lanka
+  email: ttalagala@sjp.ac.lk
+abstract: "Distance Education has a long history.
+However COVID-19 has made a new era of distance education.
+With the increasing demand, various different distance learning solutions have been introduced for different education purposes.
+In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions.
+Both visual and analytical approaches were used to  analyse global level web search querie during COVID-19 pandemic.
+The findings of this study include a Google Trend footprint of various distance leraning solutions for diffrent education purposes that can be used as a proxy for measuring global attention towards different distance learning solutions.
+The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes.
+R code and data to reproduce the results of this work are available in the online supplementary materials."
+keywords:  "Online-learning, Distance education solutions, COVID-19 Pandemic,  Google Trend search queries "
 wpnumber: 1
-jelcodes: C10,C14,C22
+jelcodes: C82,I20,I21
 blind: false

### 2021-12-26 Paper_online_Learning.tex
update introduction

@@
             pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
-            pdfkeywords={Online-learning, Distance education solutions, Covid-19 Pandemic, Google trend search queries},
+            pdfkeywords={Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries},
             pdfborder={0 0 0},
@@
 \begin{abstract}
-Distance Education has long history.
-However Coid-19 has made a new era of distance education.
-With the increasing demand, many different distance learning solutions have been introduced for different education purposes.
-In this study we invetigated the impact of Covid 19 on ditance education and the possibility of using Google analytics search index series as a proxy to quantify the popularity and the public interest for different distance education solutions.
-Both visual and analytical approaches were used to analyse web search queries at the global level during Covid-19 pandemic.This study intorduced a Google trend footprint of various ditance leraning solutions for diffrent education purposes, that can be used as a proxy for measuring global attension towards different learning solutions.
-The findings provide a fast first step to identify the most popular tools available for different educational purposes.
-R code and data are available in the online supplementary materials.
+Distance Education has a long history.
+However COVID-19 has made a new era of distance education.
+With the increasing demand, various different distance learning solutions have been introduced for different education purposes.
+In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the popularity and the public interest towards different distance education solutions.
+Both visual and analytical approaches were used to analyse global level web search querie during COVID-19 pandemic.
+The findings of this study include a Google Trend footprint of various distance leraning solutions for diffrent education purposes that can be used as a proxy for measuring global attention towards different distance learning solutions.
+The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes.
+R code and data to reproduce the results of this work are available in the online supplementary materials.
 \end{abstract}
 \begin{keywords}
-Online-learning, Distance education solutions, Covid-19 Pandemic, Google trend search queries
+Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries
 \end{keywords}
@@
 
-The ongoing pandemic of COVID-19 is still one of the most important priorities of governments and media of many countries all around the world.
-Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}.
+COVID-19 is still one of the most important priorities of governments and media of many countries all around the world.
+Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}.
 As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}.
@@
 
-The ongoing COVID-19 pandemic has brought the world to a standstill.
-Many sectors have been affected ever since the outbreak of COVID 19, worldwide.
+Many sectors have been affected ever since the outbreak of COVID-19, worldwide.
 Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}.
@@
 With the start of the pandemic a near total closure of schools were observed all around the world.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools where classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
 
@@
 Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
-alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}.
-Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.
-Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.
+alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}.
+Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.
+Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.
 
-The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders: teachers, students and parents.
-Teachers were not fully aware of their obligations and how to maintain connections with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID 19 pandemic.
+The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as teachers, students and parents.
+Teachers were not fully aware of their obligations and how to connect with students to support thier learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID-19 pandemic.
 Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services, long-term educational disengagement, digital exclusion, poor technology management, increased psychosocial challenges, decreases in mathematical achievement, a lack of engagement, increased early marriages , psychosocial challenges \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}.
+Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}.
 This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home.
@@
 
+According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.
 In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure.
@@
 Yet very little is known about the uptake of different solutions or about their effectiveness.
-The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content \autocite{moore2011learning}.
-However, getting familiar with all these available distance solutions is not feasible or practical due to limited time availability and need of urgent responses.
-According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.
+Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and need of urgent responses.
 
-Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting distance education solutions.
-Further, according to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as popularity can indicate the prominence of the product in the market, its usability and its impact.
+Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting suitable distance education solutions.
+According to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as it can indicate the prominence of the product in the market, its usability and its impact.
 Further, according to \textcite{willis2020using}, developers also tend to improve their products and services in response to increasing online search.
-Therefore, the internet represents a great opportunity to learn about public attention towards a product or service.
+Therefore, the internet represents a great opportunity to learn about the popularity and public attention towards a product or service.
 It is also a way to narrow down the search space and thereby identify most suitable product and services that meets customer needs.
 
-Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior \autocite{nuti2014use}.
-As there is still the main struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
-In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID 19 pandemic on education?
-(2) What solutions are in place to meet the need of distance education in terms during COVID 19 pandemic?
+Google Trends is an open source web analytics tool that allows users to get access to Internet search data that has the ability to provide deep insights into population behavior \autocite{nuti2014use}.
+As there is still a struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Trends search quiries can be used as a proxy of the popularity and the public interest in different distance education solutions.
+In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID-19 pandemic on education?
+(2) What solutions are in place to meet the need of distance education in terms during COVID-19 pandemic?
 (3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?
-We primarily analyze quantitative digital footprint data on the internet from December 2019 to August 2021.
-The resulted google trend footprint provides a fast first step to identify the most popular distance education tools available for different education purposes.
+
+We primarily analyzed quantitative digital footprint data on the internet from December 2019 to August 2021.
+The resulted Google Trend footprint provides a fast first step guide to identify the most popular distance education tools available for different education purposes.
 It also allows the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching.
@@
 Quality education is essential to sustainable development and it is a key United Nations (UN) sustainable development goal.
-Having the right tool is imperative to successful completion of a task at hand.
-The findings of this study provides an initial guidance to select a right tool.
+Having the most appropriate tool is imperative to successful completion of a task at hand.
+The findings of this study provides an essential initial guidance to select the most appropriate tool for different distance education needs.
 Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.
@@
 
-To identify the impact of covid 19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO \autocite{unesco2020covid}.
-Data related to Covid-19 cases was retrieved from the R software \autocite{rsoftware} package \texttt{coronavirus} \autocite{coronavirusr}.
+To identify the impact of COVID-19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO \autocite{unesco2020covid}.
+Data related to COVID-19 cases was retrieved from the R software \autocite{rsoftware} package \texttt{coronavirus} \autocite{coronavirusr}.
 The package pulls data from the Johns
@@
 In this work we also used Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
-We used weekly Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
-We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of Covid-19 pandemic.
+We used weekly Google Analytics search index series of various distance learning solutions during COVID-19 pandemic.
+We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of COVID-19 pandemic.
 Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also studied the cross correlation between weekly web searches and the weekly cases of covid-19.
+We also studied the cross correlation between weekly web searches and the weekly cases of COVID-19.
 
@@
 Google Trends allows to search for up to five queries each time \autocite{vaughan2014web}.
-We did our analysis under 11 sub segmentations namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
-The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to Covid 19 pandemic.
+We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
+The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.
 
@@
 
-Both visual and analytical approaches were used to analyse the changes in web search queries at the global level over the entire period of the Covid-19 pandemic to identify the emerging evidence about the impact of COVID-19 pandemic on distance education.
-We also evaluated the relationship between weekly web searches and the number of weekly global cases of covid-19.
-Cross correlation analysis and the dynamic time warping algorithm \autocite{giorgino2009computing} were performed to investigate the relationship and similarities between the signatures of covid-19 related search volume data and distance education related search volume data.
+Both visual and analytical approaches were used to analyse the changes in web search queries at the global level over the entire period of the COVID-19 pandemic to identify the emerging evidence about the impact of COVID-19 pandemic on distance education.
+We also evaluated the relationship between weekly web searches and the number of weekly global cases of COVID-19.
+Cross correlation analysis and the dynamic time warping algorithm \autocite{giorgino2009computing} were performed to investigate the relationship and similarities between the signatures of COVID-19 related search volume data and distance education related search volume data.
 
@@
 
-Search volume index graphs for Covid 19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic.
+Search volume index graphs for COVID-19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic.
 Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic.
@@
 However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``never version'' and ``improved version'' of distance learning.
-The google search trend pattern corresponds to the term ``online learning'' (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout Covid-19 pandemic.
+The google search trend pattern corresponds to the term ``online learning'' (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout COVID-19 pandemic.
 Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``Distance learning'' and ``online teaching'' demonstrate a similar trend during the study period.
-Covid-19 related web searches declined in the months after the WHO announcement of the Covid-19 pandemic, despite the increase in Covid-19 cases.
+COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases.
 A similar decline could be detected in distance education related terms.
-further, a second was observed only in education related terms after the WHO announcement of the Covid-19 pandemic.
-This could be due to the near total closures of schools happened with the declaration of Covid-19 as a pandemic.
+further, a second was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
+This could be due to the near total closures of schools happened with the declaration of COVID-19 as a pandemic.
 
@@
 The distance between the two series is then computed by summing the distances of individual aligned elements \autocite{giorgino2009computing}.
-Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact from covid 19 on distance education.
+Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact from COVID-19 on distance education.
 Both pairwise comparison show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)).
@@
 
-The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to covid 19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after covid 19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
 
@@
 By the time it was know as ``corona'', it was not a global concern and was mostly limited to China.
-During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to Covid 19 pandemic all around the world.
+During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic all around the world.
 This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis.
@@
 With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
-With this sudden unexpected movement, online learning also became an equally important global concern, similar to covid 19 pandemic.
+With this sudden unexpected movement, online learning also became an equally important global concern, similar to COVID-19 pandemic.
 This similar behaviour in public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line from approximately 120 to 150 on the Query Index axis.
-This also confirms the impact of covid 19 on distance education and online learning.
+This also confirms the impact of COVID-19 on distance education and online learning.
 
-Then, a cross correlation analysis was performed to estimate the lag between covid 19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
-A close relationship was evident between covid related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''.
-The significant cross-correlation coefficients at lag zero confirms contemporaneous relationship between the covid related search terms and distance learning related search terms.
-The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
-However, no significant synchronization was observed between the Covid 19 related search trends and the search patterns of ``distance education'' or ``online proctoring'' (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
+Then, a cross correlation analysis was performed to estimate the lag between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
+A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''.
+The significant cross-correlation coefficients at lag zero confirms contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms.
+The significant cross correlation values at lag one or two also confirms the sudden impact of COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
+However, no significant synchronization was observed between the COVID-19 related search trends and the search patterns of ``distance education'' or ``online proctoring'' (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
 
@@
 
-\caption{Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic}\label{fig:ccfAnalysis}
+\caption{Cross-correlation analysis between distance learning related online search terms and COVID-19 relate online search terms during COVID-19 pandemic}\label{fig:ccfAnalysis}
 \end{figure}
@@
 
-Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during covid 19 pandemic (Figure \ref{fig:plot1} (a)).
+Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during COVID-19 pandemic (Figure \ref{fig:plot1} (a)).
 This might be due to it's easy and fast setup and usage \autocite{sudarsana2019use}.
@@
 A significant peak in `Ubongo' series can be observed during 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
-Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during covid-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
-The can be a huge relief for financially distressed families during Covid-19 pandemic.
+Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+The can be a huge relief for financially distressed families during COVID-19 pandemic.
 Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
@@
 The second highest option only provides support for language learning.
-`Quizlet'`s ability provide support for multiple subject might be the reason for it to outperform it competitor in the market.
-Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
+Quizlet's ability provide support for multiple subject might be the reason for it to outperform it competitor in the market.
+Under mobile reading applications, `Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
 Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
@@
 At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
-These peaks aligns with the near-total closures of schools due to covid 19 in Figure \ref{fig:covidImpactWorld}.
+These peaks aligns with the near-total closures of schools due to COVID-19 in Figure \ref{fig:covidImpactWorld}.
 The major concerns that arise with the unexpected sudden movement from physical classroom to virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
 Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
-The inability to reopen school fully due to the increasing Covid-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.
+The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.
 
@@
 The attention in the education sector is low compared to all category.
-However a significant peak can be observed during the first academic break after declaring Covid 19 as a pandemic.
+However a significant peak can be observed during the first academic break after declaring COVID-19 as a pandemic.
 Similar peak can be observed in 2021 .
@@
 In this study we investigated whether the Google trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
-This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during covid 19 pandemic.
+This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic.
 A few previous published studies \autocite{vaughan2014web,kansal2021google} have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
 
-Our findings suggest that web search interest in distance learning solutions at the global was strongly influenced by the Covid-19 pandemic.
-Google trend search queries showed a strong cross correlation between covid-19 related search terms and distance education related search terms.
-The high popularity and increasing public attention towards different distance learning solutions during Covid-19 pandemic was also investigated through a detailed google trend footprint analysis.
+Our findings suggest that web search interest in distance learning solutions at the global was strongly influenced by the COVID-19 pandemic.
+Google trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
+The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed google trend footprint analysis.
 The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
@@
 The finding will also helpful for developers to identify their competitors in the market.
-High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the covid-19 pandemic.
+High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the COVID-19 pandemic.
 These findings offer important insights to various stakeholders of education to make better decision with less effort and time.
 
-Google Trends only captures the search behaviour of a a sub population with Internet access, eventhough it is the most common worldwide search engine.
-Our study focused on global attention toward distance learning solutions during Covid-19 pandemic.
+Google Trends only captures the search behaviour of a sub population with Internet access, eventhough it is the most common worldwide search engine.
+Our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
 Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.

### 2021-12-26 main.Rmd
update introduction

@@
 
-The ongoing pandemic of COVID-19 is still one of the most important priorities of governments and  media of many countries all around the world.
-Due to the alarming levels of spread and severity, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline].
+COVID-19 is still one of the most important priorities of governments and  media of many countries all around the world.
+Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline].
 As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].
@@
 
-The ongoing COVID-19 pandemic has brought the world to a standstill.
-Many sectors have been affected ever since the outbreak of COVID 19, worldwide.
+Many sectors have been affected ever since the outbreak of COVID-19, worldwide.
 Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education].
@@
 With the start of the pandemic a near total closure of schools were observed all around the world.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools where classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
+
+```{r distance_learning_world, eval=FALSE}
+# Do not run
+# Download Data
+```
+
+```{r covid19_trend_world, eval=FALSE}
+# Do not run
+# Download Data
+```
+
+```{r covid_epidemic_datasave, eval=FALSE}
+# Do not run
+# Download Data
+```
 
@@
 Eventhough distance education has a long history that goes back to almost two centuries with significant  modifications,
-alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID 19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously [@richmond2020critical].
-Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID 19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
-Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing Covid-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to Covid-19 pandemic.
+alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously [@richmond2020critical].
+Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
+Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.
 
-The speed of the physical classroom closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders: teachers, students and parents.
-Teachers were not fully aware of their obligations and how to maintain connections with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID 19 pandemic.
+The speed of the physical classroom closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as  teachers, students and parents.
+Teachers were not fully aware of their obligations and how to connect with students to support thier learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID-19 pandemic.
 Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services, long-term educational disengagement, digital exclusion, poor technology management, increased psychosocial challenges, decreases in mathematical achievement, a lack of engagement, increased early marriages ,  psychosocial challenges [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher].
+Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher].
 This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home.
@@
 
+According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught.
 In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure.
@@
 Yet very little is known about the uptake of different solutions or about their effectiveness.
-The design of different types of distance learning solutions can depend on the learning objective, target audience, access, and type of content [@moore2011learning].
-However, getting familiar with all these available distance solutions  is  not feasible or practical due to limited time availability and need of urgent responses.
-According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught.
+Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and need of urgent responses.
 
-Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting distance education solutions.
-Further, according to @ahn2006utilizing, the popularity of a product can also greatly influence consumer purchasing decisions as popularity can  indicate the prominence of the product in the market, its usability and its impact.
+Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting suitable distance education solutions.
+According to @ahn2006utilizing, the popularity of a product can also greatly influence consumer purchasing decisions as it can  indicate the prominence of the product in the market, its usability and its impact.
 Further, according to @willis2020using, developers also tend to improve their products and services in response to increasing online search.
-Therefore, the internet represents a great opportunity to learn about public attention towards a product or service.
+Therefore, the internet represents a great opportunity to learn about the popularity and public attention towards a product or service.
 It is also a way to narrow down the search space and  thereby identify most suitable product and services that meets customer needs.
 
-Google Trends is an open source web analytics tool that allows users to interact with Internet search data, which can provide deep insights into population behavior [@nuti2014use].
-As there is still the main struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Analytics search index series can be used as a proxy of the popularity and the public interest in different distance education solutions.
-In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID 19 pandemic on education?
-(2) What solutions are in place to meet the need of distance education in terms during COVID 19 pandemic?
+Google Trends is an open source web analytics tool that allows users to get access to Internet search data that has the ability to provide deep insights into population behavior [@nuti2014use].
+As there is still a struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Trends search quiries  can be used as a proxy of the popularity and the public interest in different distance education solutions.
+In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID-19 pandemic on education?
+(2) What solutions are in place to meet the need of distance education in terms during COVID-19 pandemic?
 (3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?
-We primarily analyze quantitative digital footprint data on the internet from December 2019 to August 2021.
-The resulted google trend footprint provides a fast first step to identify the most popular distance education tools available for different education purposes.
+
+We primarily analyzed quantitative digital footprint data on the internet from December 2019 to August 2021.
+The resulted Google Trend footprint provides a fast first step guide to identify the most popular distance education tools available for different education purposes.
 It  also allows the teachers to narrow down the search space and  deepen their exploration on  prominent distance education solutions to support their online teaching.
@@
 Quality education is essential to sustainable development and it is a key  United Nations (UN) sustainable development goal.
-Having the right tool is imperative to successful completion of a task at hand.
-The findings of this study provides an initial guidance to select a right tool.
+Having the most appropriate tool is imperative to successful completion of a task at hand.
+The findings of this study provides an essential initial guidance to select the most appropriate tool for different distance education needs.
 Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.
@@
 
-To identify the impact of covid 19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO [@unesco2020covid].
-Data related to Covid-19 cases was retrieved from the R software [@rsoftware] package `coronavirus` [@coronavirusr].
+To identify the impact of COVID-19 on education, data on the evolution of school closures and re-openings was obtained from a repository maintained by UNESCO [@unesco2020covid].
+Data related to COVID-19 cases was retrieved from the R software [@rsoftware] package `coronavirus` [@coronavirusr].
 The  package pulls data from the Johns
@@
 In this work we also used  Google Analytics search index series  as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
-We used weekly  Google Analytics search index series of various distance learning solutions during covid-19 pandemic.
-We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the  period of Covid-19 pandemic.
+We used weekly  Google Analytics search index series of various distance learning solutions during COVID-19 pandemic.
+We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the  period of COVID-19 pandemic.
 Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also studied the cross correlation between weekly web searches and the  weekly  cases of covid-19.
+We also studied the cross correlation between weekly web searches and the  weekly  cases of COVID-19.
 
@@
 
-<!-- completed up to page 30-->
-
 Google Trends allows to search for up to five queries each time [@vaughan2014web].
-We did our analysis under 11 sub segmentations namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
-The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to Covid 19 pandemic.
+We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
+The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.
 
@@
 
-Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level over the entire period of the Covid-19 pandemic to identify the emerging evidence about the impact of  COVID-19 pandemic on distance education.
-We also evaluated the  relationship between weekly web searches and the number of  weekly global cases of covid-19.
-Cross correlation analysis and  the dynamic time warping algorithm [@giorgino2009computing] were performed  to investigate the relationship and similarities between the signatures of  covid-19 related search volume data  and distance education related search volume data.
+Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level over the entire period of the COVID-19 pandemic to identify the emerging evidence about the impact of  COVID-19 pandemic on distance education.
+We also evaluated the  relationship between weekly web searches and the number of  weekly global cases of COVID-19.
+Cross correlation analysis and  the dynamic time warping algorithm [@giorgino2009computing] were performed  to investigate the relationship and similarities between the signatures of  COVID-19 related search volume data  and distance education related search volume data.
 
@@
 However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version", "never version" and "improved version" of distance learning.
-The google search trend pattern corresponds to the term "online learning"  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout  Covid-19 pandemic.
+The google search trend pattern corresponds to the term "online learning"  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout  COVID-19 pandemic.
 Despite quantitative differences in terms of relative search volume, the series generated by the two terms "Distance learning" and "online teaching" demonstrate a similar trend during the study period.
-Covid-19 related web searches declined in the months after the WHO announcement of the Covid-19 pandemic, despite the increase in Covid-19 cases.
+COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases.
 A similar decline could be detected in distance education related terms.
-further,  a second was observed only in education related terms  after the WHO announcement of the Covid-19 pandemic.
-This could be due to the near total closures of schools happened  with the declaration of  Covid-19 as a pandemic.
+further,  a second was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
+This could be due to the near total closures of schools happened  with the declaration of  COVID-19 as a pandemic.
 
@@
 The distance between the two series is then computed by summing the distances of individual aligned elements [@giorgino2009computing].
-Two pairwise comparisons, "corona"  with "online learning" and   "covid" with "online learning" were conducted to see the impact  from covid 19 on distance education.
+Two pairwise comparisons, "corona"  with "online learning" and   "covid" with "online learning" were conducted to see the impact  from COVID-19 on distance education.
 Both pairwise comparison show a significant  areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and  \ref{fig:dtw2} (a)).
@@
 
-The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to covid 19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after covid 19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
 
@@
 By the time it was know as "corona", it was not a global concern and was mostly limited to China.
-During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to Covid 19 pandemic all around the world.
+During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic all around the world.
 This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis.
@@
 With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
-With this sudden unexpected movement, online learning  also became an equally important global concern, similar to covid 19 pandemic.
+With this sudden unexpected movement, online learning  also became an equally important global concern, similar to COVID-19 pandemic.
 This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line  from approximately 120 to 150 on the Query Index axis.
-This also confirms the impact of covid 19 on distance education and online learning.
+This also confirms the impact of COVID-19 on distance education and online learning.
 
-Then, a cross correlation analysis was performed to estimate the lag between covid 19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
-A close relationship was evident between covid related search patterns and the search patterns of the terms "Online learning", "Distance learning" and "Online teaching".
-The significant cross-correlation coefficients  at lag zero confirms  contemporaneous relationship between the covid related search terms and distance learning related search terms.
-The significant cross correlation values at lag one or two also confirms the sudden impact of covid 19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
-However, no significant  synchronization was observed  between the Covid 19 related search trends and the search patterns of "distance education" or "online proctoring" (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
+Then, a cross correlation analysis was performed to estimate the lag between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
+A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms "Online learning", "Distance learning" and "Online teaching".
+The significant cross-correlation coefficients  at lag zero confirms  contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms.
+The significant cross correlation values at lag one or two also confirms the sudden impact of COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
+However, no significant  synchronization was observed  between the COVID-19 related search trends and the search patterns of "distance education" or "online proctoring" (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).
 
-```{r ccfAnalysis, fig.height=7, fig.cap= "Cross-correlation analysis between distance learning related online search terms and covid 19 relate online search terms during Covid-19 pandemic"}
+```{r ccfAnalysis, fig.height=7, fig.cap= "Cross-correlation analysis between distance learning related online search terms and COVID-19 relate online search terms during COVID-19 pandemic"}
 ```
@@
 
-Under digital learning management systems "Google Classroom", which  helps  teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during covid 19 pandemic (Figure \ref{fig:plot1} (a)).
+Under digital learning management systems "Google Classroom", which  helps  teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during COVID-19 pandemic (Figure \ref{fig:plot1} (a)).
 This might be due to it's easy and fast setup and usage [@sudarsana2019use].
@@
 A significant peak in  'Ubongo' series can be observed during 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
-Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during covid-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
-The can be a huge relief for financially distressed families during Covid-19 pandemic.
+Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+The can be a huge relief for financially distressed families during COVID-19 pandemic.
 Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
@@
 The second highest option only provides  support for language learning.
-'Quizlet''s ability provide support for multiple subject might be the reason for it to outperform it  competitor in the market.
+Quizlet's ability provide support for multiple subject might be the reason for it to outperform it  competitor in the market.
 Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
@@
 At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a))  and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
-These peaks aligns with the near-total closures of schools due to covid 19 in Figure \ref{fig:covidImpactWorld}.
+These peaks aligns with the near-total closures of schools due to COVID-19 in Figure \ref{fig:covidImpactWorld}.
 The major concerns that arise with the unexpected sudden movement from physical classroom to  virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
 Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
-The inability to reopen school fully due to the increasing Covid-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reason for this considerable public attention around August, 2020.
+The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reason for this considerable public attention around August, 2020.
 
 ```{r Digital_lms_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r mobile_phones_apps_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r offline_functionality_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r mooc_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r Self_directed_learning_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r Mobile_reading_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r Collaboration_platforms_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r Tools_for_teachers_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r External_repositories_DL_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r Online_proctoring_data_download, eval=FALSE}
+# Do not run
+# Download Data
 ```
@@
 ```{r Online_proctoring_analysis, dependson="External_repositories_DL_analysis"}
-```
-
-```{r psychosocial_support_data_download, eval=FALSE}
 ```
@@
 The attention in the education sector is low compared to all category.
-However a significant peak can be observed during the first academic break  after declaring  Covid 19 as a pandemic.
+However a significant peak can be observed during the first academic break  after declaring  COVID-19 as a pandemic.
 Similar peak can be observed in 2021 .
 But that significant compared to 2020 peak.
+
+```{r psychosocial_support_data_download, eval=FALSE}
+# Do not run
+# Download Data
+```
 
@@
 In this study  we investigated whether the Google  trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
-This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during covid 19 pandemic.
+This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic.
 A few previous published studies [@vaughan2014web; @kansal2021google] have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
 
-Our findings suggest that web search interest in distance learning solutions at the global  was strongly influenced by the Covid-19 pandemic.
-Google trend search queries showed a strong cross correlation between covid-19 related search terms and distance education related search terms.
-The high popularity and increasing public attention towards different distance learning solutions during Covid-19 pandemic was also investigated through a detailed google trend footprint analysis.
+Our findings suggest that web search interest in distance learning solutions at the global  was strongly influenced by the COVID-19 pandemic.
+Google trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
+The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed google trend footprint analysis.
 The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
@@
 The finding will also helpful for developers to identify their competitors in the market.
-High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the covid-19 pandemic.
+High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the COVID-19 pandemic.
 These findings offer important insights to  various stakeholders of education  to make better decision with less effort and time.
 
-Google Trends only captures the search behaviour of a a sub population with Internet access, eventhough it  is the most common worldwide search engine.
-Our study focused on global attention toward distance learning solutions during Covid-19 pandemic.
+Google Trends only captures the search behaviour of a sub population with Internet access, eventhough it  is the most common worldwide search engine.
+Our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
 Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.

### 2021-12-26 src/main.R

@@ -12,14 +12,12 @@ library(countrycode)
 library(coronavirus)
 library(patchwork)
 #set the time window
-# Consider 2020 
 time <- ("2019-12-01 2021-08-15")
 
 ## ---- covidImpactWorld
 ## COVID-19 Impact on Education (global data)
 ## Data source: https://en.unesco.org/covid19/educationresponse
 ## Data download: 16-08-2021
-## Write (download from the above link)
 world_impact <- read.csv(
   here::here("data", "covid_impact_education_full.csv"))
 wimpact <- world_impact %>%
@@ -34,7 +32,6 @@ wimpact <- wimpact %>% select(-ISO) %>%
 p <- wimpact %>% autoplot(Count) +
   geom_line(size=1) +
   xlab ("Time") +
-  # viridis::scale_colour_viridis(discrete = TRUE)+
   scale_color_brewer(palette="Dark2")+
   ylab("Number of Countries")+
   theme_bw() + 
@@ -87,7 +84,6 @@ p <- wimpact %>%
   facet_grid(vars(Continent), scales = "free_y") +   # only assigns the row variable
   xlab("Time)") +
   scale_color_brewer(palette="Dark2")+
-  #viridis::scale_colour_viridis(discrete = TRUE, begin = 0, end = 1)+
   theme_bw() + 
   theme( text = element_text(size = 14), legend.position = "bottom")+
   guides(colour = guide_legend(nrow = 2))+
@@ -128,18 +124,6 @@ write.csv(trends, here::here( "data",
   "covid_search_world.csv"))
 
 
-## ---- covid19_vaccine_trend_world
-## Offline data retrieval
-keywords= c("COVID-19 vaccine", "covid vaccine")
-#set channels 
-channel <- 'web'
-trends <-  gtrends(keywords, gprop =channel,
-                   time = "2017-12-01 2021-08-15" )
-trends <- trends$interest_over_time
-write.csv(trends, here::here("data",
-  "covid_vaccine_search_world.csv"))
-
-
 ## ---- covid_epidemic_datasave
 # Data retrieved from coronavirus R package on 19-01-2021
 library(coronavirus)
@@ -156,6 +140,21 @@ covid_cases <- coronavirus %>%
 covid_cases <- covid_cases %>% as_tsibble(index = date)
 write.csv(covid_cases, here::here(
   "data", "covid_cases.csv"))
+
+
+
+## ---- covid19_vaccine_trend_world
[…]

### 2021-12-26 Paper_online_Learning.tex
edit typo

@@
 
-The term distance education is known by a variety of names. The terms such as distance learning, e-Learning, online learning environments, home study, independent study, external
-study, correspondence education, off campus study, open learning and open education are used commonly and interchangeably \autocite{moore2011learning}.
-It often describes the effort of providing access to learning for those who are geographically distant.
-The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
-Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places and/or varying forms of instructional materials \autocite{moore2011learning}.
-There are different types of interactions that facilitate learners to construct knowledge in distance education environments.
-These are learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction \autocite{wallace2003online}.
+The term distance education is known by a variety of names such as distance learning, e-Learning, online learning, home study, independent study, external study, correspondence education, off campus study, open learning and open education \autocite{moore2011learning}.
+Few common features found in all these terms are, they all involve some form knowledge sharing between two parties; learners and educators, different times, different geographical locations and/or varying forms of instructional materials and tools \autocite{moore2011learning}.
+Further, there are different types of interactions such are learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction that can facilitate learners to construct knowledge in distance education environments.
+These \autocite{wallace2003online}.
 

### 2021-12-26 main.Rmd
edit typo

@@
 
-The term distance education is known by a variety of names. The terms such as distance learning, e-Learning,  online learning environments, home study, independent study, external
-study, correspondence education, off campus study, open learning and open education  are used commonly and interchangeably  [@moore2011learning].
-It often describes the effort of providing access to learning for those who are geographically distant.
-The definition also stated that distance education uses emerging media and associated experiences to produce distributed learning opportunities.
-Few common features found in all these terms is they all involve some form of instruction occurs between two parties; learners and teachers, different times, places  and/or  varying forms of instructional materials [@moore2011learning].
-There are different types  of interactions that facilitate learners to construct knowledge in distance education environments.
-These are learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction [@wallace2003online].
+The term distance education is known by a variety of names such as distance learning, e-Learning,  online learning, home study, independent study, external study, correspondence education, off campus study, open learning and open education [@moore2011learning].
+Few common features found in all these terms are, they all involve some form knowledge sharing between two parties; learners and educators, different times, different geographical locations  and/or  varying forms of instructional materials and tools [@moore2011learning].
+Further, there are different types  of interactions such are learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction that can facilitate learners to construct knowledge in distance education environments.
+These [@wallace2003online].
 

### 2021-12-26 Paper_online_Learning.tex
updated methodology

@@
 
-Google Trends, is a freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages \autocite{carneiro2009google}.
+Google Trends is a freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages \autocite{carneiro2009google}.
 Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution \autocite{alicino2015assessing,cook2011assessing}.
 According to \textcite{jarynowski2020perception} the public concerns on different matters can also take an epidemic nature starting with the phase of growing interest which is known as ``early adoption'', to the phase of general interest which is known as ``majority'' to eventually lose popularity which is known as ``lagers stage''.
-Figure \ref{fig:distanceLearningWorldAnalysis} also confirms this life cycle explanation.
+Figure \ref{fig:distanceLearningWorldAnalysis} (b) and (c) also confirm this life cycle explanation of Google Trend search patterns.
 
-The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding} and health care \autocite{nuti2014use,alicino2015assessing,arora2019google,carneiro2009google,cook2011assessing} as it provides data about social phenomena in a more timely manner than traditional data collection processes \autocite{vaughan2014web}.
+The use of Google Trend search queries has become a popular data source in monitoring and modeling the dynamics of different application domains such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding} and health care \autocite{nuti2014use,alicino2015assessing,arora2019google,carneiro2009google,cook2011assessing} as it has the ability to provide data about social phenomena in a more timely manner than traditional data collection processes \autocite{vaughan2014web}.
 According to the study conducted by \textcite{willis2020using} on taxi preferences found that online queries can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
@@
 The study conducted by \textcite{husnayain2020applications} focuses on search terms related to the coronavirus, handwashing, and face masks to monitor public restlessness toward COVID-19.
-However, there are only a limited number of research attempts related to education section using Google trend search query data \autocite{vaughan2014web}.
-The study conducted by \textcite{vaughan2014web} have used google trend search queries to predict academic fame by testing the correlation between search volume data and university ranking data.
+However, there are only a limited number of research attempts related to education sector using Google Trend search query data \autocite{vaughan2014web}.
+The study conducted by \textcite{vaughan2014web} have used Google Trend search queries to predict academic fame by testing the correlation between search volume data and university ranking data.
 \textcite{kansal2021google} have used a text mining approach to identify emerging patterns of words and phrases during COVID-19 pandemic, keeping education prospects as the focus.
 However, their study was limited to a small number of learning platforms.
-However, according to \textcite{moore2011learning} the design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
+According to \textcite{moore2011learning}, the design of different types of learning environments can depend on the learning outcomes, target audience, access (physical, virtual and/or both), and/or type of content.
 Therefore, this study provides a detailed investigation on distance education solutions focusing on different distance learning needs.
-In contrast to the work in \textcite{kansal2021google}, the work in our paper has a very specific focus: the distance learning solutions for various education purposes.
+In contrast to the work in \textcite{kansal2021google}, the work in our paper has a very specific focus and provides a thorough investigation on different distance learning solutions available for various education purposes.
 
-Eventhough google trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool.
-It may contain inaccuracies \autocite{carneiro2009google}.
-According to \textcite{nuti2014use} the data retrieval process should be transparent.
+Eventhough Google Trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool as it may contain inaccuracies \autocite{carneiro2009google}.
+According to \textcite{nuti2014use}, the data retrieval process should be transparent.
 This will increase the trustworthiness of both the results and the generalizability of the findings.
-Furthermore, researchers should clearly document the rationale and data retrieval process to ensure the reproducibility of results \autocite{nuti2014use}.
-.
+Furthermore, \textcite{nuti2014use} believe that researchers should clearly document the rationale and data retrieval process to ensure the reproducibility of results.
 
@@
 
-In this work we also used Google Analytics search index series as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
-We used weekly Google Analytics search index series of various distance learning solutions during COVID-19 pandemic.
+In this work we also used Google Trend search quiries as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
+We used weekly Google Trend search quiries of various distance learning solutions during COVID-19 pandemic.
 We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of COVID-19 pandemic.
 Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also studied the cross correlation between weekly web searches and the weekly cases of COVID-19.
 
@@
 \autocite{alicino2015assessing,vaughan2014web}.
-A value of 100 represents the peak popularity of the term.
+As such, a value of 100 represents the peak popularity of the term.
 
@@
 Eventhough, the solutions they had recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base and evidence of impact \autocite{unesco2020DLsolutions}.
-Different solutions are categorized based on distance learning needs.
 For each keyword, the worldwide search queries were performed with a keyword being used as the ``search term''.
@@
 However, according to \textcite{vaughan2014web}, short forms or acronyms had higher search volumes than the corresponding complete names, in general.
-However we did not use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
+However, we did not use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
 
-Google Trends allows to search for up to five queries each time \autocite{vaughan2014web}.
 We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
@@
 
-In google trends, it is possible to search for only up to five queries at a time.
+In google trends, it is possible to search for only up to five queries at a time \autocite{vaughan2014web}.
 Therefore, under each segmentation we entered up to five distance learning tools at a time and recorded their relative ranking scores.
@@
 
-Both visual and analytical approaches were used to analyse the changes in web search queries at the global level over the entire period of the COVID-19 pandemic to identify the emerging evidence about the impact of COVID-19 pandemic on distance education.
+Both visual and analytical approaches were used to analyse the changes in web search queries at the global level to identify the emerging evidence about the impact of COVID-19 pandemic on distance education.
 We also evaluated the relationship between weekly web searches and the number of weekly global cases of COVID-19.

### 2021-12-26 main.Rmd
updated methodology

@@
 
-Google Trends, is a  freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages [@carneiro2009google].
+Google Trends is a  freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages [@carneiro2009google].
 Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution  [@alicino2015assessing; @cook2011assessing].
 According to @jarynowski2020perception the public concerns on different matters can also take an epidemic nature  starting with the phase of growing interest which is known as "early adoption", to the phase of general interest which is known as "majority" to eventually lose popularity which is known as "lagers stage".
-Figure \ref{fig:distanceLearningWorldAnalysis} also confirms this life cycle explanation.
+Figure \ref{fig:distanceLearningWorldAnalysis} (b) and (c) also confirm this life cycle explanation of Google Trend search patterns.
 
-The use of google trend search queries has become a popular data source in monitoring and modeling the dynamics of different domains such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding] and health care [@nuti2014use; @alicino2015assessing; @arora2019google; @carneiro2009google; @cook2011assessing] as it provides data about social phenomena in a more timely manner than traditional data collection processes [@vaughan2014web].
+The use of Google Trend search queries has become a popular data source in monitoring and modeling the dynamics of different application domains such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding] and health care [@nuti2014use; @alicino2015assessing; @arora2019google; @carneiro2009google; @cook2011assessing] as it has the ability to provide data about social phenomena in a more timely manner than traditional data collection processes [@vaughan2014web].
 According to the study conducted by @willis2020using on taxi preferences found that online queries can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
@@
 The study conducted by @husnayain2020applications focuses on search terms related to the coronavirus, handwashing, and face masks  to monitor public restlessness toward COVID-19.
-However, there are only a limited number of research attempts related to education section  using Google trend search query data [@vaughan2014web].
-The study conducted by @vaughan2014web have used  google trend search queries to predict academic fame by testing the correlation between search volume data and university ranking data.
+However, there are only a limited number of research attempts related to education sector  using Google Trend search query data [@vaughan2014web].
+The study conducted by @vaughan2014web have used  Google Trend search queries to predict academic fame by testing the correlation between search volume data and university ranking data.
 @kansal2021google have used a text mining approach to identify emerging patterns of words and phrases during COVID-19 pandemic, keeping education prospects as the focus.
 However, their study was limited to a small number of  learning platforms.
-However, according to @moore2011learning the design of different types of learning environments can depend on the learning objective, target audience, access (physical, virtual and/or both), and type of content.
+According to @moore2011learning, the design of different types of learning environments can depend on the learning outcomes, target audience, access (physical, virtual and/or both), and/or type of content.
 Therefore, this study    provides a detailed investigation on distance  education solutions focusing on different distance learning needs.
-In contrast to the work in @kansal2021google, the work in our paper has a very specific focus: the distance learning solutions for various education purposes.
+In contrast to the work in @kansal2021google, the work in our paper has a very specific focus and provides a thorough investigation on different distance learning solutions available for various education purposes.
 
-Eventhough google trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool.
-It may contain inaccuracies [@carneiro2009google].
-According to @nuti2014use the data retrieval process should be transparent.
+Eventhough Google Trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool as it may contain inaccuracies [@carneiro2009google].
+According to @nuti2014use, the data retrieval process should be transparent.
 This will increase the trustworthiness of both the results and the generalizability of the findings.
-Furthermore, researchers should  clearly document the rationale and data retrieval process to ensure the reproducibility of results [@nuti2014use].
-.
+Furthermore, @nuti2014use believe that researchers should  clearly document the rationale and data retrieval process to ensure the reproducibility of results.
 
@@
 
-In this work we also used  Google Analytics search index series  as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
-We used weekly  Google Analytics search index series of various distance learning solutions during COVID-19 pandemic.
+In this work we also used  Google Trend search quiries   as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
+We used weekly  Google Trend search quiries of various distance learning solutions during COVID-19 pandemic.
 We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the  period of COVID-19 pandemic.
 Through this we wanted to capture the volume of interest in the public realm about distance education solutions and how this changes over time.
-We also studied the cross correlation between weekly web searches and the  weekly  cases of COVID-19.
 
@@
 [@alicino2015assessing; @vaughan2014web].
-A value of 100 represents the peak popularity of the term.
+As such, a value of 100 represents the peak popularity of the term.
 
@@
 Eventhough, the solutions they had recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base and evidence of impact [@unesco2020DLsolutions].
-Different solutions are categorized based on distance learning needs.
 For each keyword, the worldwide search queries were performed with a keyword being used  as the "search term".
@@
 However, according to @vaughan2014web,  short forms or acronyms had higher search volumes than the corresponding complete names, in general.
-However we did not  use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
+However, we did not  use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
 
-Google Trends allows to search for up to five queries each time [@vaughan2014web].
 We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
@@
 
-In google trends, it is possible to search for only up to five queries at a  time.
+In google trends, it is possible to search for only up to five queries at a  time [@vaughan2014web].
 Therefore, under each segmentation we entered up to five distance learning tools at a time and recorded their relative ranking scores.
@@
 
-Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level over the entire period of the COVID-19 pandemic to identify the emerging evidence about the impact of  COVID-19 pandemic on distance education.
+Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level to identify the emerging evidence about the impact of  COVID-19 pandemic on distance education.
 We also evaluated the  relationship between weekly web searches and the number of  weekly global cases of COVID-19.

### 2021-12-26 Paper_online_Learning.tex
update paper

@@
 
-Search volume index graphs for COVID-19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic.
-Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic.
+The visual representation of Google Trend search quiries of COVID-19 related terms worldwide shows that the highest peak coincided with the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (b)).
+Surprisingly, the search volume index graph for distanced education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
+This highlight the impact of COVID-19 pandemic on ditance education.
+
 According to \textcite{moore2011learning}, ``distance education'' is the most renowned descriptor used when referencing distance learning.
-However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``never version'' and ``improved version'' of distance learning.
-The google search trend pattern corresponds to the term ``online learning'' (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout COVID-19 pandemic.
+However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``newer version'' and ``improved version'' of distance learning.
+The Google Trend search patterns corresponds to the term ``online learning'' confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
 Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``Distance learning'' and ``online teaching'' demonstrate a similar trend during the study period.
 COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases.
-A similar decline could be detected in distance education related terms.
-further, a second was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
+A similar decline can be observed in distance education related terms (Figures \ref{fig:distanceLearningWorldAnalysis}).
+This confirms the life cycle of Google Trend search pattern introduced in \textcite{jarynowski2020perception}.
+Further, a second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
 This could be due to the near total closures of schools happened with the declaration of COVID-19 as a pandemic.
@@
 
-The pairwise comparison performed using Dynamic Time Warping algorithm.
-Under this approach the two time series are stretched or compressed locally in order to make one resemble the other as much as possible.
+A pairwise comparison was performed using Dynamic Time Warping algorithm.
+Under this approach, two time series are stretched or compressed locally in order to make one resemble the other as much as possible.
 The distance between the two series is then computed by summing the distances of individual aligned elements \autocite{giorgino2009computing}.
-Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact from COVID-19 on distance education.
-Both pairwise comparison show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)).
+Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact of COVID-19 on distance education.
+Both pairwise comparisons show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)).
 The normalized distance between the two series in the two comparisons are 0.1326197 and 0.1828299, respectively.
-This also provides further confirmation for the similarity of the google trend search profiles.
+This also provides further confirmation for the similarity of the Google Trend search profiles.
+this highlight the sudden public attention towards both aspects, COVID-19 and distance education, eventhough distance education has a long history in contrast to COVID-19.
 
@@
 series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-In 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide.
-By the time it was know as ``corona'', it was not a global concern and was mostly limited to China.
-During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic all around the world.
-This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis.
+In later 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was know as ``corona'' it was not a global concern and was mostly limited to China.
+During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic.
+This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis of the Figure \ref{fig:dtw1} (b).
 
-Later with the spread of the virus worldwide, WHO named it as COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
+Later, with the spread of the virus worldwide, WHO named it as COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
 With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
-With this sudden unexpected movement, online learning also became an equally important global concern, similar to COVID-19 pandemic.
-This similar behaviour in public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line from approximately 120 to 150 on the Query Index axis.
+With this sudden unexpected massive shutdown, online learning also became an equally important global concern, similar to COVID-19 pandemic.
+This similar behaviour in public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line from approximately 120 to 150 on the Query Index axis of the Figure \ref{fig:dtw2} (b).
 This also confirms the impact of COVID-19 on distance education and online learning.
 
-Then, a cross correlation analysis was performed to estimate the lag between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
+Next, a cross correlation analysis was performed to estimate the lag time between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
 A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''.
@@
 This provides a fast first step to identify the most popular tools available for different educational purposes.
-Under each and every category Figure \ref{fig:plot1} (a) - (e) and Figure \ref{fig:plot2} (f) - (j), one or two tools have got a significantly higher level of attention in comparison to the remaining options available undue the same category.
-The time series across multiple panels are not comparable as they represent ``relative'' search volume of given terms and they are comparable only with a given panel.
-The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each category.
+Under each and every panel, one or two tools have got a significantly higher level of attention in comparison to the remaining options available under the same category (Figure \ref{fig:plot1} (a) - (e) and Figure \ref{fig:plot2} (f) - (j)).
+However, the Google Trend series across multiple panels are not comparable as they represent ``relative'' search volume of given terms.
+They are comparable only within a given panel.
+The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each panel.
 
@@
 Under systems built for use on basic mobile phones, both `kaiOS' and `Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
-`KaiOs' has the ability provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
+`KaiOs' has the ability to provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
 `Ubongo' has also received a considerable attention, eventhough its target groups are school-age children and their parents in Africa.
@@
 Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
-The can be a huge relief for financially distressed families during COVID-19 pandemic.
+These distance learning solutions can provide a huge relief for financially distressed families during COVID-19 pandemic.
 Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
@@
 The second highest option only provides support for language learning.
-Quizlet's ability provide support for multiple subject might be the reason for it to outperform it competitor in the market.
+Quizlet's ability to provide support for multiple subject might be the reason for it to outperform its competitor in the market.
 Under mobile reading applications, `Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
 Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
-However, `WhatsApp' has outperformed `Zoom' overtime through with increasing public attention during the pandemic.
+However, `WhatsApp' has outperformed `Zoom' with increasing public attention during the pandemic.
 In contrary, the global attention towards `Zoom' has decreased over time.
-`Teams' has also got considerable attention in comparison to the other remaining options available int he market.
+`Teams' has also got considerable attention in comparison to the other remaining options available in the market.
 Under tools for teachers to create of digital learning content, `Trello' which provides a visual collaboration platform for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
@@
 
-At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
-These peaks aligns with the near-total closures of schools due to COVID-19 in Figure \ref{fig:covidImpactWorld}.
+At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
+These peaks aligns with the near-total closures of schools due to COVID-19 pandemic (Figure \ref{fig:covidImpactWorld}).
 The major concerns that arise with the unexpected sudden movement from physical classroom to virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
-Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
+Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
 The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.
@@
 
-\caption{Google search trends footprint analysis}\label{fig:plot1}
+\caption{Google Trend footprint analysis of distance learning solutions during COVID-19 pandemic}\label{fig:plot1}
 \end{figure}
@@
 
-\caption{Google search trends footprint analysis}\label{fig:plot2}
+\caption{Google Trend footprint analysis of distance learning solutions during COVID-19 pandemic}\label{fig:plot2}
 \end{figure}
-
-The attention in the education sector is low compared to all category.
-However a significant peak can be observed during the first academic break after declaring COVID-19 as a pandemic.
-Similar peak can be observed in 2021 .
-But that significant compared to 2020 peak.
 
@@
 
-{\centering \includegraphics[width=1\textwidth]{figure/psychosocialSupportAnalysis-1}
+{\centering \includegraphics[width=1\textwidth]{figure/psychosocialSupport-1}
 
@@
 
-\caption{Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' search category}\label{fig:psychosocialSupportAnalysis}
+\caption{Goolge search trend of the term psychosocial support under All categories and Education category during COVID-19 pandemic}\label{fig:psychosocialSupport}
 \end{figure}
@@
 Good psychosocial support is also vital to improve both students' and teachers' mental health and well-being during a pandemic.
-This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two sudden spikes in the search volume index graphs of the term `Psychosocial support' which correspond well with the academic breaks in August in both years 2020 and 2021, under `Education' category.
+This is evident from the two sudden spikes in the search volume index graphs of the term `Psychosocial support' ( Figure \ref{fig:psychosocialSupport} ) which correspond well with the academic breaks in August in both years, 2020 and 2021, under `Education' category.
 This highlight the urgent need of attention and solutions towards these aspects.
-This type of an analysis based on secondary data is very vital as these types of aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
+This type of an analysis based on secondary data is vital as these aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
 
@@
 
-In this study we investigated whether the Google trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
+In this study we investigated whether the Google Trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
 This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic.
@@
 
-Our findings suggest that web search interest in distance learning solutions at the global was strongly influenced by the COVID-19 pandemic.
-Google trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
-The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed google trend footprint analysis.
+Our findings suggest that web search interest in distance education at the global level was strongly influenced by the COVID-19 pandemic.
+Google Trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
+The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed Google Trend footprint analysis.
 The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
@@
 High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the COVID-19 pandemic.
-These findings offer important insights to various stakeholders of education to make better decision with less effort and time.
+Therefore, the findings of this study offer important insights to various stakeholders of education to make better decision with less time and effort.
 

### 2021-12-26 main.Rmd
update paper

@@
 
-Search volume index graphs for COVID-19 related terms worldwide (Figures \ref{fig:distanceLearningWorldAnalysis} (b)) shows that the highest peak coincided with the start of the pandemic.
-Surprisingly, the search volume index graph for distanced education related terms (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) also shows a similar spike at the start of the pandemic.
+The visual representation of Google Trend search quiries of COVID-19 related terms worldwide shows that the highest peak coincided with the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (b)).
+Surprisingly, the search volume index graph for distanced education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
+This highlight the impact of COVID-19 pandemic on ditance education.
+
 According to @moore2011learning, "distance education" is the most renowned descriptor used when referencing distance learning.
-However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version", "never version" and "improved version" of distance learning.
-The google search trend pattern corresponds to the term "online learning"  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)) confirms the above claim as it is the most searched distance education related term throughout  COVID-19 pandemic.
+However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version", "newer version" and "improved version" of distance learning.
+The Google Trend search patterns corresponds to the term "online learning" confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
 Despite quantitative differences in terms of relative search volume, the series generated by the two terms "Distance learning" and "online teaching" demonstrate a similar trend during the study period.
 COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases.
-A similar decline could be detected in distance education related terms.
-further,  a second was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
+A similar decline can be observed in distance education related terms  (Figures \ref{fig:distanceLearningWorldAnalysis}).
+This confirms the life cycle of Google Trend search pattern introduced in @jarynowski2020perception.
+Further,  a second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
 This could be due to the near total closures of schools happened  with the declaration of  COVID-19 as a pandemic.
@@
 
-The pairwise comparison performed  using  Dynamic Time Warping algorithm.
-Under this approach the two time series are stretched or compressed  locally in order to make one resemble the other as much as possible.
+A pairwise comparison was performed  using  Dynamic Time Warping algorithm.
+Under this approach, two time series are stretched or compressed  locally in order to make one resemble the other as much as possible.
 The distance between the two series is then computed by summing the distances of individual aligned elements [@giorgino2009computing].
-Two pairwise comparisons, "corona"  with "online learning" and   "covid" with "online learning" were conducted to see the impact  from COVID-19 on distance education.
-Both pairwise comparison show a significant  areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and  \ref{fig:dtw2} (a)).
+Two pairwise comparisons, "corona"  with "online learning" and   "covid" with "online learning" were conducted to see the impact  of COVID-19 on distance education.
+Both pairwise comparisons show a significant  areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and  \ref{fig:dtw2} (a)).
 The normalized distance between the two series in the two comparisons are  0.1326197 and 0.1828299, respectively.
-This also provides further confirmation for  the similarity of the google trend  search profiles.
+This also provides further confirmation for  the similarity of the Google Trend  search profiles.
+this highlight the sudden public attention towards both aspects, COVID-19 and distance education, eventhough distance education has a long history in contrast to COVID-19.
 
@@
 series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-In 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide.
-By the time it was know as "corona", it was not a global concern and was mostly limited to China.
-During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic all around the world.
-This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis.
+In later 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide [@ciotti2019covid].When the virus was know as "corona" it was not a global concern and was mostly limited to China.
+During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic.
+This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis of the Figure \ref{fig:dtw1} (b).
 
-Later with the spread of the virus worldwide, WHO  named it as  COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
+Later, with the spread of the virus worldwide, WHO  named it as  COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
 With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
-With this sudden unexpected movement, online learning  also became an equally important global concern, similar to COVID-19 pandemic.
-This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that fall on near 45-degree line  from approximately 120 to 150 on the Query Index axis.
+With this sudden unexpected massive shutdown, online learning  also became an equally important global concern, similar to COVID-19 pandemic.
+This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line  from approximately 120 to 150 on the Query Index axis of the Figure \ref{fig:dtw2} (b).
 This also confirms the impact of COVID-19 on distance education and online learning.
 
-Then, a cross correlation analysis was performed to estimate the lag between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
+Next, a cross correlation analysis was performed to estimate the lag time between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
 A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms "Online learning", "Distance learning" and "Online teaching".
@@
 This provides a fast first step to identify the most popular tools available for different educational purposes.
-Under each and every category Figure \ref{fig:plot1} (a) - (e) and  Figure \ref{fig:plot2} (f) - (j), one or two  tools have got a significantly higher level of attention in comparison to the remaining options available undue the same category.
-The time series across multiple panels are not comparable as they represent "relative" search volume of given terms and they are comparable only with a given panel.
-The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each category.
+Under each and every panel, one or two  tools have got a significantly higher level of attention in comparison to the remaining options available under the same category (Figure \ref{fig:plot1} (a) - (e) and  Figure \ref{fig:plot2} (f) - (j)).
+However, the Google Trend series across multiple panels are not comparable as they represent "relative" search volume of given terms.
+They are comparable only within a given panel.
+The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each panel.
 
@@
 Under systems built for use on basic mobile phones, both 'kaiOS' and  'Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
-'KaiOs'  has the ability provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
+'KaiOs'  has the ability to provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
 'Ubongo' has also received a considerable attention, eventhough its target groups are school-age children and their parents in Africa.
@@
 Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
-The can be a huge relief for financially distressed families during COVID-19 pandemic.
+These distance learning solutions can provide a huge relief for financially distressed families during COVID-19 pandemic.
 Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
@@
 The second highest option only provides  support for language learning.
-Quizlet's ability provide support for multiple subject might be the reason for it to outperform it  competitor in the market.
+Quizlet's ability to provide support for multiple subject might be the reason for it to outperform its  competitor in the market.
 Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
 Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
-However, 'WhatsApp' has outperformed 'Zoom' overtime through with increasing public attention during the pandemic.
+However, 'WhatsApp' has outperformed 'Zoom'  with increasing public attention during the pandemic.
 In contrary, the global attention towards 'Zoom'  has decreased over time.
-'Teams' has also got considerable attention in comparison to the other remaining options available int he market.
+'Teams' has also got considerable attention in comparison to the other remaining options available in the market.
 Under tools for teachers to create of digital learning content, 'Trello'  which provides a visual collaboration platform  for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
@@
 
-At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both 'digital learning management systems (Figure \ref{fig:plot1} (a))  and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
-These peaks aligns with the near-total closures of schools due to COVID-19 in Figure \ref{fig:covidImpactWorld}.
+At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both digital learning management systems (Figure \ref{fig:plot1} (a))  and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
+These peaks aligns with the near-total closures of schools due to COVID-19 pandemic (Figure \ref{fig:covidImpactWorld}).
 The major concerns that arise with the unexpected sudden movement from physical classroom to  virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
-Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create of digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
+Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
 The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reason for this considerable public attention around August, 2020.
@@
 
-```{r plot1, fig.height=12, fig.cap="Google search trends footprint analysis"}
+```{r plot1, fig.height=12, fig.cap="Google Trend footprint analysis of distance learning solutions during COVID-19 pandemic"}
 ```
 
-```{r plot2, fig.height=12, fig.cap="Google search trends footprint analysis"}
+```{r plot2, fig.height=12, fig.cap="Google Trend footprint analysis of distance learning solutions during COVID-19 pandemic"}
 ```
-
-The attention in the education sector is low compared to all category.
-However a significant peak can be observed during the first academic break  after declaring  COVID-19 as a pandemic.
-Similar peak can be observed in 2021 .
-But that significant compared to 2020 peak.
 
@@
 
-```{r psychosocialSupportAnalysis, fig.height= 3, fig.cap="Goolge search trend of the term 'Psychosocial support' under 'All categories' and 'Education' search category"}
+```{r  psychosocialSupport, fig.cap = "Goolge search trend of the term psychosocial support under All categories and Education category during COVID-19 pandemic",  fig.height= 4}
+
 ```
@@
 Good psychosocial support is also vital to improve both students' and  teachers' mental health and well-being during a pandemic.
-This is evident in Figure \ref{fig: psychosocialSupportAnalysis} from the two sudden spikes in the search volume index graphs of the term 'Psychosocial support'  which correspond well with  the academic breaks in August in both years 2020 and 2021, under 'Education' category.
+This is evident from the two sudden spikes in the search volume index graphs of the term 'Psychosocial support' ( Figure \ref{fig:psychosocialSupport} ) which correspond well with  the academic breaks in August in both years, 2020 and 2021, under 'Education' category.
 This highlight the urgent need of attention and solutions towards these aspects.
-This type of an analysis based on secondary data is very vital as these types of aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
+This type of an analysis based on secondary data is vital as these aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
 
@@
 
-In this study  we investigated whether the Google  trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
+In this study  we investigated whether the Google  Trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
 This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic.
@@
 
-Our findings suggest that web search interest in distance learning solutions at the global  was strongly influenced by the COVID-19 pandemic.
-Google trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
-The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed google trend footprint analysis.
+Our findings suggest that web search interest in distance education at the global level was strongly influenced by the COVID-19 pandemic.
+Google Trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
+The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed Google Trend footprint analysis.
 The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
@@
 High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the COVID-19 pandemic.
-These findings offer important insights to  various stakeholders of education  to make better decision with less effort and time.
+Therefore, the findings of this study offer important insights to  various stakeholders of education  to make better decision with less  time and effort.
 

### 2021-12-26 src/main.R

@@ -1113,7 +1113,7 @@ data_world <- bind_rows(data_world_all, data_world_edu )
 write.csv(data_world, here::here( "data", "psychosocial_support_world.csv"))
 
 
-## ---- psychosocialSupportAnalysis
+## ---- psychosocialSupport
 keywords <- c("psychosocial support")
 data_world <- read.csv( here::here( "data", "psychosocial_support_world.csv"))
 
@@ -1121,24 +1121,15 @@ dataW <- data_world %>%
   select(date, category, hits) %>%
   mutate(category = factor(category, levels = c(0, 74), labels = c("All", "Education")))
 dataW$date = as.Date(dataW$date, "%Y-%m-%d")
-p1 <- dataW %>% 
-  ggplot(aes(x=date,y=category,fill=hits))+
-  geom_tile(size=0.1, colour = "grey50")+
-  scale_y_discrete(expand=c(0,0))+
-  ggtitle("All")+
-  labs(x="",y="")+
-  scale_fill_viridis()+
-  theme(panel.grid = element_blank(), text = element_text(size = 16))
 
-
-p11 <- dataW %>%
+p1 <- dataW %>%
   as_tsibble(index= date, key = category) %>%
   autoplot(hits, size= 1) +
   theme(legend.position = "bottom",  
         text = element_text(size = 12), 
         legend.title = element_blank(), 
         legend.text = element_text(size = 12))+
-  scale_colour_viridis_d(guide = "colourbar", direction = -1, end = 0.6) 
-
-print(p11)
+  scale_colour_viridis_d(guide = "colourbar",
+                         direction = -1, end = 0.6) 
+print(p1)
 

### 2021-12-26 Paper_online_Learning.Rmd
edit abstract

@@
 However COVID-19 has made a new era of distance education.
-With the increasing demand, various different distance learning solutions have been introduced for different education purposes.
+With the increasing demand various distance learning solutions have been introduced for different education purposes.
 In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions.
 Both visual and analytical approaches were used to  analyse global level web search querie during COVID-19 pandemic.
-The findings of this study include a Google Trend footprint of various distance leraning solutions for diffrent education purposes that can be used as a proxy for measuring global attention towards different distance learning solutions.
+The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions.
 The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes.

### 2021-12-26 Paper_online_Learning.tex
edit abstract

@@
 However COVID-19 has made a new era of distance education.
-With the increasing demand, various different distance learning solutions have been introduced for different education purposes.
+With the increasing demand various distance learning solutions have been introduced for different education purposes.
 In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the popularity and the public interest towards different distance education solutions.
 Both visual and analytical approaches were used to analyse global level web search querie during COVID-19 pandemic.
-The findings of this study include a Google Trend footprint of various distance leraning solutions for diffrent education purposes that can be used as a proxy for measuring global attention towards different distance learning solutions.
+The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions.
 The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes.

### 2021-12-26 Paper_online_Learning.Rmd
update keywords

@@
 ---
-title: "Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends"
+title: "COVID-19 and Online Learning Tools"
 author:
@@
 R code and data to reproduce the results of this work are available in the online supplementary materials."
-keywords:  "Online-learning, Distance education solutions, COVID-19 Pandemic,  Google Trend search queries "
+keywords:  "Online-learning, Distance education solutions, COVID-19 Pandemic,  Google Trend search queries, Online-teaching "
 wpnumber: 1

### 2021-12-26 Paper_online_Learning.tex
update keywords

@@
 \hypersetup{
-            pdftitle={Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends},
-            pdfkeywords={Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries},
+            pdftitle={COVID-19 and Online Learning Tools},
+            pdfkeywords={Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries, Online-teaching},
             pdfborder={0 0 0},
@@
 
-\title{Attention Towards Distance Education Tools During COVID-19 Pandemic: Evidence from Google Trends}
+\title{COVID-19 and Online Learning Tools}
 
@@
 \begin{keywords}
-Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries
+Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries, Online-teaching
 \end{keywords}

### 2021-12-26 Paper_online_Learning.Rmd
edit typos

@@
 However COVID-19 has made a new era of distance education.
-With the increasing demand various distance learning solutions have been introduced for different education purposes.
-In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions.
-Both visual and analytical approaches were used to  analyse global level web search querie during COVID-19 pandemic.
+Due to the  increasing demand, various distance learning solutions have been introduced for different distance education purposes.
+In this study we investigated the impact of COVID-19 on distance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions.
+Both visual and analytical approaches were used to  analyse global level web search queries during COVID-19 pandemic.
 The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions.
-The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes.
+This can provide a fast first step guide to identify the most popular online learning tools available for different educational purposes.
 R code and data to reproduce the results of this work are available in the online supplementary materials."
-keywords:  "Online-learning, Distance education solutions, COVID-19 Pandemic,  Google Trend search queries, Online-teaching "
+keywords:  "Online learning, Online teaching Distance education solutions, COVID-19 Pandemic,  Google Trend search queries"
 wpnumber: 1

### 2021-12-26 Paper_online_Learning.tex
edit typos

@@
             pdftitle={COVID-19 and Online Learning Tools},
-            pdfkeywords={Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries, Online-teaching},
+            pdfkeywords={Online learning, Online teaching Distance education solutions, COVID-19 Pandemic, Google Trend search queries},
             pdfborder={0 0 0},
@@
 However COVID-19 has made a new era of distance education.
-With the increasing demand various distance learning solutions have been introduced for different education purposes.
-In this study we invetigated the impact of COVID-19 on ditance education and the possibility of using Google Trend search queries as a proxy to quantify the popularity and the public interest towards different distance education solutions.
-Both visual and analytical approaches were used to analyse global level web search querie during COVID-19 pandemic.
+Due to the increasing demand, various distance learning solutions have been introduced for different distance education purposes.
+In this study we investigated the impact of COVID-19 on distance education and the possibility of using Google Trend search queries as a proxy to quantify the popularity and the public interest towards different distance education solutions.
+Both visual and analytical approaches were used to analyse global level web search queries during COVID-19 pandemic.
 The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions.
-The can provide a fast first step guide to identify the most popular distance learning tools available for different educational purposes.
+This can provide a fast first step guide to identify the most popular online learning tools available for different educational purposes.
 R code and data to reproduce the results of this work are available in the online supplementary materials.
@@
 \begin{keywords}
-Online-learning, Distance education solutions, COVID-19 Pandemic, Google Trend search queries, Online-teaching
+Online learning, Online teaching Distance education solutions, COVID-19 Pandemic, Google Trend search queries
 \end{keywords}
@@
 According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.
-In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure.
+In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
 There are many different distance learning solutions.
@@
 Few common features found in all these terms are, they all involve some form knowledge sharing between two parties; learners and educators, different times, different geographical locations and/or varying forms of instructional materials and tools \autocite{moore2011learning}.
-Further, there are different types of interactions such are learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction that can facilitate learners to construct knowledge in distance education environments.
-These \autocite{wallace2003online}.
+Further, distance education has different types of interactions such as learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction that can facilitate learners to construct knowledge \autocite{wallace2003online}.
+Different tools are required to fulfill all these requirements in a distance learning environment.
 
@@
 
-We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
+We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
 The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.
@@
 Surprisingly, the search volume index graph for distanced education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
-This highlight the impact of COVID-19 pandemic on ditance education.
+This highlights the impact of COVID-19 pandemic on distance education.
 
@@
 A similar decline can be observed in distance education related terms (Figures \ref{fig:distanceLearningWorldAnalysis}).
-This confirms the life cycle of Google Trend search pattern introduced in \textcite{jarynowski2020perception}.
+This confirms the life cycle of Google Trend search patterns introduced in \textcite{jarynowski2020perception}.
 Further, a second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
@@
 
-The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The entirety of the blue warping path flows through green pastures further confirms the similarity between the search patterns related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after the COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
 
@@
 series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-In later 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was know as ``corona'' it was not a global concern and was mostly limited to China.
+In late 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was know as ``corona'' it was not a global concern and was mostly limited to China.
 During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic.
-This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis of the Figure \ref{fig:dtw1} (b).
+This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).
 
@@
 With this sudden unexpected massive shutdown, online learning also became an equally important global concern, similar to COVID-19 pandemic.
-This similar behaviour in public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line from approximately 120 to 150 on the Query Index axis of the Figure \ref{fig:dtw2} (b).
+This similar behaviour in public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line from approximately 120 to 150 on the Query Index axis in Figure \ref{fig:dtw2} (b).
 This also confirms the impact of COVID-19 on distance education and online learning.
@@
 Figure \ref{fig:plot1} and Figure \ref{fig:plot2} provide a Google trend footprint that can be used as a proxy for measuring global attention towards different distance learning solutions.
-This provides a fast first step to identify the most popular tools available for different educational purposes.
+This provides a fast first step guide to identify the most popular tools available for different educational purposes.
 Under each and every panel, one or two tools have got a significantly higher level of attention in comparison to the remaining options available under the same category (Figure \ref{fig:plot1} (a) - (e) and Figure \ref{fig:plot2} (f) - (j)).
@@
 Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
-The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reason for this considerable public attention around August, 2020.
+The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reasons for this considerable public attention around August, 2020.
 
@@
 Google Trends only captures the search behaviour of a sub population with Internet access, eventhough it is the most common worldwide search engine.
-Our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
+Further, our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
 Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.

### 2021-12-26 main.Rmd
edit typos

@@
 According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught.
-In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student distance learning during the period of school closure.
+In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
 There are many different distance learning solutions.
@@
 Few common features found in all these terms are, they all involve some form knowledge sharing between two parties; learners and educators, different times, different geographical locations  and/or  varying forms of instructional materials and tools [@moore2011learning].
-Further, there are different types  of interactions such are learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction that can facilitate learners to construct knowledge in distance education environments.
-These [@wallace2003online].
+Further, distance education has different types  of interactions such as learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction that can facilitate learners to construct knowledge  [@wallace2003online].
+Different tools are required to fulfill all these requirements in a distance learning environment.
 
@@
 
-We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, Collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
+We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
 The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.
@@
 A similar decline can be observed in distance education related terms  (Figures \ref{fig:distanceLearningWorldAnalysis}).
-This confirms the life cycle of Google Trend search pattern introduced in @jarynowski2020perception.
+This confirms the life cycle of Google Trend search patterns introduced in @jarynowski2020perception.
 Further,  a second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
@@
 
-The entirety of the blue warping path flows through green pastures further confirms the similarity between the search pattern related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The entirety of the blue warping path flows through green pastures further confirms the similarity between the search patterns related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after the COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
 
@@
 series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-In later 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide [@ciotti2019covid].When the virus was know as "corona" it was not a global concern and was mostly limited to China.
+In late 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide [@ciotti2019covid].When the virus was know as "corona" it was not a global concern and was mostly limited to China.
 During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic.
-This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis of the Figure \ref{fig:dtw1} (b).
+This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).
 
@@
 With this sudden unexpected massive shutdown, online learning  also became an equally important global concern, similar to COVID-19 pandemic.
-This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line  from approximately 120 to 150 on the Query Index axis of the Figure \ref{fig:dtw2} (b).
+This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line  from approximately 120 to 150 on the Query Index axis in Figure \ref{fig:dtw2} (b).
 This also confirms the impact of COVID-19 on distance education and online learning.
@@
 Figure \ref{fig:plot1} and  Figure \ref{fig:plot2} provide a Google trend footprint that can be used as a proxy for measuring global attention towards different distance learning solutions.
-This provides a fast first step to identify the most popular tools available for different educational purposes.
+This provides a fast first step guide to identify the most popular tools available for different educational purposes.
 Under each and every panel, one or two  tools have got a significantly higher level of attention in comparison to the remaining options available under the same category (Figure \ref{fig:plot1} (a) - (e) and  Figure \ref{fig:plot2} (f) - (j)).
@@
 Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
-The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reason for this considerable public attention around August, 2020.
+The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reasons for this considerable public attention around August, 2020.
 
@@
 Google Trends only captures the search behaviour of a sub population with Internet access, eventhough it  is the most common worldwide search engine.
-Our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
+Further, our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
 Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.

### 2021-12-27 Paper_online_Learning.Rmd
updated abstract

@@
   email: ttalagala@sjp.ac.lk
-abstract: "Distance Education has a long history.
-However COVID-19 has made a new era of distance education.
+abstract: "Distance education has a long history.
+However, COVID-19 has made a new era of distance education.
 Due to the increasing demand, various distance learning solutions have been introduced for different distance education purposes.
-In this study we investigated the impact of COVID-19 on distance education and the possibility of using Google Trend search queries as a proxy to quantify the  popularity and the public interest towards different distance education solutions.
-Both visual and analytical approaches were used to  analyse global level web search queries during COVID-19 pandemic.
-The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions.
-This can provide a fast first step guide to identify the most popular online learning tools available for different educational purposes.
-R code and data to reproduce the results of this work are available in the online supplementary materials."
+In this study, we investigated the impact of COVID-19 on global attention towards different distance learning-teaching tools.
+We used Google Trend search queries as a proxy to quantify the  popularity and public interest towards different distance education solutions.
+Both visual and analytical approaches were used to  analyse global level web search queries during the COVID-19 pandemic.
+This can provide a fast first step guide to identifying the most popular online learning tools available for different educational purposes.
+The R code and data to reproduce the results of this work are available in the online supplementary materials."
 keywords:  "Online learning, Online teaching Distance education solutions, COVID-19 Pandemic,  Google Trend search queries"

### 2021-12-27 Paper_online_Learning.tex
updated abstract

@@
 \begin{abstract}
-Distance Education has a long history.
-However COVID-19 has made a new era of distance education.
+Distance education has a long history.
+However, COVID-19 has made a new era of distance education.
 Due to the increasing demand, various distance learning solutions have been introduced for different distance education purposes.
-In this study we investigated the impact of COVID-19 on distance education and the possibility of using Google Trend search queries as a proxy to quantify the popularity and the public interest towards different distance education solutions.
-Both visual and analytical approaches were used to analyse global level web search queries during COVID-19 pandemic.
-The findings of this study include a Google Trend footprint of various distance leraning solutions that can be used as a proxy for measuring global attention towards different distance learning solutions.
-This can provide a fast first step guide to identify the most popular online learning tools available for different educational purposes.
-R code and data to reproduce the results of this work are available in the online supplementary materials.
+In this study, we investigated the impact of COVID-19 on global attention towards different distance learning-teaching tools.
+We used Google Trend search queries as a proxy to quantify the popularity and public interest towards different distance education solutions.
+Both visual and analytical approaches were used to analyse global level web search queries during the COVID-19 pandemic.
+This can provide a fast first step guide to identifying the most popular online learning tools available for different educational purposes.
+The R code and data to reproduce the results of this work are available in the online supplementary materials.
 \end{abstract}

### 2021-12-27 Paper_online_Learning.tex
working on intro section

@@
 
-COVID-19 is still one of the most important priorities of governments and media of many countries all around the world.
-Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 \autocite{world2020timeline}.
-As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}.
-In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment to minimize the spread of the virus.
-In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.
+COVID-19 is still one of the most important priorities of governments and media in many countries all around the world.
+Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on January 30, 2020, and then a pandemic on March 11, 2020 \autocite{world2020timeline}.
+As of September 29, 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths \autocite{dong2020interactive}.
+In response to the pandemic, health authorities worldwide have taken many steps, including vaccination development and deployment to minimize the spread of the virus.
+In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, and reduction of mass gatherings to reduce the spread of the virus \autocite{chang2020modelling}.
 
@@
 Many sectors have been affected ever since the outbreak of COVID-19, worldwide.
-Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world \autocite{daniel2020education}.
-Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 \autocite{unesco2020covid}.
-With the start of the pandemic a near total closure of schools were observed all around the world.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
+Among these many sectors, education is one of the most affected sectors with a near-total closure of schools, colleges, and universities all around the world \autocite{daniel2020education}.
+Figures \ref{fig:covidImpactWorld} show the evolution of global school closures and reopenings since mid February 2020 \autocite{unesco2020covid}.
+With the start of the pandemic, a near total closure of schools was observed all around the world.
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are held exclusively in person (Figures \ref{fig:covidImpactContinent}).
 
@@
 
-Eventhough distance education has a long history that goes back to almost two centuries with significant modifications,
-alterations, and addition in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously \autocite{richmond2020critical}.
+Even though distance education has a long history that goes back almost two centuries with significant modifications,
+alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}.
 Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.
-Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.
+Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.
 
-The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as teachers, students and parents.
-Teachers were not fully aware of their obligations and how to connect with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, lack of awareness of students' ongoing problems were some of the major issues they encountered during COVID-19 pandemic.
-Students on the other hand who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}.
-This also put an extra burden on parents, specially with limited education and resources, as they were expected to facilitate the required learning environment at home.
-Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures.
-Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown.
-This situation left internet as the only medium to support all these educational processes.
+The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders, such as teachers, students, and parents.
+Teachers were not fully aware of their obligations and how to connect with students to support their learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing tools and techniques and their effectiveness, lack of interaction and communication with the students, and lack of awareness of students' ongoing problems were some of the major issues they encountered during the COVID-19 pandemic.
+Students on the other hand who tend to have fewer educational opportunities beyond schools and universities are severely affected by this sudden and unexpected movement.
+Further, the impact was not limited to students' learning but also to other aspects of their lives, such as student debt, digital exclusion, lack of technology, lack of access to internet-enabled devices or a stable internet connection, long-term educational disengagement, poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts and lack of disability services \autocite{drane2020impact,daniel2020education,unescoadverse2020,richmond2020critical,carter2021teacher}.
+This also put an extra burden on parents, especially those with limited education and resources, as they were expected to facilitate the required learning environment at home.
+Working parents also find it difficult to work during a school closure due to childcare obligations that result from unexpected school closures.
+Furthermore, every single step, including planning, developing new tools and techniques, conducting awareness programmes about distance education, and shifting towards distance education, happened online, due to the unexpected massive global shutdown.
+This situation left the internet as the only medium to support all these educational processes.
 
 According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.
-In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
+In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
 There are many different distance learning solutions.
 Yet very little is known about the uptake of different solutions or about their effectiveness.
-Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and need of urgent responses.
+Further, getting familiar with all these available distance learning solutions is also not feasible or practical due to limited time availability and the need for urgent responses.
 

### 2021-12-27 main.Rmd
working on intro section

@@
 
-COVID-19 is still one of the most important priorities of governments and  media of many countries all around the world.
-Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on 30 January 2020, and then a pandemic on 11 March 2020 [@world2020timeline].
-As of 29 September 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].
-In response to the pandemic, health authorities worldwide have taken many steps including vaccination development and deployment  to minimize the spread of the virus.
-In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as  travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, reduction of mass gatherings to reduce the spread of the virus [@chang2020modelling].
+COVID-19 is still one of the most important priorities of governments and media in many countries all around the world.
+Due to the alarming levels of spread and severity of the virus, the World Health Organization (WHO) declared the outbreak a Public Health Emergency of International Concern (PHEIC) on January 30, 2020, and then a pandemic on March 11, 2020 [@world2020timeline].
+As of September 29, 2021, more than 232.7 million cases of COVID-19 have been reported in over 192 countries and territories, resulting in more than 4.7 million deaths [@dong2020interactive].
+In response to the pandemic, health authorities worldwide have taken many steps, including vaccination development and deployment  to minimize the spread of the virus.
+In addition, authorities worldwide have also taken many non-pharmaceutical interventions and preventive measures such as  travel restrictions, lock-downs, workplace hazard controls, school/university closures, facility closures, and reduction of mass gatherings to reduce the spread of the virus [@chang2020modelling].
 
@@
 Many sectors have been affected ever since the outbreak of COVID-19, worldwide.
-Among these many sectors education is one of the most affected sectors with a near-total closures of schools, colleges and universities, all around the world [@daniel2020education].
-Figures \ref{fig:covidImpactWorld} shows the evolution of global school closures and reopening since mid February 2020 [@unesco2020covid].
-With the start of the pandemic a near total closure of schools were observed all around the world.
-Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are being held exclusively in person (Figures \ref{fig:covidImpactContinent}).
+Among these many sectors, education is one of the most affected sectors with a near-total closure of schools, colleges, and universities all around the world [@daniel2020education].
+Figures \ref{fig:covidImpactWorld} show the evolution of global school closures and reopenings since mid February 2020 [@unesco2020covid].
+With the start of the pandemic, a near total closure of schools was observed all around the world.
+Over time, Africa and Oceania demonstrated better recovery in comparison to other regions with their increasing number of fully open schools in which the classes are held exclusively in person (Figures \ref{fig:covidImpactContinent}).
 
@@
 
-Eventhough distance education has a long history that goes back to almost two centuries with significant  modifications,
-alterations, and addition in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly made a new era of distance education while encouraging various stakeholders of education to take the concept of distance education seriously [@richmond2020critical].
+Even though distance education has a long history that goes back  almost two centuries with significant  modifications,
+alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].
 Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
-Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), eventhough distance education has a long history in contrast to COVID-19 pandemic.
+Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.
 
-The speed of the physical classroom closures and the rapid move to  online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders such as  teachers, students and parents.
-Teachers were not fully aware of their obligations and how to connect with students to support thier learning.
-Lack of preparation, lack of tools and techniques for distance education, lack of  awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, lack of awareness of students' ongoing problems  were some of the major issues they encountered during COVID-19 pandemic.
-Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities severely affected from this sudden unexpected movement.
-Further, the impact was not limited to students' learning but also to other aspects of their lives such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled device or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher].
-This also put an extra burden on parents, specially with limited education and resources,  as they were  expected to facilitate the required learning environment at home.
-Working parents also found it difficult to work during school closure due to childcare obligations that result from unexpected school closures.
-Further, every single step including planning, developing new tools and techniques, conducting awareness programmes about distance education, shifting towards distance education, happened through online, due to the unexpected massive global shutdown.
-This situation left internet as the only medium to support all these educational processes.
+The speed of the physical classroom closures and the rapid move to online delivery of education allowed little time for planning or reflection on potential risks that could happen to its various stakeholders, such as  teachers, students, and parents.
+Teachers were not fully aware of their obligations and how to connect with students to support their learning.
+Lack of preparation, lack of tools and techniques for distance education, lack of awareness of the availability of the existing  tools and techniques and their effectiveness, lack of interaction and  communication with the students, and lack of awareness of students' ongoing problems  were some of the major issues they encountered during the COVID-19 pandemic.
+Students  on the other hand  who tend to have fewer educational opportunities beyond schools and universities are severely affected by this sudden and unexpected movement.
+Further, the impact was not limited to students' learning but also to other aspects of their lives, such as  student debt, digital exclusion,  lack of technology, lack of access to internet-enabled devices or a stable internet connection, long-term educational disengagement,  poor nutrition and food insecurity, increased psychological challenges, childcare problems, exploitation, school dropouts  and lack of disability services [@drane2020impact; @daniel2020education; @unescoadverse2020; @richmond2020critical; @carter2021teacher].
+This also put an extra burden on parents, especially those with limited education and resources,  as they were  expected to facilitate the required learning environment at home.
+Working parents also find it difficult to work during a school closure due to childcare obligations that result from unexpected school closures.
+Furthermore, every single step, including planning, developing new tools and techniques, conducting awareness programmes about distance education, and shifting towards distance education, happened online, due to the unexpected massive global shutdown.
+This situation left the internet as the only medium to support all these educational processes.
 
 According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught.
-In response to this gray situation and urgent requirement of massive transformation from physical classroom to virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
+In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
 There are many different distance learning solutions.
 Yet very little is known about the uptake of different solutions or about their effectiveness.
-Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and need of urgent responses.
+Further, getting familiar with all these available distance learning solutions  is also not feasible or practical due to limited time availability and the need for urgent responses.
 

### 2021-12-27 Paper_online_Learning.tex
completed intro

@@
 This can provide a fast first step guide to identifying the most popular online learning tools available for different educational purposes.
+The results allow the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching.
 The R code and data to reproduce the results of this work are available in the online supplementary materials.
@@
 
-Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting suitable distance education solutions.
-According to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as it can indicate the prominence of the product in the market, its usability and its impact.
+Connectivity, flexibility, and the ability to promote varied interactions are some of the critical aspects educators consider when selecting suitable distance education solutions.
+According to \textcite{ahn2006utilizing}, the popularity of a product can also greatly influence consumer purchasing decisions as it can indicate the prominence of the product in the market, its usability, and its impact.
 Further, according to \textcite{willis2020using}, developers also tend to improve their products and services in response to increasing online search.
-Therefore, the internet represents a great opportunity to learn about the popularity and public attention towards a product or service.
-It is also a way to narrow down the search space and thereby identify most suitable product and services that meets customer needs.
+Therefore, the internet represents a great opportunity to learn about the popularity and public attention of a product or service.
+It is also a way to narrow down the search space and thereby identify the most suitable products and services that meet customer needs.
 
 Google Trends is an open source web analytics tool that allows users to get access to Internet search data that has the ability to provide deep insights into population behavior \autocite{nuti2014use}.
-As there is still a struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Trends search quiries can be used as a proxy of the popularity and the public interest in different distance education solutions.
-In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of COVID-19 pandemic on education?
-(2) What solutions are in place to meet the need of distance education in terms during COVID-19 pandemic?
-(3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?
+As there is still a struggle as to what technologies should be used to facilitate student learning, in this study, we test whether the Google Trends search quiries can be used as a proxy of the popularity and public interest in different distance education solutions.
+In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of the COVID-19 pandemic on education?
+(2) What solutions are in place to meet the needs of distance education during COVID-19 pandemic?
+(3) Which distance learning solutions have received wide attention and public interest during the COVID-19 pandemic?
 
 We primarily analyzed quantitative digital footprint data on the internet from December 2019 to August 2021.
-The resulted Google Trend footprint provides a fast first step guide to identify the most popular distance education tools available for different education purposes.
+The resulted Google Trend footprint provides a fast first step guide to identifying the most popular distance education tools available for different educational purposes.
 It also allows the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching.
-Further, identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
+Furthermore, identifying the most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
 
 Quality education is essential to sustainable development and it is a key United Nations (UN) sustainable development goal.
-Having the most appropriate tool is imperative to successful completion of a task at hand.
-The findings of this study provides an essential initial guidance to select the most appropriate tool for different distance education needs.
-Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.
+Having the most appropriate tool is imperative to the successful completion of a task at hand.
+The findings of this study provide essential initial guidance to selecting the most appropriate tool for different distance education needs.
+Therefore, the findings of this study directly contribute to the UN's sustainable development goals of quality education.
 
-The remainder of this paper is organized as follows.
-Section 2 presents the related work to lay the foundation for the Google trend footprint analysis of different distance learning solutions.
+The remainder of this paper is organized as follows: Section 2 presents the related work to lay the foundation for the Google trend footprint analysis of different distance learning solutions.
 Section 3 presents the methodology followed in the study.

### 2021-12-27 main.Rmd
completed intro

@@
 
-Connectivity, flexibility and ability to promote varied interactions are some critical aspects, educators consider when selecting suitable distance education solutions.
-According to @ahn2006utilizing, the popularity of a product can also greatly influence consumer purchasing decisions as it can  indicate the prominence of the product in the market, its usability and its impact.
+Connectivity, flexibility, and the ability to promote varied interactions are some of the critical aspects educators consider when selecting suitable distance education solutions.
+According to @ahn2006utilizing, the popularity of a product can also greatly influence consumer purchasing decisions as it can indicate the prominence of the product in the market, its usability, and its impact.
 Further, according to @willis2020using, developers also tend to improve their products and services in response to increasing online search.
-Therefore, the internet represents a great opportunity to learn about the popularity and public attention towards a product or service.
-It is also a way to narrow down the search space and  thereby identify most suitable product and services that meets customer needs.
+Therefore, the internet represents a great opportunity to learn about the popularity and public attention of a product or service.
+It is also a way to narrow down the search space and thereby identify the most suitable products and services that meet customer needs.
 
 Google Trends is an open source web analytics tool that allows users to get access to Internet search data that has the ability to provide deep insights into population behavior [@nuti2014use].
-As there is still a struggle as to what technologies should be used to facilitate student learning, in this study we test whether the Google Trends search quiries  can be used as a proxy of the popularity and the public interest in different distance education solutions.
-In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of  COVID-19 pandemic on education?
-(2) What solutions are in place to meet the need of distance education in terms during COVID-19 pandemic?
-(3) Which distance learning solutions have a wide attention and public interest during the COVID-19 pandemic?
+As there is still a struggle as to what technologies should be used to facilitate student learning, in this study, we test whether the Google Trends search quiries  can be used as a proxy of the popularity and public interest in different distance education solutions.
+In line with that, this paper makes three fundamental contributions to distance education by exploring three main questions: (1) What is the impact of the COVID-19 pandemic on education?
+(2) What solutions are in place to meet the needs of distance education during COVID-19 pandemic?
+(3) Which distance learning solutions have received wide attention and public interest during the COVID-19 pandemic?
 
 We primarily analyzed quantitative digital footprint data on the internet from December 2019 to August 2021.
-The resulted Google Trend footprint provides a fast first step guide to identify the most popular distance education tools available for different education purposes.
+The resulted Google Trend footprint provides a fast first step guide to identifying the most popular distance education tools available for different educational purposes.
 It also allows the teachers to narrow down the search space and  deepen their exploration on  prominent distance education solutions to support their online teaching.
-Further, identifying most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
+Furthermore, identifying the most popular tools and techniques will allow everyone to focus on those methods and develop a sustainable learning environment by advancing the existing features.
 
 Quality education is essential to sustainable development and it is a key United Nations (UN) sustainable development goal.
-Having the most appropriate tool is imperative to successful completion of a task at hand.
-The findings of this study provides an essential initial guidance to select the most appropriate tool for different distance education needs.
-Therefore, the findings of this study directly contribute to UN's sustainable development goals of quality education.
+Having the most appropriate tool is imperative to the successful completion of a task at hand.
+The findings of this study provide essential initial guidance to selecting the most appropriate tool for different distance education needs.
+Therefore, the findings of this study directly contribute to the UN's sustainable development goals of quality education.
 
-The remainder of this paper is organized as follows.
-Section 2 presents the related work to lay the foundation for the Google trend footprint analysis of different distance learning solutions.
+The remainder of this paper is organized as follows: Section 2 presents the related work to lay the foundation for the Google trend footprint analysis of different distance learning solutions.
 Section 3 presents the methodology followed in the study.

### 2021-12-27 Paper_online_Learning.tex
completed introduction

@@
 
-The term distance education is known by a variety of names such as distance learning, e-Learning, online learning, home study, independent study, external study, correspondence education, off campus study, open learning and open education \autocite{moore2011learning}.
-Few common features found in all these terms are, they all involve some form knowledge sharing between two parties; learners and educators, different times, different geographical locations and/or varying forms of instructional materials and tools \autocite{moore2011learning}.
-Further, distance education has different types of interactions such as learner--content interactions, learner--instructor interactions, learner--learner interactions and learner--interface interaction that can facilitate learners to construct knowledge \autocite{wallace2003online}.
+The term distance education is known by a variety of names, such as distance learning, e-Learning, online learning, home study, independent study, external study, correspondence education, off-campus study, open learning, and open education \autocite{moore2011learning}.
+A few common features found in all these terms are that they all involve some form of knowledge sharing between two parties; learners and educators, different times, different geographical locations and/or varying forms of instructional materials and tools \autocite{moore2011learning}.
+Further, distance education has different types of interactions, such as learner--content interactions, learner--instructor interactions, learner--learner interactions, and learner--interface interactions that can facilitate learners to construct knowledge \autocite{wallace2003online}.
 Different tools are required to fulfill all these requirements in a distance learning environment.
 
-Google Trends is a freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages \autocite{carneiro2009google}.
+Google Trends is a freely accessible website sponsored by Google that analyzes spatio-temporal patterns of search queries in Google Search, across various regions, subject domains, and languages \autocite{carneiro2009google}.
 Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution \autocite{alicino2015assessing,cook2011assessing}.
-According to \textcite{jarynowski2020perception} the public concerns on different matters can also take an epidemic nature starting with the phase of growing interest which is known as ``early adoption'', to the phase of general interest which is known as ``majority'' to eventually lose popularity which is known as ``lagers stage''.
+According to \textcite{jarynowski2020perception}, public concerns on different matters can also take on an epidemic nature, starting with the phase of growing interest, which is known as ``early adoption'', to the phase of general interest, which is known as ``majority'', and eventually lose popularity which is known as ``lagers stage''.
 Figure \ref{fig:distanceLearningWorldAnalysis} (b) and (c) also confirm this life cycle explanation of Google Trend search patterns.
 
-The use of Google Trend search queries has become a popular data source in monitoring and modeling the dynamics of different application domains such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding} and health care \autocite{nuti2014use,alicino2015assessing,arora2019google,carneiro2009google,cook2011assessing} as it has the ability to provide data about social phenomena in a more timely manner than traditional data collection processes \autocite{vaughan2014web}.
-According to the study conducted by \textcite{willis2020using} on taxi preferences found that online queries can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+Google Trend search queries have become a popular data source in monitoring and modeling the dynamics of various application domains, such as disaster management \autocite{kam2019monitoring}, transportation \autocite{willis2020using}, business \autocite{chumnumpan2019understanding}, and health care \autocite{nuti2014use,alicino2015assessing,arora2019google,carneiro2009google,cook2011assessing}, because they can provide data about social phenomena in a more timely manner than than traditional data collection processes \autocite{vaughan2014web}.
+
+According to the study conducted by \textcite{willis2020using} on taxi preferences, online queries can provide a proxy for the quality of taxi services and thereby provide a means to investigate service quality and public concerns about the service provided.
 In November, 2008, Google launched an internet-based surveillance tool, Google Flu Trends (GFT), that uses aggregated Google search data to provide a near real-time support in predicting influenza outbreaks \autocite{cook2011assessing}.
 Several recent studies have also been conducted to investigate the relation between Google search trend patterns and the COVID-19 related concerns \autocite{husnayain2020applications,effenberger2020association}.
-The study conducted by \textcite{husnayain2020applications} focuses on search terms related to the coronavirus, handwashing, and face masks to monitor public restlessness toward COVID-19.
+The study conducted by \textcite{husnayain2020applications} focuses on search terms related to the corona virus, hand washing, and face masks to monitor public restlessness toward COVID-19.
 However, there are only a limited number of research attempts related to education sector using Google Trend search query data \autocite{vaughan2014web}.
@@
 
-Eventhough Google Trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool as it may contain inaccuracies \autocite{carneiro2009google}.
+Even though Google Trends search queries provide a massive amount of data, more care should be given when utilizing it as a research tool as it may contain inaccuracies \autocite{carneiro2009google}.
 According to \textcite{nuti2014use}, the data retrieval process should be transparent.

### 2021-12-27 main.Rmd
completed introduction

@@
 
-The term distance education is known by a variety of names such as distance learning, e-Learning,  online learning, home study, independent study, external study, correspondence education, off campus study, open learning and open education [@moore2011learning].
-Few common features found in all these terms are, they all involve some form knowledge sharing between two parties; learners and educators, different times, different geographical locations  and/or  varying forms of instructional materials and tools [@moore2011learning].
-Further, distance education has different types  of interactions such as learner–content interactions, learner–instructor interactions, learner–learner interactions and  learner–interface interaction that can facilitate learners to construct knowledge  [@wallace2003online].
+The term distance education is known by a variety of names, such as distance learning, e-Learning,  online learning, home study, independent study, external study, correspondence education, off-campus study, open learning, and open education [@moore2011learning].
+A few common features found in all these terms are that they all involve some form of knowledge sharing between two parties; learners and educators, different times, different geographical locations  and/or  varying forms of instructional materials and tools [@moore2011learning].
+Further, distance education has different types  of interactions, such as learner–content interactions, learner–instructor interactions, learner–learner interactions, and  learner–interface interactions that can facilitate learners to construct knowledge  [@wallace2003online].
 Different tools are required to fulfill all these requirements in a distance learning environment.
 
-Google Trends is a  freely accessible website sponsored by Google that analyzes spatio and temporal patterns of search queries in Google Search, across various regions, subject domains and languages [@carneiro2009google].
+Google Trends is a freely accessible website sponsored by Google that analyzes spatio-temporal patterns of search queries in Google Search, across various regions, subject domains, and languages [@carneiro2009google].
 Therefore, Google trends provide a quantifiable and valuable measure of emerging public concerns and trending topics together with their geographic distribution  [@alicino2015assessing; @cook2011assessing].
-According to @jarynowski2020perception the public concerns on different matters can also take an epidemic nature  starting with the phase of growing interest which is known as "early adoption", to the phase of general interest which is known as "majority" to eventually lose popularity which is known as "lagers stage".
+According to @jarynowski2020perception, public concerns on different matters can also take on an epidemic nature, starting with the phase of growing interest, which is known as "early adoption", to the phase of general interest, which is known as "majority", and eventually lose popularity which is known as "lagers stage".
 Figure \ref{fig:distanceLearningWorldAnalysis} (b) and (c) also confirm this life cycle explanation of Google Trend search patterns.
 
-The use of Google Trend search queries has become a popular data source in monitoring and modeling the dynamics of different application domains such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding] and health care [@nuti2014use; @alicino2015assessing; @arora2019google; @carneiro2009google; @cook2011assessing] as it has the ability to provide data about social phenomena in a more timely manner than traditional data collection processes [@vaughan2014web].
-According to the study conducted by @willis2020using on taxi preferences found that online queries can provide a proxy of quality of taxi services and thereby provide a means to investigate service quality and public concerns on the service provided.
+Google Trend search queries have become a popular data source in monitoring and modeling the dynamics of various application domains, such as disaster management [@kam2019monitoring], transportation [@willis2020using], business [@chumnumpan2019understanding], and health care [@nuti2014use; @alicino2015assessing; @arora2019google; @carneiro2009google; @cook2011assessing], because they can provide data about social phenomena in a more timely manner than than traditional data collection processes [@vaughan2014web].
+
+According to the study conducted by @willis2020using on taxi preferences, online queries can provide a proxy for the quality of taxi services and thereby provide a means to investigate service quality and public concerns about the service provided.
 In November, 2008, Google launched an internet-based surveillance tool, Google Flu Trends  (GFT), that uses aggregated Google search data to provide a near real-time support in predicting influenza outbreaks [@cook2011assessing].
 Several recent studies have also been conducted  to investigate the relation between Google search trend patterns and the COVID-19 related concerns [@husnayain2020applications; @effenberger2020association].
-The study conducted by @husnayain2020applications focuses on search terms related to the coronavirus, handwashing, and face masks  to monitor public restlessness toward COVID-19.
+The study conducted by @husnayain2020applications focuses on search terms related to the corona virus, hand washing, and face masks  to monitor public restlessness toward COVID-19.
 However, there are only a limited number of research attempts related to education sector  using Google Trend search query data [@vaughan2014web].
@@
 
-Eventhough Google Trend search quiries provide massive amount of data, more care should be given when utilizing it as a research tool as it may contain inaccuracies [@carneiro2009google].
+Even though Google Trends search queries provide a massive amount of data, more care should be given when utilizing it as a research tool as it may contain inaccuracies [@carneiro2009google].
 According to @nuti2014use, the data retrieval process should be transparent.

### 2021-12-28 Paper_online_Learning.tex
updated methodology

@@
 
-In this work we also used Google Trend search quiries as a measurement proxy to investigate emerging evidence about distance learning and the popularity and the public interest in various distance education solutions.
-We used weekly Google Trend search quiries of various distance learning solutions during COVID-19 pandemic.
+In this work, we also used Google Trend search queries as a measurement proxy to investigate emerging evidence about distance learning and the popularity and public interest in various distance education solutions.
+We used weekly Google Trends search queries of various distance learning solutions during COVID-19 pandemic.
 We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the period of COVID-19 pandemic.
@@
 Google Trends determines categories based on search patterns.
-According to \textcite{vaughan2014web} more relevant accurate data can be obtained by limiting the search in Google Trends to a specific category as it helps to reduce noise in the data.
-In this work our focus was given to education category.
+According to \textcite{vaughan2014web} more relevant and accurate data can be obtained by limiting the search in Google Trends to a specific category as it helps to reduce noise in the data.
+In this work, our focus was given to the education category.
 Google Trends does not provide the absolute search volume of a given search term.
-Instead it provides a relative search volume of a particular term adjusted according to the total searches of the geography and time range it represents.
-The resulted series scale from 0 to 100, where each data point in the series represents the search interest relative to the highest point on the series for the selected region and time.
+Instead, it provides the relative search volume of a particular term, adjusted according to the total searches of the geography and time range it represents.
+The resulted series scales from 0 to 100, where each data point in the series represents the search interest relative to the highest point in the series for the selected region and time.
 \autocite{alicino2015assessing,vaughan2014web}.
@@
 It is important to decide what term to search for in Google Trends as different terms have different search volumes.
-In the past literature several strategies have been used to select suitable search terms for a given study.
+In the past, several strategies have been used to select suitable search terms for a given study.
 Some studies have selected search terms based on intuition and some through brainstorming processes \autocite{vaughan2014web}.
-In this study our choice of search term took into account the list of distance learning solutions published by UNESCO \autocite{unesco2020DLsolutions}.
-Eventhough, the solutions they had recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base and evidence of impact \autocite{unesco2020DLsolutions}.
-For each keyword, the worldwide search queries were performed with a keyword being used as the ``search term''.
-That allowed us to search the exact string of text typed by the user.
+In this study, our choice of search term took into account the list of distance learning solutions published by UNESCO \autocite{unesco2020DLsolutions}.
+Even though, the solutions they have recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base, and evidence of impact \autocite{unesco2020DLsolutions}.
+For each keyword, the worldwide search queries were performed with the keyword being used as the ``search term''.
+That allowed us to search for the exact string of text typed by the user.
 However, according to \textcite{vaughan2014web}, short forms or acronyms had higher search volumes than the corresponding complete names, in general.
-However, we did not use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
+However, we did not use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs, and acronyms could be confused with another entity.
 
-We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
-The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.
+We did our analysis under 11 sub-segments namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality, Massive Open Online Course (MOOC) platforms, self-directed learning content, mobile reading applications, collaboration platforms that support live-video communication, tools for teachers to create digital learning content, external repositories of distance learning solutions, tools for online proctoring, and resources to provide psychosocial support.
+The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to the COVID-19 pandemic.
 
-In google trends, it is possible to search for only up to five queries at a time \autocite{vaughan2014web}.
-Therefore, under each segmentation we entered up to five distance learning tools at a time and recorded their relative ranking scores.
-An iterative pairwise comparison was first used to identify the series that has highest search volume during the study period.
+In Google trends, it is possible to search for only up to five queries at a time \autocite{vaughan2014web}.
+Therefore, under each segmentation, we entered up to five distance learning tools at a time and recorded their relative ranking scores.
+An iterative pairwise comparison was first used to identify the series that had the highest search volume during the study period.
 Through this iterative process of relative comparison, we were able to identify the tool with the highest relative ranking score.
-Then the relative ranking scores for all the other tools in the given segmentation was obtained relative to the previously detected series with the highest ranking scores \autocite{vaughan2014web}.
+Then the relative ranking scores for all the other tools in the given segmentation were obtained relative to the previously detected series with the highest ranking scores \autocite{vaughan2014web}.
 
-Both visual and analytical approaches were used to analyse the changes in web search queries at the global level to identify the emerging evidence about the impact of COVID-19 pandemic on distance education.
-We also evaluated the relationship between weekly web searches and the number of weekly global cases of COVID-19.
+Both visual and analytical approaches were used to analyse the changes in web search queries at the global level to identify the emerging evidence about the impact of the COVID-19 pandemic on distance education.
+We also evaluated the relationship between weekly web searches and the number of global cases of COVID-19.
 Cross correlation analysis and the dynamic time warping algorithm \autocite{giorgino2009computing} were performed to investigate the relationship and similarities between the signatures of COVID-19 related search volume data and distance education related search volume data.

### 2021-12-28 main.Rmd
updated methodology

@@
 
-In this work we also used  Google Trend search quiries   as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and the public interest in various distance education solutions.
-We used weekly  Google Trend search quiries of various distance learning solutions during COVID-19 pandemic.
+In this work, we also used  Google Trend search queries   as a measurement proxy to investigate  emerging evidence about  distance learning and the popularity and public interest in various distance education solutions.
+We used weekly  Google Trends search queries of various distance learning solutions during COVID-19 pandemic.
 We limited all of our searches to the period from 1 December 2019 to 15 August 2021 to match the  period of COVID-19 pandemic.
@@
 Google Trends determines categories based on search patterns.
-According to @vaughan2014web  more relevant accurate data can be obtained by  limiting the search in Google Trends to a specific category as it helps  to reduce noise in the data.
-In this work our focus was given to education category.
+According to @vaughan2014web  more relevant and accurate data can be obtained by limiting the search in Google Trends to a specific category as it helps to reduce noise in the data.
+In this work, our focus was given to the education category.
 Google Trends does not provide the absolute search volume of a given search term.
-Instead it provides a relative search volume of a  particular term adjusted according to  the total searches of the geography and time range it represents.
-The resulted series  scale from 0 to 100, where each data point in the series represents the search interest relative to the highest point on the series for the selected region and time.
+Instead, it provides the relative search volume of a  particular term, adjusted according to  the total searches of the geography and time range it represents.
+The resulted series  scales from 0 to 100, where each data point in the series represents the search interest relative to the highest point in the series for the selected region and time.
 [@alicino2015assessing; @vaughan2014web].
@@
 It is important to decide what term to search for in Google Trends as different terms have different search volumes.
-In the past literature several strategies have been used to select suitable search terms for a given study.
+In the past, several strategies have been used to select suitable search terms for a given study.
 Some studies have  selected search terms based on intuition and some through  brainstorming processes [@vaughan2014web].
-In this study our choice of search term  took into account the list of distance learning solutions published by UNESCO [@unesco2020DLsolutions].
-Eventhough, the solutions they had recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base and evidence of impact [@unesco2020DLsolutions].
-For each keyword, the worldwide search queries were performed with a keyword being used  as the "search term".
-That allowed us to search the exact string of text typed by the user.
+In this study, our choice of search term  took into account the list of distance learning solutions published by UNESCO [@unesco2020DLsolutions].
+Even though, the solutions they have recommended do not carry their explicit endorsement, they tend to have a wide reach, a strong user-base, and evidence of impact [@unesco2020DLsolutions].
+For each keyword, the worldwide search queries were performed with the keyword being used  as the "search term".
+That allowed us to search for the exact string of text typed by the user.
 However, according to @vaughan2014web,  short forms or acronyms had higher search volumes than the corresponding complete names, in general.
-However, we did not  use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs and acronyms could be confused with another entity.
+However, we did not  use the acronym in this study, as our focus is on specific tools and techniques available in the market to fulfill distance education needs, and acronyms could be confused with another entity.
 
-We did our analysis under 11 sub segmentation namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) Platforms, self-directed learning content, mobile reading applications, collaboration platforms that support live-video communication, tools for teachers to create of digital learning content, external repositories of distance learning solutions, tools for online proctoring and Resources to provide psychosocial support.
-The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to COVID-19 pandemic.
+We did our analysis under 11 sub-segments namely, digital learning management systems, systems built for use on basic mobile phones, systems with strong offline functionality,  Massive Open Online Course (MOOC) platforms, self-directed learning content, mobile reading applications, collaboration platforms that support live-video communication, tools for teachers to create digital learning content, external repositories of distance learning solutions, tools for online proctoring, and resources to provide psychosocial support.
+The segmentation was mainly motivated by the list of distance learning solutions published by UNESCO in response to the COVID-19 pandemic.
 
-In google trends, it is possible to search for only up to five queries at a  time [@vaughan2014web].
-Therefore, under each segmentation we entered up to five distance learning tools at a time and recorded their relative ranking scores.
-An iterative pairwise comparison was first used to identify the series that has highest search volume during the study period.
+In Google trends, it is possible to search for only up to five queries at a  time [@vaughan2014web].
+Therefore, under each segmentation, we entered up to five distance learning tools at a time and recorded their relative ranking scores.
+An iterative pairwise comparison was first used to identify the series that had the highest search volume during the study period.
 Through this iterative process of relative comparison, we were able to identify the tool with the highest relative ranking score.
-Then the relative ranking scores for all the other tools in the given segmentation was obtained relative to the previously detected series with the highest ranking scores [@vaughan2014web].
+Then the relative ranking scores for all the other tools in the given segmentation were obtained relative to the previously detected series with the highest ranking scores [@vaughan2014web].
 
-Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level to identify the emerging evidence about the impact of  COVID-19 pandemic on distance education.
-We also evaluated the  relationship between weekly web searches and the number of  weekly global cases of COVID-19.
+Both visual and analytical approaches were used to  analyse the changes in web search queries at the global level to identify the emerging evidence about the impact of  the COVID-19 pandemic on distance education.
+We also evaluated the  relationship between weekly web searches and the number of global cases of COVID-19.
 Cross correlation analysis and  the dynamic time warping algorithm [@giorgino2009computing] were performed  to investigate the relationship and similarities between the signatures of  COVID-19 related search volume data  and distance education related search volume data.

### 2021-12-28 Paper_online_Learning.tex
working on results

@@
 
-The visual representation of Google Trend search quiries of COVID-19 related terms worldwide shows that the highest peak coincided with the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (b)).
-Surprisingly, the search volume index graph for distanced education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
-This highlights the impact of COVID-19 pandemic on distance education.
+The visual representation of Google Trend search queries of COVID-19 related terms worldwide shows that the highest peak coincided with the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (b)).
+Surprisingly, the search volume index graph for distance education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
+This highlights the impact of the COVID-19 pandemic on distance education.
 
-According to \textcite{moore2011learning}, ``distance education'' is the most renowned descriptor used when referencing distance learning.
-However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version'', ``newer version'' and ``improved version'' of distance learning.
-The Google Trend search patterns corresponds to the term ``online learning'' confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
-Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``Distance learning'' and ``online teaching'' demonstrate a similar trend during the study period.
+According to \textcite{moore2011learning}, ``distance education'' is the most renowned term used when referencing distance learning.
+However, both \textcite{benson2002usability} and \textcite{conrad2002deep} introduced ``online learning'' as the most ``recent version,'' ``newer version'' and ``improved version'' of distance learning.
+The Google Trend search patterns correspond to the term ``online learning'' confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
+Despite quantitative differences in terms of relative search volume, the series generated by the two terms ``distance learning'' and ``online teaching'' demonstrate a similar trend during the study period.
 COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases.
-A similar decline can be observed in distance education related terms (Figures \ref{fig:distanceLearningWorldAnalysis}).
+A similar decline can be observed in distance education-related terms (Figures \ref{fig:distanceLearningWorldAnalysis}).
 This confirms the life cycle of Google Trend search patterns introduced in \textcite{jarynowski2020perception}.
 Further, a second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
-This could be due to the near total closures of schools happened with the declaration of COVID-19 as a pandemic.
+This could be due to the nearly total closures of schools that happened with the declaration of COVID-19 as a pandemic.
 
@@
 
-A pairwise comparison was performed using Dynamic Time Warping algorithm.
+A pairwise comparison was performed using the Dynamic Time Warping algorithm.
 Under this approach, two time series are stretched or compressed locally in order to make one resemble the other as much as possible.
@@
 Two pairwise comparisons, ``corona'' with ``online learning'' and ``covid'' with ``online learning'' were conducted to see the impact of COVID-19 on distance education.
-Both pairwise comparisons show a significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)).
-The normalized distance between the two series in the two comparisons are 0.1326197 and 0.1828299, respectively.
-This also provides further confirmation for the similarity of the Google Trend search profiles.
-this highlight the sudden public attention towards both aspects, COVID-19 and distance education, eventhough distance education has a long history in contrast to COVID-19.
+Both pairwise comparisons show significant areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and \ref{fig:dtw2} (a)).
+The normalized distance between the two series in the two comparisons is 0.1326197 and 0.1828299, respectively.
+This also provides further confirmation of the similarity of the Google Trend search profiles.
+This highlights the sudden public attention towards both aspects, COVID-19 and distance education, even though distance education has a long history in contrast to COVID-19.
 
-The entirety of the blue warping path flows through green pastures further confirms the similarity between the search patterns related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after the COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The entirety of the blue warping path flows through green pastures, further confirming the similarity between the search patterns related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of the series in question if mapped through this region demonstrates a significant difference in online search behaviour before and after the COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
 
-The degree of digression of blue line from the ideal 45-degree straight line indicates a dissimilarity of the two time
+The degree of digression of the blue line from the ideal 45-degree straight line indicates a dissimilarity between the two time
 series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-In late 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was know as ``corona'' it was not a global concern and was mostly limited to China.
-During that time people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic.
+In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was known as ``corona'', it was not a global concern and was mostly limited to China.
+During that time, people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic.
 This is evident from the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).
 
-Later, with the spread of the virus worldwide, WHO named it as COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
-With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
-With this sudden unexpected massive shutdown, online learning also became an equally important global concern, similar to COVID-19 pandemic.
-This similar behaviour in public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line from approximately 120 to 150 on the Query Index axis in Figure \ref{fig:dtw2} (b).
+Later, with the spread of the virus worldwide, WHO named it COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it a pandemic on March 11, 2020.
+With that, a near total closure of schools was observed all around the world (Figures \ref{fig:covidImpactWorld}).
+With this sudden and unexpected massive shutdown, online learning also became an equally important global concern, similar to the COVID-19 pandemic.
+This similar behaviour in public attention towards both aspects is confirmed by the blue line that falls on a near 45-degree line from approximately 120 to 150 on the Query Index axis in Figure \ref{fig:dtw2} (b).
 This also confirms the impact of COVID-19 on distance education and online learning.
@@
 Next, a cross correlation analysis was performed to estimate the lag time between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
-A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms ``Online learning'', ``Distance learning'' and ``Online teaching''.
-The significant cross-correlation coefficients at lag zero confirms contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms.
-The significant cross correlation values at lag one or two also confirms the sudden impact of COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
+A close relationship was evident between COVID-19 related search patterns and the search patterns for the terms ``Online learning'', ``Distance learning'' and ``Online teaching''.
+The significant cross-correlation coefficients at lag zero confirm the contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms.
+The significant cross correlation values at lag one or two also confirm the sudden impact of the COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
 However, no significant synchronization was observed between the COVID-19 related search trends and the search patterns of ``distance education'' or ``online proctoring'' (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).

### 2021-12-28 main.Rmd
working on results

@@
 
-The visual representation of Google Trend search quiries of COVID-19 related terms worldwide shows that the highest peak coincided with the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (b)).
-Surprisingly, the search volume index graph for distanced education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
-This highlights the impact of COVID-19 pandemic on distance education.
+The visual representation of Google Trend search queries of COVID-19 related terms worldwide shows that the highest peak coincided with the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (b)).
+Surprisingly, the search volume index graph for distance education related terms also shows a similar spike at the start of the pandemic (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
+This highlights the impact of the COVID-19 pandemic on distance education.
 
-According to @moore2011learning, "distance education" is the most renowned descriptor used when referencing distance learning.
-However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version", "newer version" and "improved version" of distance learning.
-The Google Trend search patterns corresponds to the term "online learning" confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
-Despite quantitative differences in terms of relative search volume, the series generated by the two terms "Distance learning" and "online teaching" demonstrate a similar trend during the study period.
+According to @moore2011learning, "distance education" is the most renowned term used when referencing distance learning.
+However, both  @benson2002usability and @conrad2002deep introduced "online learning"  as the most "recent version," "newer version" and "improved version" of distance learning.
+The Google Trend search patterns correspond to the term "online learning" confirm the above claim as it is the most searched distance education related term throughout the COVID-19 pandemic  (Figures \ref{fig:distanceLearningWorldAnalysis} (c)).
+Despite quantitative differences in terms of relative search volume, the series generated by the two terms "distance learning" and "online teaching" demonstrate a similar trend during the study period.
 COVID-19 related web searches declined in the months after the WHO announcement of the COVID-19 pandemic, despite the increase in COVID-19 cases.
-A similar decline can be observed in distance education related terms  (Figures \ref{fig:distanceLearningWorldAnalysis}).
+A similar decline can be observed in distance education-related terms  (Figures \ref{fig:distanceLearningWorldAnalysis}).
 This confirms the life cycle of Google Trend search patterns introduced in @jarynowski2020perception.
 Further,  a second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
-This could be due to the near total closures of schools happened  with the declaration of  COVID-19 as a pandemic.
+This could be due to the nearly total closures of schools that happened  with the declaration of  COVID-19 as a pandemic.
 
@@
 
-A pairwise comparison was performed  using  Dynamic Time Warping algorithm.
+A pairwise comparison was performed  using the Dynamic Time Warping algorithm.
 Under this approach, two time series are stretched or compressed  locally in order to make one resemble the other as much as possible.
@@
 Two pairwise comparisons, "corona"  with "online learning" and   "covid" with "online learning" were conducted to see the impact  of COVID-19 on distance education.
-Both pairwise comparisons show a significant  areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and  \ref{fig:dtw2} (a)).
-The normalized distance between the two series in the two comparisons are  0.1326197 and 0.1828299, respectively.
-This also provides further confirmation for  the similarity of the Google Trend  search profiles.
-this highlight the sudden public attention towards both aspects, COVID-19 and distance education, eventhough distance education has a long history in contrast to COVID-19.
+Both pairwise comparisons show significant  areas of overlap and a similar dynamic pattern (Figures \ref{fig:dtw1} (a) and  \ref{fig:dtw2} (a)).
+The normalized distance between the two series in the two comparisons is  0.1326197 and 0.1828299, respectively.
+This also provides further confirmation of  the similarity of the Google Trend  search profiles.
+This highlights the sudden public attention towards both aspects, COVID-19 and distance education, even though distance education has a long history in contrast to COVID-19.
 
-The entirety of the blue warping path flows through green pastures further confirms the similarity between the search patterns related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-The orange volcanic topography that highlights regions of nonalignment of series in question if mapped through this region demonstrate a significant difference in online search behaviours before and after the COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The entirety of the blue warping path flows through green pastures, further confirming the similarity between the search patterns related to COVID-19 and distance education (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
+The orange volcanic topography that highlights regions of nonalignment of the series in question if mapped through this region demonstrates a significant difference in online search behaviour before and after the COVID-19 pandemic (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
 
-The degree of digression of blue line from the ideal 45-degree straight line indicates a dissimilarity of the two time
+The degree of digression of the blue line from the ideal 45-degree straight line indicates a dissimilarity between the two time
 series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-In late 2019 the infection was recognized as a virus which is genetically related to the corona virus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003 which infected over 8,000 people from different countries that resulted more than 700 deaths worldwide [@ciotti2019covid].When the virus was know as "corona" it was not a global concern and was mostly limited to China.
-During that time people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to COVID-19 pandemic.
+In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide [@ciotti2019covid].When the virus was known as "corona", it was not a global concern and was mostly limited to China.
+During that time, people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic.
 This is evident from  the rough flat line from approximately 120 to 170 on the Query Index axis in Figure \ref{fig:dtw1} (b).
 
-Later, with the spread of the virus worldwide, WHO  named it as  COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it as a pandemic on March 11, 2020.
-With that, a near total closure of schools were observed all around the world (Figures \ref{fig:covidImpactWorld}).
-With this sudden unexpected massive shutdown, online learning  also became an equally important global concern, similar to COVID-19 pandemic.
-This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that falls on near 45-degree line  from approximately 120 to 150 on the Query Index axis in Figure \ref{fig:dtw2} (b).
+Later, with the spread of the virus worldwide, WHO  named it COVID-19 (COrona VIrus Diagnosed in 2019) and then declared it a pandemic on March 11, 2020.
+With that, a near total closure of schools was observed all around the world (Figures \ref{fig:covidImpactWorld}).
+With this sudden and unexpected massive shutdown, online learning  also became an equally important global concern, similar to the COVID-19 pandemic.
+This similar  behaviour in  public attention towards both aspects is confirmed by the blue line that falls on a near 45-degree line  from approximately 120 to 150 on the Query Index axis in Figure \ref{fig:dtw2} (b).
 This also confirms the impact of COVID-19 on distance education and online learning.
@@
 Next, a cross correlation analysis was performed to estimate the lag time between COVID-19 related terms and the distance education related terms (Figures \ref{fig:ccfAnalysis}).
-A close relationship was evident between COVID-19 related search patterns and the search patterns of the terms "Online learning", "Distance learning" and "Online teaching".
-The significant cross-correlation coefficients  at lag zero confirms  contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms.
-The significant cross correlation values at lag one or two also confirms the sudden impact of COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
+A close relationship was evident between COVID-19 related search patterns and the search patterns for the terms "Online learning", "Distance learning" and "Online teaching".
+The significant cross-correlation coefficients  at lag zero confirm the contemporaneous relationship between the COVID-19 related search terms and distance learning related search terms.
+The significant cross correlation values at lag one or two also confirm the sudden impact of the COVID-19 pandemic on distance education related public concerns (Figures \ref{fig:ccfAnalysis} (a) - (c) and (f) - (h)).
 However, no significant  synchronization was observed  between the COVID-19 related search trends and the search patterns of "distance education" or "online proctoring" (Figures \ref{fig:ccfAnalysis} (d), (i), (e), (j)).

### 2021-12-28 Paper_online_Learning.tex
completed results section

@@
 
-The list of distance education solutions published by UNESCO provides many different distance education solutions for different education purposes.
+The list of distance education solutions published by UNESCO provides many different distance education solutions for different educational purposes.
 Figure \ref{fig:plot1} and Figure \ref{fig:plot2} provide a Google trend footprint that can be used as a proxy for measuring global attention towards different distance learning solutions.
 This provides a fast first step guide to identify the most popular tools available for different educational purposes.
-Under each and every panel, one or two tools have got a significantly higher level of attention in comparison to the remaining options available under the same category (Figure \ref{fig:plot1} (a) - (e) and Figure \ref{fig:plot2} (f) - (j)).
-However, the Google Trend series across multiple panels are not comparable as they represent ``relative'' search volume of given terms.
+In each panel, one or two tools have received significantly more attention than the remaining options in the same category (Figure \ref{fig:plot1} (a) - (e) and Figure \ref{fig:plot2} (f) - (j)).
+However, the Google Trend series across multiple panels is not comparable as they represent ``relative'' search volume for given terms.
 They are comparable only within a given panel.
-The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each panel.
+The value of 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each panel.
 
-Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during COVID-19 pandemic (Figure \ref{fig:plot1} (a)).
-This might be due to it's easy and fast setup and usage \autocite{sudarsana2019use}.
-Under systems built for use on basic mobile phones, both `kaiOS' and `Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
-`KaiOs' has the ability to provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
-`Ubongo' has also received a considerable attention, eventhough its target groups are school-age children and their parents in Africa.
-It facilitates localized, multi-platform Entertainment-Education to African families at low cost and massive scale and is available in both Kiswahili and English.
-A significant peak in `Ubongo' series can be observed during 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
-Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
-These distance learning solutions can provide a huge relief for financially distressed families during COVID-19 pandemic.
-Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
+Under digital learning management systems ``Google Classroom'', which helps teachers and students connect remotely, communicate, and stay-organized, has gotten a wide attention and public interest during the COVID-19 pandemic (Figure \ref{fig:plot1} (a)).
+This might be due to its easy and fast setup and usage \autocite{sudarsana2019use}.
+Under systems built for use on basic mobile phones, both `kaiOS' and `Ubongo' have received wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
+`KaiOs' has the ability to provide smartphone capabilities to inexpensive, affordable devices and thereby help open portals to learning opportunities.
+`Ubongo' has also received considerable attention, even though its target groups are school-age children and their parents in Africa.
+It facilitates localized, multi-platform entertainment-education to African families at a low cost and on a massive scale, and is available in both Kiswahili and English.
+A significant peak in `Ubongo' series can be observed during the 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
+Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic.
+Under Massive Open Online Course (MOOC) platforms, `Canvas', which provides free access for teachers to facilitate learning and professional development, has attracted considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
 `Udemy' has got the second highest attention.
 However, `Canvas' has received far more attention than `Udemy'.
-Under self-directed learning content `Quizlet' has got a considerably high attention throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)).
-It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and languages.
+Under self-directed learning content, `Quizlet' has received considerably high attention throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)).
+It provides learning flashcards, an AI based learning assistant, and games to support student learning in multiple subjects and languages.
 The second highest option only provides support for language learning.
-Quizlet's ability to provide support for multiple subject might be the reason for it to outperform its competitor in the market.
-Under mobile reading applications, `Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
-Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
-However, `WhatsApp' has outperformed `Zoom' with increasing public attention during the pandemic.
-In contrary, the global attention towards `Zoom' has decreased over time.
-`Teams' has also got considerable attention in comparison to the other remaining options available in the market.
-Under tools for teachers to create of digital learning content, `Trello' which provides a visual collaboration platform for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
-`Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).
+Quizlet's ability to provide support for multiple subjects might be the reason for its ability to outperform its competitors in the market.
+Under mobile reading applications, `Reads', which provides digital stories with illustrations in multiple languages has consistently been the most popular option (Figure \ref{fig:plot2} (f)).
+Under collaboration platforms that support live-video communication, both `WhatsApp' and `Zoom' had received a lot of attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
+However, `WhatsApp' has outperformed `Zoom' in attracting public attention during the pandemic.
+In contrast, global attention towards `Zoom' has decreased over time.
+`Teams' has also gotten considerable attention in comparison to the other remaining options available in the market.
+During the pandemic, tools for teachers to create digital learning content, such as ``Trello,'' which provides a visual collaboration platform for teachers for coursework planning, classroom organization, and collaboration, received a lot of attention.
+`Nearpod', which promotes active learning and student engagement through informative and interactive learning activities has also gotten considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).
 A significant peak can be observed in the `Nearpod' series, during the first academic break.
-Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got a significant global attention (Figure \ref{fig:plot2} (i)).
-`Brookings' provides a large catalogue of learning innovations eventhough its focus is not limited to distance learning solutions.
+Under external repositories of distance learning solutions, both `Brookings' and `UNHCR' have got significant global attention (Figure \ref{fig:plot2} (i)).
+`Brookings' provides a large catalogue of learning innovations, though its focus is not limited to distance learning solutions.
 `UNHCR' also provides an extensive list of distance learning solutions from the United Nations agency for refugees.
 Student assessment is vital in ensuring the quality of education.
-However, due unexpected sudden school closure, administering examinations at a distance became a major concern.
+However, due to an unexpected and sudden school closure, administering examinations at a distance became a major concern.
 Under tools for online proctoring, `Pearson VUE' which provides computer based testing services in
-over 150 countries worldwide with more than 5,000 authorized test centers has got a considerably high public attention during the pandemic (Figure \ref{fig:plot2} (j)).
+over 150 countries worldwide with more than 5,000 authorized test centers has attracted considerably high public attention during the pandemic (Figure \ref{fig:plot2} (j)).
 
 At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both digital learning management systems (Figure \ref{fig:plot1} (a)) and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
-These peaks aligns with the near-total closures of schools due to COVID-19 pandemic (Figure \ref{fig:covidImpactWorld}).
-The major concerns that arise with the unexpected sudden movement from physical classroom to virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
-Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
-The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the need of urgent responses to connect with students to somehow facilitate their learning could be the reasons for this considerable public attention around August, 2020.
+These peaks align with the near-total closures of schools due to the COVID-19 pandemic (Figure \ref{fig:covidImpactWorld}).
+The major concerns that arise with the unexpected, sudden movement from a physical classroom to a virtual learning environment, such as a lack of experience in connecting and sharing learning materials with students could be the reason for this notable public attention towards these two aspects.
+Further, a significant peak can be observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020 (Figure \ref{fig:covidImpactWorld}).
+The inability to reopen schools fully due to the increasing COVID-19 cases worldwide and the need for urgent responses to connect with students to somehow facilitate their learning could be the reasons for this considerable public attention around August, 2020.
 

### 2021-12-28 main.Rmd
completed results section

@@
 
-The list of distance education solutions published by UNESCO provides many different distance education solutions for different education purposes.
+The list of distance education solutions published by UNESCO provides many different distance education solutions for different educational purposes.
 Figure \ref{fig:plot1} and  Figure \ref{fig:plot2} provide a Google trend footprint that can be used as a proxy for measuring global attention towards different distance learning solutions.
 This provides a fast first step guide to identify the most popular tools available for different educational purposes.
-Under each and every panel, one or two  tools have got a significantly higher level of attention in comparison to the remaining options available under the same category (Figure \ref{fig:plot1} (a) - (e) and  Figure \ref{fig:plot2} (f) - (j)).
-However, the Google Trend series across multiple panels are not comparable as they represent "relative" search volume of given terms.
+In each panel, one or two tools have received significantly more attention than the remaining options in the same category (Figure \ref{fig:plot1} (a) - (e) and  Figure \ref{fig:plot2} (f) - (j)).
+However, the Google Trend series across multiple panels is not comparable as they represent "relative" search volume for given terms.
 They are comparable only within a given panel.
-The value 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each panel.
+The value of 100 is also not comparable across multiple panels and indicates the maximum popularity for the term within each panel.
 
-Under digital learning management systems "Google Classroom", which  helps  teachers and students connect remotely, communicate and stay-organized, has got a wide attention and public interest during COVID-19 pandemic (Figure \ref{fig:plot1} (a)).
-This might be due to it's easy and fast setup and usage [@sudarsana2019use].
-Under systems built for use on basic mobile phones, both 'kaiOS' and  'Ubongo' have received a wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
-'KaiOs'  has the ability to provide smartphone capabilities to inexpensive affordable devices and thereby helps open portals to learning opportunities.
-'Ubongo' has also received a considerable attention, eventhough its target groups are school-age children and their parents in Africa.
-It facilitates localized, multi-platform Entertainment-Education  to African families at low cost and massive scale and is  available in both Kiswahili and English.
-A significant peak in  'Ubongo' series can be observed during 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
-Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, supporting universal education, has got a considerable attention during COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
-These distance learning solutions can provide a huge relief for financially distressed families during COVID-19 pandemic.
-Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has got a considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
+Under digital learning management systems "Google Classroom", which  helps  teachers and students connect remotely, communicate, and stay-organized, has gotten a wide attention and public interest during the COVID-19 pandemic (Figure \ref{fig:plot1} (a)).
+This might be due to its easy and fast setup and usage [@sudarsana2019use].
+Under systems built for use on basic mobile phones, both 'kaiOS' and  'Ubongo' have received wide attention during the pandemic (Figure \ref{fig:plot1} (b)).
+'KaiOs'  has the ability to provide smartphone capabilities to inexpensive, affordable devices and thereby help open portals to learning opportunities.
+'Ubongo' has also received considerable attention, even though its target groups are school-age children and their parents in Africa.
+It facilitates localized, multi-platform entertainment-education  to African families at a low cost and on a massive scale, and is  available in both Kiswahili and English.
+A significant peak in  'Ubongo' series can be observed during the 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
+Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic.
+Under Massive Open Online Course (MOOC) platforms, 'Canvas', which provides free access for teachers to  facilitate  learning and professional development, has attracted considerable attention in comparison to other options available (Figure \ref{fig:plot1} (d)).
 'Udemy' has got the second highest attention.
 However, 'Canvas' has received far more attention than 'Udemy'.
-Under self-directed learning content 'Quizlet' has got a considerably high attention  throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)).
-It provides learning flashcards, AI based learning assistant and games to support student learning in multiple subjects and languages.
+Under self-directed learning content, 'Quizlet' has received considerably high attention  throughout the pandemic in comparison to the other options available under the same category (Figure \ref{fig:plot1} (e)).
+It provides learning flashcards, an AI based learning assistant, and games to support student learning in multiple subjects and languages.
 The second highest option only provides  support for language learning.
-Quizlet's ability to provide support for multiple subject might be the reason for it to outperform its  competitor in the market.
-Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently being the most popular option (Figure \ref{fig:plot2} (f)).
-Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had got a considerable attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
-However, 'WhatsApp' has outperformed 'Zoom'  with increasing public attention during the pandemic.
-In contrary, the global attention towards 'Zoom'  has decreased over time.
-'Teams' has also got considerable attention in comparison to the other remaining options available in the market.
-Under tools for teachers to create of digital learning content, 'Trello'  which provides a visual collaboration platform  for teachers for coursework planning, classroom organization and collaboration, has received a wide attention during the pandemic.
-'Nearpod', that promotes active learning and student engagement through informative and interactive learning activities has also got a considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).
+Quizlet's ability to provide support for multiple subjects might be the reason for its ability to outperform its  competitors in the market.
+Under mobile reading applications, 'Reads', which provides digital stories with illustrations in multiple languages has consistently been the most popular option (Figure \ref{fig:plot2} (f)).
+Under collaboration platforms that support live-video communication, both 'WhatsApp' and 'Zoom' had received a lot of attention at the start of the pandemic in March 2020 (Figure \ref{fig:plot2} (g)).
+However, 'WhatsApp' has outperformed 'Zoom'  in attracting public attention during the pandemic.
+In contrast, global attention towards 'Zoom'  has decreased over time.
+'Teams' has also gotten considerable attention in comparison to the other remaining options available in the market.
+During the pandemic, tools for teachers to create digital learning content, such as "Trello," which provides a visual collaboration platform for teachers for coursework planning, classroom organization, and collaboration, received a lot of attention.
+'Nearpod', which promotes active learning and student engagement through informative and interactive learning activities has also gotten considerable attention during the pandemic (Figure \ref{fig:plot2} (h)).
 A significant peak can be observed in the 'Nearpod' series, during the first academic break.
-Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got a significant global attention (Figure \ref{fig:plot2} (i)).
-'Brookings' provides a large catalogue of learning innovations eventhough its focus is not limited to distance learning solutions.
+Under external repositories of distance learning solutions, both 'Brookings' and 'UNHCR' have got significant global attention (Figure \ref{fig:plot2} (i)).
+'Brookings' provides a large catalogue of learning innovations, though its focus is not limited to distance learning solutions.
 'UNHCR' also  provides an extensive list of distance learning solutions from the United Nations agency for refugees.
 Student assessment is vital in ensuring the quality of education.
-However, due unexpected sudden school closure, administering examinations at a distance became a major concern.
+However, due to an unexpected and sudden school closure, administering examinations at a distance became a major concern.
 Under tools for online proctoring, 'Pearson VUE' which provides computer based testing services in
-over 150 countries worldwide with more than 5,000 authorized test centers has got  a considerably high public attention during the pandemic (Figure \ref{fig:plot2} (j)).
+over 150 countries worldwide with more than 5,000 authorized test centers has attracted considerably high public attention during the pandemic (Figure \ref{fig:plot2} (j)).
 
 At the start of the pandemic in March 2020, a significant peak of public attention can be observed in both digital learning management systems (Figure \ref{fig:plot1} (a))  and collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)).
-These peaks aligns with the near-total closures of schools due to COVID-19 pandemic (Figure \ref{fig:covidImpactWorld}).
-The major concerns that arise with the unexpected sudden movement from physical classroom to  virtual learning environment, such as lack of experience to connect and share learning materials with students could be the reasons for this notable public attention towards these two aspects.
-Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), Self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
-The inability to reopen school fully due to the increasing COVID-19 cases worldwide and the  need of urgent responses to connect with students to somehow facilitate their learning  could be the reasons for this considerable public attention around August, 2020.
+These peaks align with the near-total closures of schools due to the COVID-19 pandemic (Figure \ref{fig:covidImpactWorld}).
+The major concerns that arise with the unexpected, sudden movement from a physical classroom to a virtual learning environment, such as a lack of experience in connecting and sharing learning materials with students could be the reason for this notable public attention towards these two aspects.
+Further, a significant peak can be  observed in digital learning management systems (Figure \ref{fig:plot1} (a)), Massive Open Online Course (MOOC) Platforms (Figure \ref{fig:plot1} (d)), self-directed learning content (Figure \ref{fig:plot1} (e)), collaboration platforms that support live-video communication (Figure \ref{fig:plot2} (g)) and tools for teachers to create digital learning content (Figure \ref{fig:plot2} (h)), immediately after the academic break in August, 2020  (Figure \ref{fig:covidImpactWorld}).
+The inability to reopen schools fully due to the increasing COVID-19 cases worldwide and the  need for urgent responses to connect with students to somehow facilitate their learning  could be the reasons for this considerable public attention around August, 2020.
 
@@
 
-Good psychosocial support is also vital to improve both students' and  teachers' mental health and well-being during a pandemic.
-This is evident from the two sudden spikes in the search volume index graphs of the term 'Psychosocial support' ( Figure \ref{fig:psychosocialSupport} ) which correspond well with  the academic breaks in August in both years, 2020 and 2021, under 'Education' category.
-This highlight the urgent need of attention and solutions towards these aspects.
-This type of an analysis based on secondary data is vital as these aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
+Good psychosocial support is also vital to improving both students' and  teachers' mental health and well-being during a pandemic.
+This is evident from the two sudden spikes in the search volume index graphs for the term 'Psychosocial support' ( Figure \ref{fig:psychosocialSupport} ) which correspond well with  the academic breaks in August in both years, 2020 and 2021, under 'Education' category.
+This highlights the urgent need for attention and solutions to these aspects.
+This type of analysis based on secondary data is vital as these aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
 

### 2021-12-28 Paper_online_Learning.tex
updated

@@
 
-Good psychosocial support is also vital to improve both students' and teachers' mental health and well-being during a pandemic.
-This is evident from the two sudden spikes in the search volume index graphs of the term `Psychosocial support' ( Figure \ref{fig:psychosocialSupport} ) which correspond well with the academic breaks in August in both years, 2020 and 2021, under `Education' category.
-This highlight the urgent need of attention and solutions towards these aspects.
-This type of an analysis based on secondary data is vital as these aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
+Good psychosocial support is also vital to improving both students' and teachers' mental health and well-being during a pandemic.
+This is evident from the two sudden spikes in the search volume index graphs for the term `Psychosocial support' ( Figure \ref{fig:psychosocialSupport} ) which correspond well with the academic breaks in August in both years, 2020 and 2021, under `Education' category.
+This highlights the urgent need for attention and solutions to these aspects.
+This type of analysis based on secondary data is vital as these aspects can be easily overlooked due to the lack of connection and communication between students and educators during school closures.
 
@@
 
-In this study we investigated whether the Google Trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
-This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic.
-A few previous published studies \autocite{vaughan2014web,kansal2021google} have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
+In this study, we investigated whether the Google Trend search queries can be used as a proxy of the popularity and public interest in different distance education solutions.
+This study is the first detailed analysis of web search behaviour related to distance learning solutions during the COVID-19 pandemic, both in quantitative and qualitative terms.
+A few previous published studies \autocite{vaughan2014web,kansal2021google} looked at Google trend search queries related to education, but only with a limited number of search terms or with different focuses.
 
@@
 Google Trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
-The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed Google Trend footprint analysis.
-The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
+The high popularity and increasing public attention towards different distance learning solutions during the COVID-19 pandemic was also investigated through a detailed Google Trend footprint analysis.
+The findings will help educators to narrow down their search space and thereby select prominent options available in the market for different educational purposes.
 According to \textcite{wallace2003online}, the learning tools can have a significant impact on how students engage in online learning.
-The finding will also helpful for developers to identify their competitors in the market.
-High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the COVID-19 pandemic.
-Therefore, the findings of this study offer important insights to various stakeholders of education to make better decision with less time and effort.
+The finding will also be helpful for developers to identify their competitors in the market.
+The high costs involved with different distance learning solutions, lack of financial support/funding, and lack of understanding about the existing distance learning solutions are some of the major issues educational institutions face with the COVID-19 pandemic.
+Therefore, the findings of this study offer important insights to various stakeholders in education to make better decisions with less time and effort.
 
-Google Trends only captures the search behaviour of a sub population with Internet access, eventhough it is the most common worldwide search engine.
-Further, our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
+Google Trends only captures the search behaviour of a sub-population with Internet access, even though it is the most common worldwide search engine.
+Furthermore, our study focused on global attention toward distance learning solutions during the COVID-19 pandemic.
 Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.
 This can lead to variations in the search output and in turn the study findings for different regions and time frames.
-Further, Google Trends data were not available as absolute values and available only in the form of relative volume.
-Therfore the scores ranges from 0 to 100 do not allow the researchers to make meaningful comparisons across different education needs.
+Furthermore, Google Trends data is not available as absolute values but only in the form of relative volume.
+Therefore, the scores range from 0 to 100, which does not allow the researchers to make meaningful comparisons across different education needs.
 Further investigation is needed at national and regional level to identify the specific measures that can be taken to ensure the quality of distance education at national and regional level.
@@
 
-This work was carried out with the aid of a grant form UNESCO and the International Development Research Centre, Ottawa, Canada.
+This work was carried out with the aid of a grant from UNESCO and the International Development Research Centre, Ottawa, Canada.
 The views expressed herein do not necessarily represent those of UNESCO, IDRC or its Board of Governors.

### 2021-12-28 main.Rmd
updated

@@
 
-In this study  we investigated whether the Google  Trend search queries can be used as a proxy of the popularity and the public interest in different distance education solutions.
-This study is the first detailed analysis of web search behaviour related to distance learning solutions, both in quantitative and qualitative terms during COVID-19 pandemic.
-A few previous published studies [@vaughan2014web; @kansal2021google] have investigated Google trend search queries related to education, but only with a limited number of search terms or different focuses.
+In this study, we investigated whether the Google  Trend search queries can be used as a proxy of the popularity and public interest in different distance education solutions.
+This study is the first detailed analysis of web search behaviour related to distance learning solutions during the COVID-19 pandemic, both in quantitative and qualitative terms.
+A few previous published studies [@vaughan2014web; @kansal2021google] looked at Google trend search queries related to education, but only with a limited number of search terms or with different focuses.
 
@@
 Google Trend search queries showed a strong cross correlation between COVID-19 related search terms and distance education related search terms.
-The high popularity and increasing public attention towards different distance learning solutions during COVID-19 pandemic was also investigated through a detailed Google Trend footprint analysis.
-The findings will help the educators to narrow down their search space and thereby select prominent options available in the market for different education purposes.
+The high popularity and increasing public attention towards different distance learning solutions during the COVID-19 pandemic was also investigated through a detailed Google Trend footprint analysis.
+The findings will help educators to narrow down their search space and thereby select prominent options available in the market for different educational purposes.
 According to @wallace2003online, the learning tools can have a significant impact on how students engage in online learning.
-The finding will also helpful for developers to identify their competitors in the market.
-High cost involved with different distance learning solutions, lack of financial support/funding and lack of understanding about the existing distance learning solutions are some of the major issues the educational institutions face with the COVID-19 pandemic.
-Therefore, the findings of this study offer important insights to  various stakeholders of education  to make better decision with less  time and effort.
+The finding will also be helpful for developers to identify their competitors in the market.
+The high costs involved with different distance learning solutions, lack of financial support/funding, and lack of understanding about the existing distance learning solutions are some of the major issues educational institutions face with the COVID-19 pandemic.
+Therefore, the findings of this study offer important insights to  various stakeholders in education  to make better decisions with less  time and effort.
 
-Google Trends only captures the search behaviour of a sub population with Internet access, eventhough it  is the most common worldwide search engine.
-Further, our study focused on global attention toward distance learning solutions during COVID-19 pandemic.
+Google Trends only captures the search behaviour of a sub-population with Internet access, even though it  is the most common worldwide search engine.
+Furthermore, our study focused on global attention toward distance learning solutions during the COVID-19 pandemic.
 Therefore, care should be given when generalizing these findings to different regions or different time periods, as different regions can have different market orientation and seasonal requirements.
 This can lead to variations in the search output and in turn the study findings for different regions and time frames.
-Further, Google Trends data were not available as absolute values and available only in the form of relative volume.
-Therfore the scores ranges from 0 to 100 do not allow the researchers to make meaningful comparisons across different education needs.
+Furthermore, Google Trends data is not available as absolute values but only in the form of relative volume.
+Therefore, the scores range from 0 to 100, which does not allow the researchers to make meaningful comparisons across different education needs.
 Further investigation is needed at national and regional level to identify the specific measures that can be taken to ensure the quality of distance education at national and regional level.

### 2021-12-28 Paper_online_Learning.tex
minor changes to the abstract

@@
 Distance education has a long history.
-However, COVID-19 has made a new era of distance education.
+However, COVID-19 has created a new era of distance education.
 Due to the increasing demand, various distance learning solutions have been introduced for different distance education purposes.
@@
 This can provide a fast first step guide to identifying the most popular online learning tools available for different educational purposes.
-The results allow the teachers to narrow down the search space and deepen their exploration on prominent distance education solutions to support their online teaching.
+The results allow the teachers to narrow down the search space and deepen their exploration of prominent distance education solutions to support their online teaching.
 The R code and data to reproduce the results of this work are available in the online supplementary materials.

### 2021-12-28 Paper_online_Learning.tex
Updated

@@
 alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}.
-Sudden unexpected movement from classrooms to home schooling at large scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.
+Sudden unexpected movement from classrooms to home schooling at large-scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.
 Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.
@@
 According to \textcite{carter2021teacher}, for educators to continue to teach inclusively, they need support to connect with the students through a variety of strategies so that students feel not only connected to their teachers, but to the subject matter being taught.
-In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
+In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment, UNESCO took immediate and timely action by publishing a long list of distance learning solutions \autocite{unesco2020DLsolutions} that can be used to facilitate student learning during the period of school closure.
 There are many different distance learning solutions.
@@
 
-\hypertarget{background}{%
-\section{Background}\label{background}}
+\hypertarget{related-work}{%
+\section{Related work}\label{related-work}}
 
@@
 This confirms the life cycle of Google Trend search patterns introduced in \textcite{jarynowski2020perception}.
-Further, a second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
+Further, the second peak was observed only in education related terms after the WHO announcement of the COVID-19 pandemic.
 This could be due to the nearly total closures of schools that happened with the declaration of COVID-19 as a pandemic.
@@
 series (Figures \ref{fig:dtw1} (b) and \ref{fig:dtw2} (b)).
-In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide \autocite{ciotti2019covid}.When the virus was known as ``corona'', it was not a global concern and was mostly limited to China.
+In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide \autocite{ciotti2019covid}.
+When the virus was known as ``corona'', it was not a global concern and was mostly limited to China.
 During that time, people did not pay much attention to online learning as conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic.
@@
 A significant peak in `Ubongo' series can be observed during the 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
-Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as the tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+Under systems with strong offline functionality, `Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic (Figure \ref{fig:plot1} (c)) as a tool focuses on facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
 These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic.

### 2021-12-28 main.Rmd
Updated

@@
 alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].
-Sudden unexpected movement from classrooms to home schooling at large scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
+Sudden unexpected movement from classrooms to home schooling at large-scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
 Surprisingly, the search spikes of distance education related searches (Figure \ref{fig:distanceLearningWorldAnalysis} (c)) coincide with increasing COVID-19 counts (Figure \ref{fig:distanceLearningWorldAnalysis} (a)) and related internet searches (Figure \ref{fig:distanceLearningWorldAnalysis} (b)), even though distance education has a long history in contrast to the COVID-19 pandemic.
@@
 According to  @carter2021teacher, for educators to continue to teach inclusively, they need support to connect with the students through  a variety of strategies so that  students feel not only connected to their teachers, but to the subject matter being taught.
-In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions that can be used to facilitate student learning during the period of school closure.
+In response to this gray situation and the urgent requirement for a massive transformation from a physical classroom to a virtual learning environment,  UNESCO took immediate and timely action by publishing a long list of distance learning solutions [@unesco2020DLsolutions] that can be used to facilitate student learning during the period of school closure.
 There are many different distance learning solutions.
@@
 
-# Background
+# Related work
 
@@
 This confirms the life cycle of Google Trend search patterns introduced in @jarynowski2020perception.
-Further,  a second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
+Further,  the second peak was observed only in education related terms  after the WHO announcement of the COVID-19 pandemic.
 This could be due to the nearly total closures of schools that happened  with the declaration of  COVID-19 as a pandemic.
@@
 series (Figures \ref{fig:dtw1} (b) and  \ref{fig:dtw2} (b)).
-In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide [@ciotti2019covid].When the virus was known as "corona", it was not a global concern and was mostly limited to China.
+In late 2019 the infection was recognized as a virus which is genetically related to the coronavirus responsible for the Severe Acute Respiratory Syndrome (SARS) outbreak of 2003, which infected over 8,000 people from different countries and resulted in more than 700 deaths worldwide [@ciotti2019covid].
+When the virus was known as "corona", it was not a global concern and was mostly limited to China.
 During that time, people did not pay much attention to online learning as  conventional classroom learning approach was the most popular teaching learning strategy prior to the COVID-19 pandemic.
@@
 A significant peak in  'Ubongo' series can be observed during the 2020 year end academic break in Africa (Figure \ref{fig:covidImpactContinent}).
-Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  the tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
+Under systems with strong offline functionality, 'Kolibri', which is available in more than 20 languages, and supports universal education, has gotten considerable attention during the COVID-19 pandemic  (Figure \ref{fig:plot1} (c)) as  a tool focuses on  facilitating learners living in underserved contexts where the internet is costly, unreliable, or unreachable.
 These distance learning solutions can provide a huge relief for financially distressed families during the COVID-19 pandemic.

### 2021-12-28 Paper_online_Learning.tex
update intro

@@
 Even though distance education has a long history that goes back almost two centuries with significant modifications,
-alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}.
+alterations, and additions in the process of delivery and communication \autocite{moore2011learning,spector2014handbook,zawacki2016mapping}, COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously \autocite{richmond2020critical}.
 Sudden unexpected movement from classrooms to home schooling at large-scale made students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) \autocite{carter2021teacher}.

### 2021-12-28 main.Rmd
update intro

@@
 Even though distance education has a long history that goes back  almost two centuries with significant  modifications,
-alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].
+alterations, and additions in the process of delivery and communication  [@moore2011learning; @spector2014handbook; @zawacki2016mapping],  COVID-19 certainly created a new era of distance education while encouraging various stakeholders in education to take the concept of distance education seriously [@richmond2020critical].
 Sudden unexpected movement from classrooms to home schooling at large-scale made  students, teachers and parents vulnerable, leading to millions of education related internet searches being performed during the period of COVID-19 pandemic (Figures \ref{fig:distanceLearningWorldAnalysis}) [@carter2021teacher].
